### 2022-04-25 proposal/proposal.tex
adding proposal to github

%
% File acl2020.tex
%
%% Based on the style files for ACL 2020, which were
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2020}
\usepackage{times}
\usepackage{latexsym}
\renewcommand{\UrlFont}{\ttfamily\small}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B\textsc{ib}\TeX}

\title{Proposal for Analyzing Language Models: Separatablity of Syntax and Semantics}

\author{Qingxia Guo \and Saiya Karamali \and Lindsay Skinner \and Gladys Wang
 \\ University of Washington \\ 
\texttt{qg07@uw.edu}\\ \texttt{karamali@uw.edu}  \\ \texttt{skinnel@uw.edu} \\ \texttt{@uw.edu}}

\date{}

\begin{document}

\maketitle
\begin{abstract}
This document contains the instructions for preparing a manuscript for the proceedings of ACL 2020.
The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like.
These instructions should be used for both papers submitted for review and for final versions of accepted papers.
Authors are asked to conform to all the directions reported in this document.
\end{abstract}

\section{Introduction}

The following instructions are directed to authors of papers submitted to ACL 2020 or accepted for publication in its proceedings.
All authors are required to adhere to these specifications.
Authors are required to provide a Portable Document Format (PDF) version of their papers.
\textbf{The proceedings are designed for printing on A4 paper.}

\section{Methods}

\section{Possible Results}
\label{sec:length}

The conference accepts submissions of long papers and short papers.
Long papers may consist of up to eight (8) pages of content plus unlimited pages for references.
Upon acceptance, final versions of long papers will be given one additional page -- up to nine (9) pages of content plus unlimited pages for references -- so that reviewers' comments can be taken into account.
Short papers may consist of up to four (4) pages of content, plus unlimited pages for references.
Upon acceptance, short papers will be given five (5) pages in the proceedings and unlimited pages for references. 
For both long and short papers, all illustrations and tables that are part of the main text must be accommodated within these page limits, observing the formatting instructions given in the present document.
Papers that do not conform to the specified length and formatting requirements are subject to be rejected without review.

The conference encourages the submission of additional material that is relevant to the reviewers but not an integral part of the paper.
There are two such types of material: appendices, which can be read, and non-readable supplementary materials, often data or code.
Additional material must be submitted as separate files, and must adhere to the same anonymity guidelines as the main paper.
The paper must be self-contained: it is optional for reviewers to look at the supplementary material.
Papers should not refer, for further detail, to documents, code or data resources that are not available to the reviewers.
Refer to Appendices~\ref{sec:appendix} and \ref{sec:supplemental} for further information.

Workshop chairs may have different rules for allowed length and whether supplemental material is welcome.
As always, the respective call for papers is the authoritative source.

\section{Divsion of Labor and Timeline}
As reviewing will be double-blind, papers submitted for review should not include any author information (such as names or affiliations). Furthermore, self-references that reveal the author's identity, \emph{e.g.},
\begin{quote}
We previously showed \citep{Gusfield:97} \ldots
\end{quote}
should be avoided. Instead, use citations such as 
\begin{quote}
\citet{Gusfield:97} previously showed\ldots
\end{quote}
Please do not use anonymous citations and do not include acknowledgements.
\textbf{Papers that do not conform to these requirements may be rejected without review.}

Any preliminary non-archival versions of submitted papers should be listed in the submission form but not in the review version of the paper.
Reviewers are generally aware that authors may present preliminary versions of their work in other venues, but will not be provided the list of previous presentations from the submission form.

Once a paper has been accepted to the conference, the camera-ready version of the paper should include the author's names and affiliations, and is allowed to use self-references.

\paragraph{\LaTeX-specific details:}
For an anonymized submission, ensure that {\small\verb|\aclfinalcopy|} at the top of this document is commented out, and that you have filled in the paper ID number (assigned during the submission process on softconf) where {\small\verb|***|} appears in the {\small\verb|\def\aclpaperid{***}|} definition at the top of this document.
For a camera-ready submission, ensure that {\small\verb|\aclfinalcopy|} at the top of this document is not commented out.

\section*{Acknowledgments}

The acknowledgments should go immediately before the references. Do not number the acknowledgments section.
Do not include this section when submitting your paper for review.

\bibliography{anthology,acl2020}
\bibliographystyle{acl_natbib}

\appendix

\section{Appendices}
\label{sec:appendix}
Appendices are material that can be read, and include lemmas, formulas, proofs, and tables that are not critical to the reading and understanding of the paper. 
Appendices should be \textbf{uploaded as supplementary material} when submitting the paper for review.
Upon acceptance, the appendices come after the references, as shown here.

\paragraph{\LaTeX-specific details:}
Use {\small\verb|\appendix|} before any appendix section to switch the section numbering over to letters.

\section{Supplemental Material}
\label{sec:supplemental}
Submissions may include non-readable supplementary material used in the work and described in the paper.
Any accompanying software and/or data should include licenses and documentation of research review as appropriate.
Supplementary material may report preprocessing decisions, model parameters, and other details necessary for the replication of the experiments reported in the paper.
Seemingly small preprocessing decisions can sometimes make a large difference in performance, so it is crucial to record such decisions to precisely characterize state-of-the-art methods.

Nonetheless, supplementary material should be supplementary (rather than central) to the paper.
\textbf{Submissions that misuse the supplementary material may be rejected without review.}
Supplementary material may include explanations or details of proofs or derivations that do not fit into the paper, lists of
features or feature templates, sample inputs and outputs for a system, pseudo-code or source code, and data.
(Source code and data should be separate uploads, rather than part of the paper).

The paper should not rely on the supplementary material: while the paper may refer to and cite the supplementary material and the supplementary material will be available to the reviewers, they will not be asked to review the supplementary material.

\end{document}

### 2022-04-26 proposal/proposal.tex
Added methods section to proposal
Added the methods section to the proposal. Still need to add citations and get feedback from others in the group regarding some statements that were made.

@@
 \section{Methods}
+
+We construct two separate probing tasks to isolate the syntactic and semantic information in word-level BERT embeddings.
+The embeddings are separated into syntactic and non-syntactic, and semantic and non-semantic components via Iterative Null-Space Projection (INLP), which is described in section 2.1.
+These embedding components are then combined to form new embeddings, which are evaluated on the same tasks that were used for probing.
+If time permits, we will also evaluate how these new embeddings perform on the language modeling task.
+%cite BERT paper maybe?
+%cite INLP paper
+
+\subsection{The Iterative Null-Space Projection method}
+
+The Iterative Null-Space Projection method (INLP from here on) is used to create a guarding function that masks all the linear information contained in a set of vectors, $X$, that can be used map each vector to its affiliated category, $c \in C$.
+This is accomplished by training a linear classifier, a matrix $W$, that is applied to each $x \in X$ in order to predict the correct category $c$ with the greatest possible accuracy.
+In other words, $Wx$ defines a distribution over the set of categories $C$ and we assign $x$ to the class $c \in C$ which is allotted the greatest probability by $Wx$.
+Note that the classifier's accuracy must be greater than random chance, otherwise $x$ contains no linear information relevant for the categorization task and thus no guarding function is needed.
+Once $W$ is determined, for any $x \in X$ we can remove the information that $W$ uses to predict $c$ by projecting $x$ onto the null-space of $W$, $N(W) = \{x | Wx=0\}$.
+Call this projection function $P_1$ and let $\hat{x} = P_1(x)$.
+This removes all of the linear information in $x$ that $W$ used to predict the category $c$.
+%Need to cite INLP paper
+
+However, this process does not necessarily remove all of the linear information in $x$ that could be used to predict $c$.
+For example, $x$ may contain redundant information and $W$ may have only used one set of this information for its prediction.
+In this case, the redundant information would still be present in $\hat{x}$.
+Thus, we must repeat the above process, defining a new linear classifier $\hat{W}$ that uses $\hat{x}$  to predict $c$.
+If $\hat{W}$ is still able to predict $c$ with a greater than random chance accuracy, then we know that $\hat{x}$ contained linear information about $c$.
+As above, we project $\hat{x}$ onto the null-space of $\hat{W}$ via the projection function $P_2$ and define a new $\hat{x} = P_2(P_1(x))$.
+
+We iteratively apply this process until no linear information remains in $\hat{x}$, i.e.
+a linear classifier is unable to predict the correct category $c$ with any probability greater than random chance.
+The final $\hat{x} = P_n(P_{n-1}(\dots P_1(x)))$ contains no linear information about the categories in $C$ and we call $P(x) = P_n(P_{n-1}(\dots P_1(x)))$ the guarding function.
+
+We will pair the INLP method with the probing tasks described in sections 2.3 and 2.4 in order to create two guarding functions that will enable us to isolate components of BERT embeddings that syntax-specific and semantics-specific information.
+
+\subsection{Data}
+
+%Description of the data set used and the information it contains relevant to the tasks of interest. Talk about pre-processing the syntax trees to get CCG tags. 
+We will use the English V4 subset of the CoNLL 2012 shared task data.
+We must perform two pre-processing steps for this data to be used for our probing and evaluation tasks.
+The first is to apply Hockenmaier and Steedman's CCG Derivation algorithm to the parse tree field in the dataset, in order to create the CCG tags for each word.
+If this proves to be too time-consuming or computationally expensive then we shall change the syntactic probing task to utilize the POS tags available in the dataset in place of CCG tags.
+The second task is to use the SRL frames in the dataset to generate (verb, BIO-argument-tag) pairs that will act as the categories for the semantic probing task.
+%cite CCG paper
+
+\subsection{Syntactic probing task}
+
+The syntactic probing task involves training a linear classifier on the final layer BERT embeddings in order to predict the CCG tag associated with each word.
+We will use this classifier in the INLP algorithm in order to create a guarding function for the information that is necessary to complete the CCG labeling task.
+For a given embedding, the projection that results from applying this guarding function to the embedding will represent the non-syntactic information contained in the embedding and will from now on be referred to as the ``non-syntactic component'' of the embedding.
+We can then determine the ``syntactic component'' of the embedding by taking the difference of the embedding vector with the non-syntactic component.
+
+\subsection{Semantic probing task}
+
+Similar to the above, the semantic probing tasks involves training a linear classifier on the final layer BERT embeddings in order to predict the semantic role tag (described in the data section) associated with each word.
+This classifier is used in the INLP algorithm in order to create a guarding function for the information necessary to complete the SRL labeling task.
+For this particular task, it is possible that a single word will have multiple SRL tags associated with it.
+In this case, we treat each (vector, SRL tag) pair as a single example in the dataset, in order to create a guarding function that works across all of the SRL tags affiliated with a particular word.
+As described in the Syntactic probing task section, we shall use the resulting guarding function to decompose the original embedding into a ``semantic component'' and a ``non-semantic component''.
+
+\subsection{Evaluation tasks}
+
+Our goal is to determine which information sets captured in the BERT embeddings are relevant for our evaluation tasks.
+We thus use the components derived from the probing tasks to create new embeddings that isolate specific types of information.
+These embeddings are then evaluated on the syntactic and semantic tasks that were used for probing, and their performance is compared to that of the original embeddings.
+
+The new embeddings to be tested include the syntactic component, the non-syntactic component, the semantic component and the non-semantic component derived from the probing tasks.
+Additionally, we can create an embedding that contains syntactic information and removes semantic information by linearly projecting the syntactic component onto the non-semantic component.
+Using a similar process, we can create an embedding that contains semantic information and removes the syntactic information present.
+Finally, we can create an embedding that contains the semantic information captured by the syntactic component, by linearly projecting the syntactic component onto the semantic component.
+Similarly, we can create an embedding that contains the syntactic information captured by the semantic component.
+
+We will assess each of these embedding types and the original BERT embeddings on the CCG and SRL labeling tasks that were used in the probes.
+We will pay particular attention to accuracy and training time when comparing the various embedding types.
+
+We have also hypothesized several additional assessment tasks that we would like to undertake, if time permits, or relegate to future work.
+The first of these tasks is to assess how each embedding type performs on the language modeling task.
+We would also like to perform the evaluation classification task using a feed-forward neural network with a single hidden layer that contains 10 nodes, in order to determine if there is any task-relevant non-linear information present in the embeddings.
+If time permits, we would also like to look for patterns in the performance of different embeddings, e.g.
+explore if a particular embedding type tends to perform better/worse on one of the evaluation tasks for words of a particular POS compared to others.
+Finally, if time permits we would like to repeat the above procedure to explore the embeddings output by different layers of the BERT model.
+%FFNN description is entirely arbitrary, it was just something we talked about during one of our meetings.
+If anyone has a particular architecture in mind here, please feel free to update this section.
 

### 2022-04-26 proposal/proposal.tex
minor updates to methods
Removed a sentence in evaluation that incorrectly claimed we would look at train time for the linear classifier. Minor grammatical changes to the rest of the methods section.

@@
 
-We will pair the INLP method with the probing tasks described in sections 2.3 and 2.4 in order to create two guarding functions that will enable us to isolate components of BERT embeddings that syntax-specific and semantics-specific information.
+We will pair the INLP method with the probing tasks described in sections 2.3 and 2.4 in order to create two guarding functions that will enable us to isolate components of BERT embeddings that contain syntax-specific and semantics-specific information.
 
@@
 The first is to apply Hockenmaier and Steedman's CCG Derivation algorithm to the parse tree field in the dataset, in order to create the CCG tags for each word.
-If this proves to be too time-consuming or computationally expensive then we shall change the syntactic probing task to utilize the POS tags available in the dataset in place of CCG tags.
+If this proves to be too time-consuming or computationally expensive then we shall change the syntactic probing task to utilize the POS tags available in the dataset, in place of CCG tags.
 The second task is to use the SRL frames in the dataset to generate (verb, BIO-argument-tag) pairs that will act as the categories for the semantic probing task.
@@
 We will assess each of these embedding types and the original BERT embeddings on the CCG and SRL labeling tasks that were used in the probes.
-We will pay particular attention to accuracy and training time when comparing the various embedding types.
-
 We have also hypothesized several additional assessment tasks that we would like to undertake, if time permits, or relegate to future work.

### 2022-04-27 proposal/proposal.tex
adding possible results

--- 2022-04-26 proposal/proposal.tex
@@
 
-\author{Qingxia Guo \and Saiya Karamali \and Lindsay Skinner \and Gladys Wang
+\author{Qingxia Guo, Saiya Karamali, Lindsay Skinner, \and Gladys Wang
  \\ University of Washington \\ 
-\texttt{qg07@uw.edu}\\ \texttt{karamali@uw.edu}  \\ \texttt{skinnel@uw.edu} \\ \texttt{@uw.edu}}
-
+\texttt{\{qg07, karamali, skinnel, qinyanw\}@uw.edu}\\ 
+}
 \date{}
@@
 
-The conference accepts submissions of long papers and short papers.
-Long papers may consist of up to eight (8) pages of content plus unlimited pages for references.
-Upon acceptance, final versions of long papers will be given one additional page -- up to nine (9) pages of content plus unlimited pages for references -- so that reviewers' comments can be taken into account.
-Short papers may consist of up to four (4) pages of content, plus unlimited pages for references.
-Upon acceptance, short papers will be given five (5) pages in the proceedings and unlimited pages for references. 
-For both long and short papers, all illustrations and tables that are part of the main text must be accommodated within these page limits, observing the formatting instructions given in the present document.
-Papers that do not conform to the specified length and formatting requirements are subject to be rejected without review.
+In this section, we consider what each evaluation task tells us about the embeddings that are being probed.
+When we isolate the syntactic component and run it on the syntactic and semantic tasks, we learn how successfully the component responsible for CCG tagging has been isolated, and we also learn how effective the syntactic component alone is on the semantic task.
+Similarly, running the semantic component on the evaluation tasks tells us how well we've isolated the semantic component and how effective it is on the syntactic task.
+Finally, running the non-syntactic and non-semantic components on the evaluation tasks tells us whether any information not identified by INLP is at all useful for the evaluation tasks.
 
-The conference encourages the submission of additional material that is relevant to the reviewers but not an integral part of the paper.
-There are two such types of material: appendices, which can be read, and non-readable supplementary materials, often data or code.
-Additional material must be submitted as separate files, and must adhere to the same anonymity guidelines as the main paper.
-The paper must be self-contained: it is optional for reviewers to look at the supplementary material.
-Papers should not refer, for further detail, to documents, code or data resources that are not available to the reviewers.
-Refer to Appendices~\ref{sec:appendix} and \ref{sec:supplemental} for further information.
-
-Workshop chairs may have different rules for allowed length and whether supplemental material is welcome.
-As always, the respective call for papers is the authoritative source.
+Next, we consider the potential results of the various projected word embeddings on the evaluation tasks.
+Each of these tell us how much overlap there is between the syntactic and semantic components of the contextual word embeddings.
+Projecting the syntactic component onto the non-semantic component removes semantic information from the syntactic component, and projecting the semantic component onto the non-syntactic component removes syntactic information from the semantic component.
+Projecting the syntactic component onto the semantic component gives us the semantic information that is also part of the syntactic component, and projecting the semantic component onto the syntactic component gives us the syntactic information that is also part of the semantic component.
+Running all of these embeddings un the semantic and syntactic tasks tells us how separated the semantic and syntactic components are, and how important the overlapping portions are to each task.
 

### 2022-04-27 proposal/proposal.tex
adding introduction section

@@
 \maketitle
-\begin{abstract}
-This document contains the instructions for preparing a manuscript for the proceedings of ACL 2020.
-The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like.
-These instructions should be used for both papers submitted for review and for final versions of accepted papers.
-Authors are asked to conform to all the directions reported in this document.
-\end{abstract}
 
@@
 
-The following instructions are directed to authors of papers submitted to ACL 2020 or accepted for publication in its proceedings.
-All authors are required to adhere to these specifications.
-Authors are required to provide a Portable Document Format (PDF) version of their papers.
-\textbf{The proceedings are designed for printing on A4 paper.}
+The boundary between semantics and syntax has always been a hotly debated topic in linguistics since generative syntax, but do NLP models make such a distinction?
+The objective of this project is to explore BERT\rq s reliance on certain syntactic information when handling a semantic task, and vice versa.
+To achieve our goal, we will construct a linear probing system for a task and then employ Iterative Nullspace Projection (INLP) \citep{inlp} to generate a new embedding devoid of information learned from probing task.
+Then we will measure the performance of another task on this new embedding.
+
+The design of our probing procedure follows \citealp{amnesia}, who employed Iterative Nullspace Projection (INLP from here) to investigate whether BERT uses part-of-speech (PoS) information when solving language modeling (LM) tasks.
+A novel method for removing information from a embedding, INLP iteratively trains a linear model on a specific task, and projects the input on the nullspace of the linear model.
+Our objective is that, by applying the INLP procedure to a syntactic task, we are able to separate the representation into a syntactic space and a non-syntactic space.
+Then we can compare the performance of a semantic task in whole space and only in the non-syntactic space to see if BERT is using syntactic information when performing the semantic task.
+Conversely, we can also first probe a semantic task and then measure with a syntactic task.
+
+To evaluate the separability of syntax and semantics, we will need two task, one task for the probing system and INLP procedure, one task for measure performance on embeddings before and after INLP.
+We choose Combinatory Categorical Grammar (CCG) tagging as the syntactic task and Semantic Role Labeling (SRL) as the semantic task.
+Section \ref{sec:method} provides description of the tasks and the construction of the experiment pipeline.
+Section \ref{sec:result} provddes possible results and interpretations.
+Finally, section \ref{sec:management} provides project management details including timeline and division of labor.
+% If the removing of syntactic information leads to a insignificant decrease in the model\rq s performance in a semantic task (or removing semantic information leads to insignificant decrease in performance in a syntactic task), we can conclude that syntactic and semantic information in BERT\rq s representation is linearly separable.
+Conversely, an significant decrease in performance will indicate low separability between syntactic and semantic information in BERT \rq s representation.
 
 \section{Methods}
+\label{sec:method}
 
-We construct two separate probing tasks to isolate the syntactic and semantic information in word-level BERT embeddings.
+We construct two separate probing tasks to isolate the syntactic and semantic information in word-level BERT embeddings\citep{bert}.
 The embeddings are separated into syntactic and non-syntactic, and semantic and non-semantic components via Iterative Null-Space Projection (INLP), which is described in section 2.1.
@@
 \section{Possible Results}
-\label{sec:length}
+\label{sec:result}
 
@@
 \section{Divsion of Labor and Timeline}
-As reviewing will be double-blind, papers submitted for review should not include any author information (such as names or affiliations).
-Furthermore, self-references that reveal the author's identity, \emph{e.g.},
-\begin{quote}
-We previously showed \citep{Gusfield:97} \ldots
-\end{quote}
-should be avoided. Instead, use citations such as 
-\begin{quote}
-\citet{Gusfield:97} previously showed\ldots
-\end{quote}
-Please do not use anonymous citations and do not include acknowledgements.
-\textbf{Papers that do not conform to these requirements may be rejected without review.}
+\label{sec:management}
+this is
 
-Any preliminary non-archival versions of submitted papers should be listed in the submission form but not in the review version of the paper.
-Reviewers are generally aware that authors may present preliminary versions of their work in other venues, but will not be provided the list of previous presentations from the submission form.
-
-Once a paper has been accepted to the conference, the camera-ready version of the paper should include the author's names and affiliations, and is allowed to use self-references.
-
-\paragraph{\LaTeX-specific details:}
-For an anonymized submission, ensure that {\small\verb|\aclfinalcopy|} at the top of this document is commented out, and that you have filled in the paper ID number (assigned during the submission process on softconf) where {\small\verb|***|} appears in the {\small\verb|\def\aclpaperid{***}|} definition at the top of this document.
-For a camera-ready submission, ensure that {\small\verb|\aclfinalcopy|} at the top of this document is not commented out.
-
-\section*{Acknowledgments}
-
-The acknowledgments should go immediately before the references. Do not number the acknowledgments section.
-Do not include this section when submitting your paper for review.
-
-\bibliography{anthology,acl2020}
+\bibliography{acl2020}
 \bibliographystyle{acl_natbib}
 
-\appendix
-
-\section{Appendices}
-\label{sec:appendix}
-Appendices are material that can be read, and include lemmas, formulas, proofs, and tables that are not critical to the reading and understanding of the paper. 
-Appendices should be \textbf{uploaded as supplementary material} when submitting the paper for review.
-Upon acceptance, the appendices come after the references, as shown here.
-
-\paragraph{\LaTeX-specific details:}
-Use {\small\verb|\appendix|} before any appendix section to switch the section numbering over to letters.
-
-\section{Supplemental Material}
-\label{sec:supplemental}
-Submissions may include non-readable supplementary material used in the work and described in the paper.
-Any accompanying software and/or data should include licenses and documentation of research review as appropriate.
-Supplementary material may report preprocessing decisions, model parameters, and other details necessary for the replication of the experiments reported in the paper.
-Seemingly small preprocessing decisions can sometimes make a large difference in performance, so it is crucial to record such decisions to precisely characterize state-of-the-art methods.
-
-Nonetheless, supplementary material should be supplementary (rather than central) to the paper.
-\textbf{Submissions that misuse the supplementary material may be rejected without review.}
-Supplementary material may include explanations or details of proofs or derivations that do not fit into the paper, lists of
-features or feature templates, sample inputs and outputs for a system, pseudo-code or source code, and data.
-(Source code and data should be separate uploads, rather than part of the paper).
-
-The paper should not rely on the supplementary material: while the paper may refer to and cite the supplementary material and the supplementary material will be available to the reviewers, they will not be asked to review the supplementary material.
-
 \end{document}

### 2022-04-27 proposal/proposal.tex
adding citation

@@
 
-To evaluate the separability of syntax and semantics, we will need two task, one task for the probing system and INLP procedure, one task for measure performance on embeddings before and after INLP.
+To evaluate the separability of syntax and semantics, we will need two task, one task for the probing system and INLP procedure, one task for evaluating performance on embeddings before and after INLP.
 We choose Combinatory Categorical Grammar (CCG) tagging as the syntactic task and Semantic Role Labeling (SRL) as the semantic task.
 Section \ref{sec:method} provides description of the tasks and the construction of the experiment pipeline.
-Section \ref{sec:result} provddes possible results and interpretations.
+Section \ref{sec:result} provides possible results and interpretations.
 Finally, section \ref{sec:management} provides project management details including timeline and division of labor.
-% If the removing of syntactic information leads to a insignificant decrease in the model\rq s performance in a semantic task (or removing semantic information leads to insignificant decrease in performance in a syntactic task), we can conclude that syntactic and semantic information in BERT\rq s representation is linearly separable.
-Conversely, an significant decrease in performance will indicate low separability between syntactic and semantic information in BERT \rq s representation.
 
@@
 %Description of the data set used and the information it contains relevant to the tasks of interest. Talk about pre-processing the syntax trees to get CCG tags. 
-We will use the English V4 subset of the CoNLL 2012 shared task data.
+We will use the English V4 subset of the CoNLL 2012 shared task data\citep{2012-conll} .
 We must perform two pre-processing steps for this data to be used for our probing and evaluation tasks.
-The first is to apply Hockenmaier and Steedman's CCG Derivation algorithm to the parse tree field in the dataset, in order to create the CCG tags for each word.
+The first is to apply \citet{ccg-bank}'s CCG Derivation algorithm to the parse tree field in the dataset, in order to create the CCG tags for each word.
 If this proves to be too time-consuming or computationally expensive then we shall change the syntactic probing task to utilize the POS tags available in the dataset, in place of CCG tags.
@@
 
+% If the removing of syntactic information leads to a insignificant decrease in the model\rq s performance in a semantic task (or removing semantic information leads to insignificant decrease in performance in a syntactic task), we can conclude that syntactic and semantic information in BERT\rq s representation is linearly separable.
+Conversely, an significant decrease in performance will indicate low separability between syntactic and semantic information in BERT \rq s representation.
+
 \section{Divsion of Labor and Timeline}
 \label{sec:management}
-this is
+There are three main parts of this project, coding the experiment, presenting related topics in class and writing the final paper. Each part has detailed roles and
 

### 2022-04-27 proposal/proposal.tex
adding the time table

@@
 \usepackage{latexsym}
+\usepackage{multirow}
 \renewcommand{\UrlFont}{\ttfamily\small}
@@
 Similarly, running the semantic component on the evaluation tasks tells us how well we've isolated the semantic component and how effective it is on the syntactic task.
-Finally, running the non-syntactic and non-semantic components on the evaluation tasks tells us whether any information not identified by INLP is at all useful for the evaluation tasks.
+If the removing of syntactic information leads to a insignificant decrease in the model\rq s performance in a semantic task (or removing semantic information leads to insignificant decrease in performance in a syntactic task), we can conclude that syntactic and semantic information in BERT\rq s representation is linearly separable.
+Conversely, an significant decrease in performance will indicate low separability between syntactic and semantic information in BERT \rq s representation.Finally, running the non-syntactic and non-semantic components on the evaluation tasks tells us whether any information not identified by INLP is at all useful for the evaluation tasks.
 
@@
 
-% If the removing of syntactic information leads to a insignificant decrease in the model\rq s performance in a semantic task (or removing semantic information leads to insignificant decrease in performance in a syntactic task), we can conclude that syntactic and semantic information in BERT\rq s representation is linearly separable.
-Conversely, an significant decrease in performance will indicate low separability between syntactic and semantic information in BERT \rq s representation.
-
 \section{Divsion of Labor and Timeline}
 \label{sec:management}
-There are three main parts of this project, coding the experiment, presenting related topics in class and writing the final paper. Each part has detailed roles and
+There are three main parts of this project, coding the experiment, presenting related topics in class and writing the final paper.
+Each part has detailed requirements and been assigned to a member to take charge of as presented in table \ref{role description}.
+It is hard to anticipate the time consumed and difficulty of each task.
+We have set up a regular weekly meeting to update the progress of each part weekly so that everyone is on the same page with the development of the project.
+
+\begin{table*}[ht]
+    \centering
+    \begin{tabular}{llll}
+    \hline
+    \textbf{Duty} & \textbf{Details} & \textbf{Who is in Charge} \\
+    \hline
+    \multirow{5}{*}{Coding the experiment} & data preprocessing & Qingxia \\
+    & building probing system \& INLP & Saiya \& Qingxia \\
+    & building evaluation system & Gladys \\
+    & merging subparts together & Saiya\\ 
+    & running different task pairs & Saiya \\
+    \hline
+    \multirow{2}{*}{Presenting related topic} & \multirow{2}{*}{presenting INLP method} & \multirow{2}{*}{Lindsay \& Gladys} \\
+    & & \\
+    \hline
+    \multirow{3}{*}{Writing paper} & method section  & Lindsay \\
+     & result section & Saiya  \\
+     & discussion/conclusion section & all\\
+    \hline
+    \end{tabular}
+    \caption{\label{role description} Role Description
+    }
+    \end{table*}
 

### 2022-04-27 proposal/proposal.tex
adding due dates to timetable

@@
 There are three main parts of this project, coding the experiment, presenting related topics in class and writing the final paper.
-Each part has detailed requirements and been assigned to a member to take charge of as presented in table \ref{role description}.
-It is hard to anticipate the time consumed and difficulty of each task.
+Each part has detailed requirements along with due dates and been assigned to a member to take charge of as presented in table \ref{role description}.
+It is hard to anticipate the time consumed and difficulty of each task, so all the due dates are not final.
 We have set up a regular weekly meeting to update the progress of each part weekly so that everyone is on the same page with the development of the project.
@@
     \hline
-    \textbf{Duty} & \textbf{Details} & \textbf{Who is in Charge} \\
+    \textbf{Duty} & \textbf{Details} & \textbf{Who is in Charge} &\textbf{Due} \\
     \hline
-    \multirow{5}{*}{Coding the experiment} & data preprocessing & Qingxia \\
-    & building probing system \& INLP & Saiya \& Qingxia \\
-    & building evaluation system & Gladys \\
-    & merging subparts together & Saiya\\ 
-    & running different task pairs & Saiya \\
+    \multirow{5}{*}{Coding the experiment} & data preprocessing & Qingxia & 05/04 \\
+    & building probing system \& INLP & Saiya \& Qingxia & 05/25 \\
+    & building evaluation system & Gladys & 05/25 \\
+    & merging subparts together & Saiya & 06/02\\ 
+    & running different task pairs & Saiya & 06/02\\
     \hline
-    \multirow{2}{*}{Presenting related topic} & \multirow{2}{*}{presenting INLP method} & \multirow{2}{*}{Lindsay \& Gladys} \\
-    & & \\
+    \multirow{2}{*}{Presenting related topic} & \multirow{2}{*}{presenting INLP method} & \multirow{2}{*}{Lindsay \& Gladys} & \multirow{2}{*}{05/11}\\
+    & & & \\
     \hline
-    \multirow{3}{*}{Writing paper} & method section  & Lindsay \\
-     & result section & Saiya  \\
-     & discussion/conclusion section & all\\
+    \multirow{3}{*}{Writing paper} & method section  & Lindsay & \multirow{3}{*}{06/09}  \\
+     & result section & Saiya &  \\
+     & discussion/conclusion section & all & \\
     \hline

### 2022-04-28 proposal/proposal.tex
fixing some citation

@@
 The boundary between semantics and syntax has always been a hotly debated topic in linguistics since generative syntax, but do NLP models make such a distinction?
-The objective of this project is to explore BERT\rq s reliance on certain syntactic information when handling a semantic task, and vice versa.
-To achieve our goal, we will construct a linear probing system for a task and then employ Iterative Nullspace Projection (INLP) \citep{inlp} to generate a new embedding devoid of information learned from probing task.
+The objective of this project is to explore BERT\rq s\citep{bert}reliance on certain syntactic information when handling a semantic task, and vice versa.
+To achieve our goal, we will construct a linear probing system for a task and then employ Iterative Nullspace Projection (INLP from here on) \citep{inlp} to generate a new embedding devoid of information learned from probing task.
 Then we will measure the performance of another task on this new embedding.
 
-The design of our probing procedure follows \citealp{amnesia}, who employed Iterative Nullspace Projection (INLP from here) to investigate whether BERT uses part-of-speech (PoS) information when solving language modeling (LM) tasks.
+The design of our probing procedure follows \citealp{amnesia}, who employed INLP to investigate whether BERT uses part-of-speech (PoS) information when solving language modeling (LM) tasks.
 A novel method for removing information from a embedding, INLP iteratively trains a linear model on a specific task, and projects the input on the nullspace of the linear model.
@@
 To evaluate the separability of syntax and semantics, we will need two task, one task for the probing system and INLP procedure, one task for evaluating performance on embeddings before and after INLP.
-We choose Combinatory Categorical Grammar (CCG) tagging as the syntactic task and Semantic Role Labeling (SRL) as the semantic task.
+We choose Combinatory Categorical Grammar (CCG from here on) tagging \citep{ccg-bank} as the syntactic task and Semantic Role Labeling (SRL from here on) as the semantic task.
 Section \ref{sec:method} provides description of the tasks and the construction of the experiment pipeline.
@@
 
-We construct two separate probing tasks to isolate the syntactic and semantic information in word-level BERT embeddings\citep{bert}.
-The embeddings are separated into syntactic and non-syntactic, and semantic and non-semantic components via Iterative Null-Space Projection (INLP), which is described in section 2.1.
+We construct two separate probing tasks to isolate the syntactic and semantic information in word-level BERT embeddings.
+The embeddings are separated into syntactic and non-syntactic, and semantic and non-semantic components via INLP which is described in section 2.1.
 These embedding components are then combined to form new embeddings, which are evaluated on the same tasks that were used for probing.

### 2022-04-28 proposal/proposal.tex
modifying intro

@@
 
-The boundary between semantics and syntax has always been a hotly debated topic in linguistics since generative syntax, but do NLP models make such a distinction?
-The objective of this project is to explore BERT\rq s\citep{bert}reliance on certain syntactic information when handling a semantic task, and vice versa.
-To achieve our goal, we will construct a linear probing system for a task and then employ Iterative Nullspace Projection (INLP from here on) \citep{inlp} to generate a new embedding devoid of information learned from probing task.
-Then we will measure the performance of another task on this new embedding.
+The boundary between semantics and syntax has always been a hotly debated topic in linguistics, but do NLP models make such a distinction?
+The objective of this project is to explore BERT\rq s \citep{bert} reliance on certain syntactic information when handling a semantic task, and vice versa.
+To achieve our goal, we will construct a linear probing system for a task and then employ Iterative Nullspace Projection (INLP from here on) \citep{inlp} to generate a new embedding devoid of information learned from the probing task.
+Then we will measure the performance of this new embedding on various tasks.
 
-The design of our probing procedure follows \citealp{amnesia}, who employed INLP to investigate whether BERT uses part-of-speech (PoS) information when solving language modeling (LM) tasks.
+The design of our probing procedure follows \citealp{amnesia}, who employed INLP to investigate whether BERT uses part-of-speech (POS) information when solving language modeling (LM) tasks.
 A novel method for removing information from a embedding, INLP iteratively trains a linear model on a specific task, and projects the input on the nullspace of the linear model.
@@
 Conversely, we can also first probe a semantic task and then measure with a syntactic task.
+Once we derive the semantic space and syntactic space from the experiment, we can further conduct analysis on the separability of the two spaces.
 
 To evaluate the separability of syntax and semantics, we will need two task, one task for the probing system and INLP procedure, one task for evaluating performance on embeddings before and after INLP.
-We choose Combinatory Categorical Grammar (CCG from here on) tagging \citep{ccg-bank} as the syntactic task and Semantic Role Labeling (SRL from here on) as the semantic task.
+We choose Combinatory Categorical Grammar (CCG from here on) tagging \citep{ccg-bank} as the syntactic task and Semantic Role Labeling (SRL from here on) \citep{propbank} as the semantic task.
 Section \ref{sec:method} provides description of the tasks and the construction of the experiment pipeline.
@@
 We construct two separate probing tasks to isolate the syntactic and semantic information in word-level BERT embeddings.
-The embeddings are separated into syntactic and non-syntactic, and semantic and non-semantic components via INLP which is described in section 2.1.
+The embeddings are separated into syntactic and non-syntactic, and semantic and non-semantic components via INLP which is described in section \ref{sec:inlp-method}.
 These embedding components are then combined to form new embeddings, which are evaluated on the same tasks that were used for probing.
@@
 \subsection{The Iterative Null-Space Projection method}
+\label{sec:inlp-method}
 
-The Iterative Null-Space Projection method (INLP from here on) is used to create a guarding function that masks all the linear information contained in a set of vectors, $X$, that can be used map each vector to its affiliated category, $c \in C$.
+The INLP is used to create a guarding function that masks all the linear information contained in a set of vectors, $X$, that can be used map each vector to its affiliated category, $c \in C$.
 This is accomplished by training a linear classifier, a matrix $W$, that is applied to each $x \in X$ in order to predict the correct category $c$ with the greatest possible accuracy.
@@
 
-\begin{table*}[ht]
+\begin{table*}[h]
     \centering
@@
 
+\pagebreak
+
 \bibliography{acl2020}

### 2022-04-28 proposal/proposal.tex
modifying proposal

@@
 
-The INLP is used to create a guarding function that masks all the linear information contained in a set of vectors, $X$, that can be used map each vector to its affiliated category, $c \in C$.
+The INLP is used to create a guarding function that masks all the linear information contained in a set of vectors, $X$, that can be used map each vector to $c \in C$, where $C$ is the set of all categories.
 This is accomplished by training a linear classifier, a matrix $W$, that is applied to each $x \in X$ in order to predict the correct category $c$ with the greatest possible accuracy.
@@
 
-We will pair the INLP method with the probing tasks described in sections 2.3 and 2.4 in order to create two guarding functions that will enable us to isolate components of BERT embeddings that contain syntax-specific and semantics-specific information.
+We will pair the INLP method with the probing tasks described in sections \ref{sec:syntactic} and \ref{sec:semantics} in order to create two guarding functions that will enable us to isolate components of BERT embeddings that contain syntax-specific and semantics-specific information.
 
 \subsection{Data}
+\label{sec:data}
 
 %Description of the data set used and the information it contains relevant to the tasks of interest. Talk about pre-processing the syntax trees to get CCG tags. 
-We will use the English V4 subset of the CoNLL 2012 shared task data\citep{2012-conll} .
+We will use the English V4 subset of the CoNLL 2012 shared task data \citep{2012-conll} .
 We must perform two pre-processing steps for this data to be used for our probing and evaluation tasks.
@@
 \subsection{Syntactic probing task}
+\label{sec:syntactic}
 
@@
 \subsection{Semantic probing task}
+\label{sec:semantics}
 
@@
 \subsection{Evaluation tasks}
+\label{sec:eval}
 
@@
 Each part has detailed requirements along with due dates and been assigned to a member to take charge of as presented in table \ref{role description}.
-It is hard to anticipate the time consumed and difficulty of each task, so all the due dates are not final.
-We have set up a regular weekly meeting to update the progress of each part weekly so that everyone is on the same page with the development of the project.
+It is hard to anticipate the time consumed and difficulty of each task, so all the due dates are temporary.
+We have set up a regular weekly meeting to update the progress of each part jso that everyone is on the same page with the development of the project.
 
@@
 
-\pagebreak
-
 \bibliography{acl2020}

### 2022-04-28 proposal/proposal.tex
Updated SRL probe description
updated to deal with multiple SRL tags

@@
 For this particular task, it is possible that a single word will have multiple SRL tags associated with it.
-In this case, we treat each (vector, SRL tag) pair as a single example in the dataset, in order to create a guarding function that works across all of the SRL tags affiliated with a particular word.
+In this case, if the word has $n$ SRL tags associated with it then we will look at the $n$ most likely tags output by the linear classifier.
+We will then treat each (vector, SRL tag) pair as a single example when calculating the classifier's accuracy in order to create a guarding function that works across all of the SRL tags affiliated with a particular word.
 As described in the Syntactic probing task section, we shall use the resulting guarding function to decompose the original embedding into a ``semantic component'' and a ``non-semantic component''.

### 2022-04-28 proposal/proposal.tex
adding tables

@@
 Our objective is that, by applying the INLP procedure to a syntactic task, we are able to separate the representation into a syntactic space and a non-syntactic space.
-Then we can compare the performance of a semantic task in whole space and only in the non-syntactic space to see if BERT is using syntactic information when performing the semantic task.
+Then we can compare the performance of a semantic task in the whole space and only in the non-syntactic space to see if BERT is using syntactic information when performing the semantic task.
 Conversely, we can also first probe a semantic task and then measure with a syntactic task.
@@
 Similarly, running the semantic component on the evaluation tasks tells us how well we've isolated the semantic component and how effective it is on the syntactic task.
-If the removing of syntactic information leads to a insignificant decrease in the model\rq s performance in a semantic task (or removing semantic information leads to insignificant decrease in performance in a syntactic task), we can conclude that syntactic and semantic information in BERT\rq s representation is linearly separable.
-Conversely, an significant decrease in performance will indicate low separability between syntactic and semantic information in BERT \rq s representation.Finally, running the non-syntactic and non-semantic components on the evaluation tasks tells us whether any information not identified by INLP is at all useful for the evaluation tasks.
+If the removing of syntactic information leads to an insignificant decrease in the model\rq s performance in a semantic task (or removing semantic information leads to insignificant decrease in performance in a syntactic task), we can conclude that syntactic and semantic information in BERT\rq s representation is linearly separable.
+Conversely, a significant decrease in performance will indicate low separability between syntactic and semantic information in BERT \rq s representation.
+Finally, running the non-syntactic and non-semantic components on the evaluation tasks tells us whether any information not identified by INLP is at all useful for the evaluation tasks.
 
@@
 Projecting the syntactic component onto the semantic component gives us the semantic information that is also part of the syntactic component, and projecting the semantic component onto the syntactic component gives us the syntactic information that is also part of the semantic component.
-Running all of these embeddings un the semantic and syntactic tasks tells us how separated the semantic and syntactic components are, and how important the overlapping portions are to each task.
+Running all of these embeddings on the semantic and syntactic tasks tells us how separated the semantic and syntactic components are, and how important the overlapping portions are to each task.
 
-\section{Divsion of Labor and Timeline}
+\section{Division of Labor and Timeline}
 \label{sec:management}
@@
     \hline
-    \multirow{3}{*}{Writing paper} & method section  & Lindsay & \multirow{3}{*}{06/09}  \\
-     & result section & Saiya &  \\
-     & discussion/conclusion section & all & \\
+    \multirow{2}{*}{Writing paper} & \multirow{2}{*}{writing out experiment } & \multirow{2}{*}{Lindsay} & \multirow{2}{*}{06/09}  \\
+    & & & \\
     \hline

### 2022-06-06 src/LinearClassifier.py
adding inlp training loop

@@ -2,6 +2,9 @@ import torch
 from sklearn.metrics import accuracy_score
 from sklearn.utils.class_weight import compute_class_weight
 import numpy as np
+import scipy
+from typing import List
+import tqdm
 
 class LinearClassifier(torch.nn.Module):
 	def __init__(self, input_embeddings,output,tag_size,dev_x=None,dev_y=None):
@@ -39,8 +42,8 @@ class LinearClassifier(torch.nn.Module):
 		return dev_pred,loss.item()
 
 	
-	def optimize(self,lr=0.01,num_epochs=1000):
-		optimizer = torch.optim.SGD(self.linear.parameters(), lr = lr)
+	def optimize(self,lr=0.01,num_epochs=100):
+		optimizer = torch.optim.AdamW(self.linear.parameters(), lr = lr)
 		best_predictions = None
 		best_loss = float('inf')
 		stop_count = 0
@@ -50,34 +53,100 @@ class LinearClassifier(torch.nn.Module):
 			optimizer.zero_grad()
 			loss.backward(retain_graph=True)
 			optimizer.step()
-			print(f'epoch: {epoch+1}, loss = {loss.item():.4f}')
+			#print(f'epoch: {epoch+1}, loss = {loss.item():.4f}')
 			#implement stopping criterion
 			dev_pred,dev_loss=self.eval()
-			if dev_loss<best_loss:
-				best_loss=dev_loss
+			#print(f'epoch: {epoch+1}, dev loss = {dev_loss:.4f}')
+			if loss.item()<best_loss:
+				best_loss=loss.item()
 				best_model= self.linear
 				best_predictions = preds
-				best_dev = dev_pred
+				#best_dev = dev_pred
 				stop_count=0
 			else:
-				if stop_count ==100:
+				if stop_count ==3:
 					break
 				else:
 					stop_count+=1
 		final_pred = torch.argmax(best_predictions,dim=1).numpy()
-		#dev_pred = best_model(self.dev_x)
-		final_dev = torch.argmax(best_dev,dim=1).numpy()
+		dev_pred = best_model(self.dev_x)
 		#final_dev = torch.argmax(best_dev,dim=1).numpy()
+		final_dev = torch.argmax(dev_pred,dim=1).numpy()
 		#final_out = output.numpy()
 		#dev_out = self.dev_y.numpy()
-		print(f'train accuracy score:{accuracy_score(self.output.numpy(),final_pred):.4f}')
[…]

### 2022-06-06 src/probe.py

@@ -13,7 +13,7 @@ from utils import bert_tokenization, get_projection_to_intersection_of_nullspace
 import numpy as np
 from tqdm import tqdm
 from sklearn import preprocessing
-from LinearClassifier import LinearClassifier
+from LinearClassifier import LinearClassifier, INLPTraining
 #logging.basicConfig(level-logging.INFO) #turn on detailed logging
 
 
@@ -80,25 +80,29 @@ dev_tokens,dev_seq,dev_mask = bert_tokenization(dev_input,tokenizer)
 train_output = torch.load('../data/pmb_gold/gold_train_embeddings.pt')
 dev_output = torch.load('../data/pmb_gold/gold_dev_embeddings.pt')
 
-train_embeddings,train_y = from_sents_to_words(train_df,'sem',train_tokens,train_output)
-dev_embeddings,dev_y = from_sents_to_words(dev_df,'sem',dev_tokens,dev_output)
+train_embeddings,train_y = from_sents_to_words(train_df,'syn',train_tokens,train_output)
+dev_embeddings,dev_y = from_sents_to_words(dev_df,'syn',dev_tokens,dev_output)
+
 
 # print(train_embeddings.shape)
 # print(train_y.shape)
 # print(dev_embeddings.shape)
 # print(dev_y.shape)
 
+print(dev_embeddings.dtype)
 num_epochs = 100
 num_tags = len(np.unique(train_y.numpy()))
 input_dim = train_embeddings.shape[1] #should be 768
 
-
-lin_class = LinearClassifier(train_embeddings,train_y,num_tags,dev_x=dev_embeddings,dev_y=dev_y)
-print(lin_class.dev_x.shape)
-print(num_tags)
-#print(lin_class.output.shape)
-bm=lin_class.optimize()
+inlp_syn = INLPTraining(train_embeddings,train_y,num_tags,dev_x=dev_embeddings,dev_y=dev_y)
+inlp_syn.run_INLP_loop(10)
 sys.exit()
+# lin_class = LinearClassifier(train_embeddings,train_y,num_tags,dev_x=dev_embeddings,dev_y=dev_y)
+# print(lin_class.dev_x.shape)
+# print(num_tags)
+# #print(lin_class.output.shape)
+# bm=lin_class.optimize()
+# sys.exit()
 
 orig_embeddings = train_embeddings
 Ws = []
@@ -107,7 +111,9 @@ INLP_iterations = 3 #arbitrary choice for now
[…]

### 2022-06-06 src/LinearClassifier.py
adding eval classifier

@@ -7,7 +7,7 @@ from typing import List
 import tqdm
 
 class LinearClassifier(torch.nn.Module):
-	def __init__(self, input_embeddings,output,tag_size,dev_x=None,dev_y=None):
+	def __init__(self, input_embeddings,output,tag_size):
 		'''
 		a linear classifier for probe
 		input_embeddings : a tensor with size [batch_size,embed_dim]
@@ -24,22 +24,12 @@ class LinearClassifier(torch.nn.Module):
 		# cls_weight = compute_class_weight('balanced',classes=np.array(range(tag_size)),y=output.numpy())
 		# cls_weight = torch.tensor(cls_weight,dtype=torch.float)
 		self.loss_func = torch.nn.CrossEntropyLoss()
-		self.dev_x = dev_x
-		self.dev_y = dev_y
 	
 	def forward(self,embeddings):
 		# embedding size = [batch_size, embed_dim]
-		# output size = [batch_size]
+		# output size = [batch_size,tag_size]
 		fc = self.linear(embeddings)
-		#TODO: maybe later for logistics regression, output should be one hot
-		#output_probs = torch.nn.functional.softmax(fc,dim=1)
 		return fc
-
-	def eval(self):
-		with torch.no_grad():
-			dev_pred = self.linear(self.dev_x)
-			loss = self.loss_func(dev_pred,self.dev_y)
-		return dev_pred,loss.item()
 
 	
 	def optimize(self,lr=0.01,num_epochs=100):
@@ -53,15 +43,12 @@ class LinearClassifier(torch.nn.Module):
 			optimizer.zero_grad()
 			loss.backward(retain_graph=True)
 			optimizer.step()
-			#print(f'epoch: {epoch+1}, loss = {loss.item():.4f}')
+			print(f'epoch: {epoch+1}, loss = {loss.item():.4f}')
 			#implement stopping criterion
-			dev_pred,dev_loss=self.eval()
-			#print(f'epoch: {epoch+1}, dev loss = {dev_loss:.4f}')
 			if loss.item()<best_loss:
 				best_loss=loss.item()
 				best_model= self.linear
 				best_predictions = preds
-				#best_dev = dev_pred
 				stop_count=0
 			else:
 				if stop_count ==3:
@@ -69,13 +56,11 @@ class LinearClassifier(torch.nn.Module):
 				else:
 					stop_count+=1
 		final_pred = torch.argmax(best_predictions,dim=1).numpy()
-		dev_pred = best_model(self.dev_x)
[…]

### 2022-06-06 src/eval_classifier.py

@@ -1,0 +1,83 @@
+from LinearClassifier import LinearClassifier
+import torch
+from torch.utils.data import TensorDataset, DataLoader, RandomSampler
+from sklearn.metrics import accuracy_score
+
+class EvalClassifier(LinearClassifier):
+    def __init__(self,input_embeddings,output,tag_size,dev_x,dev_y):
+        '''
+        A Eval Classifier for eval task, inherited from Linear Classifer
+        But With batching and different logistics regression
+        '''
+        super().__init__(input_embeddings,output,tag_size)
+        self.dev_x = dev_x
+        self.dev_y = dev_y
+        # TODO: tranfer output to one hot vector
+        #self.output_vecs = 
+        #self.loss_func = torch.nn.NLLLoss()
+    
+    def forward(self,embeddings):
+        fc = self.linear(embeddings)
+        #output_probs = torch.nn.functional.softmax(fc,dim=1)
+        return fc
+    
+    def batching_data(self,dset='train',batch_size=32):
+        if dset == 'train':
+            data = TensorDataset(self.embeddings,self.output)
+        elif dset == 'dev':
+            data = TensorDataset(self.dev_x,self.dev_y)
+        sampler = RandomSampler(data)
+        dataloader = DataLoader(data,sampler=sampler,batch_size=batch_size)
+        return dataloader
+    
+    def eval(self):
+        with torch.no_grad():
+            dev_pred = self.forward(self.dev_x)
+            loss = self.loss_func(dev_pred,self.dev_y)
+        return dev_pred, loss.item()
+    
+    def optimize(self,batch_size=32, lr=0.01,num_epochs=100):
+        optimizer = torch.optim.LBFGS(self.linear.parameters(), lr = lr)
+        best_predictions_tr = None
+        best_predictions_dv = None
+        best_loss=float('inf')
+        stop_count = 0
+        for epoch in range(num_epochs):
+            train_batch = self.batching_data(batch_size=batch_size)
+            train_loss = 0
+            train_preds = []
+            dev_preds = []
+            for emb, label in train_batch:
+                pred = self.forward(emb)
+                train_preds.append(pred)
+                loss = self.loss_func(pred,label)
+                self.linear.zero_grad()
+                loss.backward(retain_graph=True)
+                # clip the gradient
+                torch.nn.utils.clip_grad_norm_(self.linear.parameters(), 1.0)
+                optimizer.step()
+                train_loss+=loss.item()
[…]

### 2022-06-06 src/LinearClassifier.py
fixed the bug in null space projection

@@ -43,7 +43,7 @@ class LinearClassifier(torch.nn.Module):
 			optimizer.zero_grad()
 			loss.backward(retain_graph=True)
 			optimizer.step()
-			print(f'epoch: {epoch+1}, loss = {loss.item():.4f}')
+			#print(f'epoch: {epoch+1}, loss = {loss.item():.4f}')
 			#implement stopping criterion
 			if loss.item()<best_loss:
 				best_loss=loss.item()
@@ -60,7 +60,7 @@ class LinearClassifier(torch.nn.Module):
 		#final_out = output.numpy()
 		#dev_out = self.dev_y.numpy()
 		train_acc = accuracy_score(self.output.numpy(),final_pred)
-		print(f'train accuracy score:{train_acc:.4f}')
+		#print(f'train accuracy score:{train_acc:.4f}')
 		#print(f'dev accuracy score:{accuracy_score(self.dev_y.numpy(),final_dev):.4f}')
 			
 		return best_model,train_acc
@@ -79,7 +79,7 @@ class INLPTraining(LinearClassifier):
 		if np.allclose(W, 0):
 			w_basis = np.zeros_like(W.T)
 		else:
-			w_basis = scipy.linalg.orth(W.T) # orthogonal basis
+			w_basis = scipy.linalg.orth(W.T) # orthogonal basis 
 		w_basis = w_basis * np.sign(w_basis[0][0]) # handle sign ambiguity
 		P_W = w_basis.dot(w_basis.T) # orthogonal projection on W's rowspace
 		return P_W
@@ -93,6 +93,7 @@ class INLPTraining(LinearClassifier):
 		:param rowspace_projection_matrices: List[np.array], a list of rowspace projections
 		:param input_dim: input dim
 		"""
+		# This is werid because rowspace is not null space
 
 		I = np.eye(self.input_dim)
 		Q = np.sum(rowspace_projection_matrices, axis=0)
@@ -113,6 +114,7 @@ class INLPTraining(LinearClassifier):
 
 	def run_INLP_loop(self,iteration,min_acc=0.0):
 		I = np.eye(self.input_dim)
+		P = I
 		Ws = []
 		rowspace_projections = []
 		for i in range(iteration):
@@ -124,10 +126,15 @@ class INLPTraining(LinearClassifier):
 				continue
 			W = bm.weight.detach().numpy()
 			Ws.append(W)
+			# Noted this is the projection space for W, not the null space
[…]

### 2022-06-06 src/probe.py

@@ -14,6 +14,7 @@ import numpy as np
 from tqdm import tqdm
 from sklearn import preprocessing
 from LinearClassifier import LinearClassifier, INLPTraining
+from eval_classifier import EvalClassifier
 #logging.basicConfig(level-logging.INFO) #turn on detailed logging
 
 
@@ -80,22 +81,33 @@ dev_tokens,dev_seq,dev_mask = bert_tokenization(dev_input,tokenizer)
 train_output = torch.load('../data/pmb_gold/gold_train_embeddings.pt')
 dev_output = torch.load('../data/pmb_gold/gold_dev_embeddings.pt')
 
-train_embeddings,train_y = from_sents_to_words(train_df,'syn',train_tokens,train_output)
-dev_embeddings,dev_y = from_sents_to_words(dev_df,'syn',dev_tokens,dev_output)
+train_embeddings,train_y = from_sents_to_words(train_df,'sem',train_tokens,train_output)
+dev_embeddings,dev_y = from_sents_to_words(dev_df,'sem',dev_tokens,dev_output)
 
-
+# l=train_y.tolist()
+# import collections
+# print(collections.Counter(l))
+# print(len(l))
+# sys.exit()
 # print(train_embeddings.shape)
 # print(train_y.shape)
 # print(dev_embeddings.shape)
 # print(dev_y.shape)
 
-print(dev_embeddings.dtype)
+#print(dev_embeddings.dtype)
 num_epochs = 100
 num_tags = len(np.unique(train_y.numpy()))
 input_dim = train_embeddings.shape[1] #should be 768
 
-inlp_syn = INLPTraining(train_embeddings,train_y,num_tags,dev_x=dev_embeddings,dev_y=dev_y)
-inlp_syn.run_INLP_loop(10)
+# sem_test = LinearClassifier(train_embeddings,train_y,num_tags)
+# sem_test.optimize()
+# eval_test = EvalClassifier(train_embeddings,train_y,num_tags,dev_embeddings,dev_y)
+# eval_test.optimize(batch_size=1000)
+inlp_syn = INLPTraining(train_embeddings,train_y,num_tags)
+P,P_is,Ws=inlp_syn.run_INLP_loop(10)
+print(f'the rank of P is :{np.linalg.matrix_rank(P)}')
+for P_i in P_is:
+	print(np.linalg.matrix_rank(P_i))
 sys.exit()
[…]

### 2022-06-06 src/LinearClassifier.py
fixing inlp

@@ -93,8 +93,7 @@ class INLPTraining(LinearClassifier):
 		:param rowspace_projection_matrices: List[np.array], a list of rowspace projections
 		:param input_dim: input dim
 		"""
-		# This is werid because rowspace is not null space
-
+		# This is werid because Q is not normalized so the N(P) = I-P does not work
 		I = np.eye(self.input_dim)
 		Q = np.sum(rowspace_projection_matrices, axis=0)
 		P = I - self.get_rowspace_projection(Q)

### 2022-06-07 src/eval_classifier.py
changing optimizer for eval

@@ -32,12 +32,13 @@ class EvalClassifier(LinearClassifier):
     
     def eval(self):
         with torch.no_grad():
+            # TODO: batching if necessary
             dev_pred = self.forward(self.dev_x)
             loss = self.loss_func(dev_pred,self.dev_y)
         return dev_pred, loss.item()
     
     def optimize(self,batch_size=32, lr=0.01,num_epochs=100):
-        optimizer = torch.optim.LBFGS(self.linear.parameters(), lr = lr)
+        optimizer = torch.optim.Adagrad(self.linear.parameters(), lr = lr)
         best_predictions_tr = None
         best_predictions_dv = None
         best_loss=float('inf')

### 2022-06-07 src/LinearClassifier.py
adding processing class

@@ -6,6 +6,9 @@ import scipy
 from typing import List
 import tqdm
 
+## defining GPU here
+device = torch.device("cuda") if torch.cuda.is_available() else torch.device('cpu')
+
 class LinearClassifier(torch.nn.Module):
 	def __init__(self, input_embeddings,output,tag_size):
 		'''
@@ -39,7 +42,9 @@ class LinearClassifier(torch.nn.Module):
 		stop_count = 0
 		for epoch in range(num_epochs):
 			preds = self.forward(self.embeddings)
+			print(preds.shape)
 			loss = self.loss_func(preds,self.output)
+			print(self.output.shape)
 			optimizer.zero_grad()
 			loss.backward(retain_graph=True)
 			optimizer.step()
@@ -55,11 +60,11 @@ class LinearClassifier(torch.nn.Module):
 					break
 				else:
 					stop_count+=1
-		final_pred = torch.argmax(best_predictions,dim=1).numpy()
+		final_pred = torch.argmax(best_predictions,dim=1).cpu().numpy()
 		#final_dev = torch.argmax(best_dev,dim=1).numpy()
 		#final_out = output.numpy()
 		#dev_out = self.dev_y.numpy()
-		train_acc = accuracy_score(self.output.numpy(),final_pred)
+		train_acc = accuracy_score(self.output.cpu().numpy(),final_pred)
 		#print(f'train accuracy score:{train_acc:.4f}')
 		#print(f'dev accuracy score:{accuracy_score(self.dev_y.numpy(),final_dev):.4f}')
 			
@@ -102,13 +107,13 @@ class INLPTraining(LinearClassifier):
 	def reinitialize_classifier(self):
 		in_size = self.linear.in_features
 		out_size = self.linear.out_features
-		self.linear = torch.nn.Linear(in_size,out_size)
+		self.linear = torch.nn.Linear(in_size,out_size,device=device)
 	
 	def apply_projection(self,P):
 		'''
 		applying projection of P to the embedding vectors
 		'''
-		P = torch.tensor(P,dtype=torch.float)
+		P = torch.tensor(P,dtype=torch.float,device=device)
 		self.embeddings =  torch.matmul(P,self.original_embedding).T
 
 	def run_INLP_loop(self,iteration,min_acc=0.0):
@@ -123,7 +128,7 @@ class INLPTraining(LinearClassifier):
 			if acc < min_acc:
 				# TODO: not sure it should be continue here
 				continue
[…]

### 2022-06-07 src/load_bert.py

@@ -8,19 +8,126 @@ import itertools
 import pdb
 import json
 import pickle
+import numpy as np
+from collections import defaultdict
 from typing import Dict, List
 from utils import bert_tokenization
+from sklearn import preprocessing
+from torch.utils.data import TensorDataset, DataLoader, RandomSampler, SequentialSampler
 #from nltk.stem import WordNetLemmatizer
 #logging.basicConfig(level-logging.INFO) #turn on detailed logging
 
 ## defining GPU here
-
 device = torch.device("cuda") if torch.cuda.is_available() else torch.device('cpu')
 
-## defining tokenizer and bert model here
-tokenizer = AutoTokenizer.from_pretrained("bert-base-uncased", use_fast=False)
-model = AutoModel.from_pretrained('bert-base-uncased')
-model = model.to(device)
+class DataProcessing:
+    '''
+    A class for processing data and get right format of input output for model
+    '''
+    def __init__(self, input_pk_path):
+        self.input_df = pd.read_pickle(input_pk_path)
+    
+    def bert_tokenize(self,max_len=32):
+        '''
+        Returns Bert tokenized objects
+        '''
+        tokenizer = AutoTokenizer.from_pretrained('bert-base-uncased', use_fast=True)
+        input_text = self.input_df['text'].tolist()
+        tokens = tokenizer(input_text,
+                    max_length=max_len,
+                    truncation=True,
+                    is_split_into_words=True,
+                    padding=True,
+                    return_tensors='pt')
+        tokens = tokens.to(device)
+        self.tokens = tokens
+        return self.tokens
+    
+    
+    def get_bert_embeddings(self,load_option='no_save',load_path=None):
+
+        if load_option == 'load':
+            output_tensor = torch.load(load_path)
+            output_tensor = output_tensor.to(device)
+        
+        else:
+            input_seqs = self.tokens['input_ids']
+            input_mask = self.tokens['attention_mask']
+            model = AutoModel.from_pretrained('bert-base-uncased')
+            model = model.to(device)
+
+            ## batching data
+            data = TensorDataset(input_seqs,input_mask)
+            loader = DataLoader(data,batch_size=1000)
+
+            outputs = []
+            for param in model.parameters():
+                param.requires_grad=False
+            model.eval()
+
[…]

### 2022-06-07 src/probe.py

@@ -8,57 +8,37 @@ import pdb
 import json
 import pickle
 from typing import Dict, List
-from collections import defaultdict
 from utils import bert_tokenization, get_projection_to_intersection_of_nullspaces, get_rowspace_projection, train_linear_classifier
 import numpy as np
 from tqdm import tqdm
-from sklearn import preprocessing
+
+from load_bert import DataProcessing
 from LinearClassifier import LinearClassifier, INLPTraining
 from eval_classifier import EvalClassifier
 #logging.basicConfig(level-logging.INFO) #turn on detailed logging
 
+## defining GPU here
+device = torch.device("cuda") if torch.cuda.is_available() else torch.device('cpu')
 
-def from_sents_to_words(input_df,task_keyword,tokens,output):
-	'''
-	this function checks the tokenization process and transfer data into word level
-	Arguments:
-		input_df: a pd.DataFrame object
-		task_keyword: {syn,sem}
-		tokens: bert tokenized object
-		output: a tensor with shape (sentence num, max_seq_len, embedding dim)
-	Returns: 
-		embeddings: a tensor with shape (word num, embedding dim)
-		y: a tensor with shape (word num,1)
-	'''
-	input_text = input_df['text'].tolist()
-	if task_keyword == 'syn':
-		input_y = input_df['ccg_tags'].tolist()
-	else:
-		input_y = input_df['semantics_tags'].tolist()
-	label_encoder = preprocessing.LabelEncoder()
-	label_encoder.fit(np.concatenate(input_y))
-	encoded_y = [label_encoder.transform(l) for l in input_y]
-	## mapping tokens back to word ids
-	word_inds = [tokens.word_ids(i) for i in range(len(input_y))]
-	d_list = []
-	embed_by_sent = []
-	skip_ind = []
-	for i in range(len(word_inds)):
-		d = defaultdict(list)
-		for ind,emb in zip(word_inds[i], output[i]):
-			if isinstance(ind, int):
-				d[ind].append(emb)
-	
-		word_in_sent = [torch.mean(torch.stack(d[k],0),0) if len(d[k])>1 else d[k][0] for k in d.keys()]
-		if len(word_in_sent)!= len(input_text[i]):
-			skip_ind.append(i)
[…]

### 2022-06-07 src/LinearClassifier.py
disable printing

@@ -42,9 +42,10 @@ class LinearClassifier(torch.nn.Module):
 		stop_count = 0
 		for epoch in range(num_epochs):
 			preds = self.forward(self.embeddings)
-			print(preds.shape)
+			# print(preds.shape)
+			# print(self.output.shape)
 			loss = self.loss_func(preds,self.output)
-			print(self.output.shape)
+			
 			optimizer.zero_grad()
 			loss.backward(retain_graph=True)
 			optimizer.step()

### 2022-06-08 src/LinearClassifier.py
fixing label issue

@@ -22,7 +22,7 @@ class LinearClassifier(torch.nn.Module):
 		super().__init__()
 		self.embeddings = input_embeddings
 		self.output = output
-		self.linear = torch.nn.Linear(input_embeddings.shape[1], tag_size)
+		self.linear = torch.nn.Linear(input_embeddings.shape[1], tag_size,device=device)
 		# class weight performs really worse
 		# cls_weight = compute_class_weight('balanced',classes=np.array(range(tag_size)),y=output.numpy())
 		# cls_weight = torch.tensor(cls_weight,dtype=torch.float)

### 2022-06-08 src/eval_classifier.py

@@ -37,7 +37,7 @@ class EvalClassifier(LinearClassifier):
             loss = self.loss_func(dev_pred,self.dev_y)
         return dev_pred, loss.item()
     
-    def optimize(self,batch_size=32, lr=0.01,num_epochs=100):
+    def optimize(self,batch_size=32, lr=0.01,num_epochs=12):
         optimizer = torch.optim.Adagrad(self.linear.parameters(), lr = lr)
         best_predictions_tr = None
         best_predictions_dv = None
@@ -66,6 +66,7 @@ class EvalClassifier(LinearClassifier):
             if dev_loss<best_loss:
                 # train_preds = torch.cat(train_preds)
                 # dev_preds = torch.cat(dev_preds)
+                best_loss = dev_loss
                 best_predictions_tr = torch.cat(train_preds)
                 best_predictions_dv = dev_preds
                 best_model=self.linear
@@ -76,9 +77,9 @@ class EvalClassifier(LinearClassifier):
                 else:
                     stop_count+=1
             
-        final_train = torch.argmax(best_predictions_tr,dim=1).numpy()
-        final_dev = torch.argmax(best_predictions_dv,dim=1).numpy()
-        print(f'train accuracy score:{accuracy_score(self.output.numpy(),final_train):.4f}')
-        print(f'dev accuracy score:{accuracy_score(self.dev_y.numpy(),final_dev):.4f}')
+        final_train = torch.argmax(best_predictions_tr,dim=1).cpu().numpy()
+        final_dev = torch.argmax(best_predictions_dv,dim=1).cpu().numpy()
+        print(f'train accuracy score:{accuracy_score(self.output.cpu().numpy(),final_train):.4f}')
+        print(f'dev accuracy score:{accuracy_score(self.dev_y.cpu().numpy(),final_dev):.4f}')
 
         

### 2022-06-08 src/load_bert.py

@@ -24,8 +24,15 @@ class DataProcessing:
     '''
     A class for processing data and get right format of input output for model
     '''
-    def __init__(self, input_pk_path):
-        self.input_df = pd.read_pickle(input_pk_path)
+    def __init__(self, standard,dset):
+        '''
+        standard = {gold,silver}
+        dset = {train,test,dev}
+        '''
+        self.standard = standard
+        self.dset = dset
+        path = '../data/pmb_'+standard+'/'+standard+'_'+dset+'.pkl'
+        self.input_df = pd.read_pickle(path)
     
     def bert_tokenize(self,max_len=32):
         '''
@@ -44,7 +51,9 @@ class DataProcessing:
         return self.tokens
     
     
-    def get_bert_embeddings(self,load_option='no_save',load_path=None):
+    def get_bert_embeddings(self,load_option='save'):
+
+        load_path = '../data/pmb_'+self.standard+'/'+self.standard+'_'+self.dset+'_emb.pt'
 
         if load_option == 'load':
             output_tensor = torch.load(load_path)
@@ -93,9 +102,10 @@ class DataProcessing:
             input_y = self.input_df['ccg_tags'].tolist()
         else:
             input_y = self.input_df['semantics_tags'].tolist()
-        label_encoder = preprocessing.LabelEncoder()
-        label_encoder.fit(np.concatenate(input_y))
-        encoded_y = [label_encoder.transform(l) for l in input_y]
+        
+        with open('../data/pmb_'+self.standard+'/'+task_keyword+'_mapping.json') as f:
+                label_encoder = json.load(f)
+
         ## mapping tokens back to word ids
         word_inds = [self.tokens.word_ids(i) for i in range(len(input_y))]
         d_list = []
@@ -114,11 +124,10 @@ class DataProcessing:
             embed_by_sent.append(torch.stack(word_in_sent,0))
             d_list.append(d)
         embeddings = torch.cat(embed_by_sent)
-        y = torch.tensor([item for i, sublist in enumerate(encoded_y) if i not in skip_ind for item in sublist ], dtype=torch.long)
-        
+        y = torch.tensor([label_encoder[item] for i, sublist in enumerate(input_y) if i not in skip_ind for item in sublist ], dtype=torch.long)
         #embeddings = embeddings.to(device)
         y = y.to(device)
-        return embeddings,y
[…]

### 2022-06-08 src/probe.py

@@ -20,17 +20,29 @@ from eval_classifier import EvalClassifier
 ## defining GPU here
 device = torch.device("cuda") if torch.cuda.is_available() else torch.device('cpu')
 
-silver_train = DataProcessing('../data/pmb_gold/gold_train.pkl')
-silver_train.bert_tokenize()
-out_sil = silver_train.get_bert_embeddings('load','../data/pmb_gold/gold_train_embeddings.pt')
-print(out_sil.shape)
+gold_train = DataProcessing('gold','train')
+gold_train.bert_tokenize()
+train_emb = gold_train.get_bert_embeddings('load')
 
-emb, y = silver_train.from_sents_to_words('syn',out_sil)
-print(emb.is_cuda)
-print(y.is_cuda)
-num_tags = len(np.unique(y.cpu().numpy()))
+emb_tr, y_tr, num_tags = gold_train.from_sents_to_words('sem',train_emb)
 
+gold_dev = DataProcessing('gold','dev')
+gold_dev.bert_tokenize()
+dev_emb = gold_dev.get_bert_embeddings('load')
+
+emb_dev, y_dev, num_tags = gold_dev.from_sents_to_words('sem',dev_emb)
+
+# print(emb.is_cuda)
+# print(y.is_cuda)
 print(num_tags)
+## TODO: weird inconsistency
+eval_gold = EvalClassifier(emb_tr,y_tr,num_tags,emb_dev,y_dev)
+eval_gold.optimize()
+
+sys.exit()
+
+
+## calling INLP
 inlp_syn = INLPTraining(emb,y,num_tags)
 inlp_syn = inlp_syn.to(device)
 P,P_is,Ws=inlp_syn.run_INLP_loop(10)
@@ -39,6 +51,10 @@ for P_i in P_is:
 	print(np.linalg.matrix_rank(P_i))
 
 sys.exit()
+
+## calling eval
+#eval_t = EvalClassifier(emb,y,num_tags,dev_x=,dev_y=)
+#eval_t.optimize()
 
 	
 

### 2022-06-08 src/load_bert.py
making everything lands in cpu

@@ -46,7 +46,6 @@ class DataProcessing:
                     is_split_into_words=True,
                     padding=True,
                     return_tensors='pt')
-        tokens = tokens.to(device)
         self.tokens = tokens
         return self.tokens
     
@@ -57,7 +56,7 @@ class DataProcessing:
 
         if load_option == 'load':
             output_tensor = torch.load(load_path)
-            output_tensor = output_tensor.to(device)
+            #output_tensor = output_tensor.to(device)
         
         else:
             input_seqs = self.tokens['input_ids']
@@ -81,9 +80,9 @@ class DataProcessing:
                     output = model(word_ids,attention_mask=mask)
                     outputs.append(output[0])
             
-            output_tensor = torch.cat(outputs).detach()
+            output_tensor = torch.cat(outputs).detach().cpu()
             if load_option == 'save':
-                torch.save(output_tensor.cpu(),load_path)
+                torch.save(output_tensor,load_path)
         
         return output_tensor
 
@@ -126,7 +125,7 @@ class DataProcessing:
         embeddings = torch.cat(embed_by_sent)
         y = torch.tensor([label_encoder[item] for i, sublist in enumerate(input_y) if i not in skip_ind for item in sublist ], dtype=torch.long)
         #embeddings = embeddings.to(device)
-        y = y.to(device)
+        #y = y.to(device)
         return embeddings,y,len(label_encoder)
 
 

### 2022-06-08 src/probe.py
adding comment

@@ -20,22 +20,23 @@ from eval_classifier import EvalClassifier
 ## defining GPU here
 device = torch.device("cuda") if torch.cuda.is_available() else torch.device('cpu')
 
-gold_train = DataProcessing('gold','train')
+gold_train = DataProcessing('silver','train')
 gold_train.bert_tokenize()
 train_emb = gold_train.get_bert_embeddings('load')
 
+## all in cpu right now
 emb_tr, y_tr, num_tags = gold_train.from_sents_to_words('sem',train_emb)
 
-gold_dev = DataProcessing('gold','dev')
+gold_dev = DataProcessing('silver','dev')
 gold_dev.bert_tokenize()
 dev_emb = gold_dev.get_bert_embeddings('load')
 
+## all in cpu right now
 emb_dev, y_dev, num_tags = gold_dev.from_sents_to_words('sem',dev_emb)
 
-# print(emb.is_cuda)
-# print(y.is_cuda)
+
 print(num_tags)
-## TODO: weird inconsistency
+
 eval_gold = EvalClassifier(emb_tr,y_tr,num_tags,emb_dev,y_dev)
 eval_gold.optimize()
 

### 2022-06-08 src/LinearClassifier.py
fixing out of memory issue

@@ -20,6 +20,7 @@ class LinearClassifier(torch.nn.Module):
 		dev_y: dev label for stopping criterion
 		'''
 		super().__init__()
+		## everything defined in GPU
 		self.embeddings = input_embeddings
 		self.output = output
 		self.linear = torch.nn.Linear(input_embeddings.shape[1], tag_size,device=device)
@@ -31,7 +32,8 @@ class LinearClassifier(torch.nn.Module):
 	def forward(self,embeddings):
 		# embedding size = [batch_size, embed_dim]
 		# output size = [batch_size,tag_size]
-		fc = self.linear(embeddings)
+		emb = embeddings.to(device)
+		fc = self.linear(emb)
 		return fc
 
 	
@@ -40,11 +42,12 @@ class LinearClassifier(torch.nn.Module):
 		best_predictions = None
 		best_loss = float('inf')
 		stop_count = 0
+		output = self.output.to(device)
 		for epoch in range(num_epochs):
 			preds = self.forward(self.embeddings)
 			# print(preds.shape)
 			# print(self.output.shape)
-			loss = self.loss_func(preds,self.output)
+			loss = self.loss_func(preds,output)
 			
 			optimizer.zero_grad()
 			loss.backward(retain_graph=True)
@@ -65,7 +68,7 @@ class LinearClassifier(torch.nn.Module):
 		#final_dev = torch.argmax(best_dev,dim=1).numpy()
 		#final_out = output.numpy()
 		#dev_out = self.dev_y.numpy()
-		train_acc = accuracy_score(self.output.cpu().numpy(),final_pred)
+		train_acc = accuracy_score(self.output.numpy(),final_pred)
 		#print(f'train accuracy score:{train_acc:.4f}')
 		#print(f'dev accuracy score:{accuracy_score(self.dev_y.numpy(),final_dev):.4f}')
 			
@@ -106,6 +109,7 @@ class INLPTraining(LinearClassifier):
 		return P
 
 	def reinitialize_classifier(self):
+		## may be empty cache here
 		in_size = self.linear.in_features
 		out_size = self.linear.out_features
 		self.linear = torch.nn.Linear(in_size,out_size,device=device)
@@ -114,7 +118,8 @@ class INLPTraining(LinearClassifier):
 		'''
 		applying projection of P to the embedding vectors
 		'''
-		P = torch.tensor(P,dtype=torch.float,device=device)
+		## may be empty cache here
[…]

### 2022-06-08 src/eval_classifier.py

@@ -3,6 +3,9 @@ import torch
 from torch.utils.data import TensorDataset, DataLoader, RandomSampler
 from sklearn.metrics import accuracy_score
 
+device = torch.device("cuda") if torch.cuda.is_available() else torch.device('cpu')
+
+
 class EvalClassifier(LinearClassifier):
     def __init__(self,input_embeddings,output,tag_size,dev_x,dev_y):
         '''
@@ -10,8 +13,8 @@ class EvalClassifier(LinearClassifier):
         But With batching and different logistics regression
         '''
         super().__init__(input_embeddings,output,tag_size)
-        self.dev_x = dev_x
-        self.dev_y = dev_y
+        self.dev_x = dev_x.to(device)
+        self.dev_y = dev_y.to(device)
         # TODO: tranfer output to one hot vector
         #self.output_vecs = 
         #self.loss_func = torch.nn.NLLLoss()
@@ -37,7 +40,7 @@ class EvalClassifier(LinearClassifier):
             loss = self.loss_func(dev_pred,self.dev_y)
         return dev_pred, loss.item()
     
-    def optimize(self,batch_size=32, lr=0.01,num_epochs=12):
+    def optimize(self,batch_size=32, lr=0.001,num_epochs=12):
         optimizer = torch.optim.Adagrad(self.linear.parameters(), lr = lr)
         best_predictions_tr = None
         best_predictions_dv = None
@@ -49,13 +52,15 @@ class EvalClassifier(LinearClassifier):
             train_preds = []
             dev_preds = []
             for emb, label in train_batch:
+                emb = emb.to(device)
+                label = label.to(device)
                 pred = self.forward(emb)
                 train_preds.append(pred)
                 loss = self.loss_func(pred,label)
                 self.linear.zero_grad()
                 loss.backward(retain_graph=True)
                 # clip the gradient
-                torch.nn.utils.clip_grad_norm_(self.linear.parameters(), 1.0)
+                #torch.nn.utils.clip_grad_norm_(self.linear.parameters(), 1.0)
                 optimizer.step()
                 train_loss+=loss.item()
             avg_loss = train_loss/len(train_batch)

### 2022-06-08 src/probe.py

@@ -34,18 +34,17 @@ dev_emb = gold_dev.get_bert_embeddings('load')
 ## all in cpu right now
 emb_dev, y_dev, num_tags = gold_dev.from_sents_to_words('sem',dev_emb)
 
-
+if torch.cuda.is_available():
+	torch.cuda.empty_cache()
 print(num_tags)
 
-eval_gold = EvalClassifier(emb_tr,y_tr,num_tags,emb_dev,y_dev)
-eval_gold.optimize()
-
+t = EvalClassifier(emb_tr,y_tr,num_tags,emb_dev,y_dev)
+t.optimize()
 sys.exit()
 
-
 ## calling INLP
-inlp_syn = INLPTraining(emb,y,num_tags)
-inlp_syn = inlp_syn.to(device)
+inlp_syn = INLPTraining(emb_tr,y_tr,num_tags)
+# inlp_syn = inlp_syn.to(device)
 P,P_is,Ws=inlp_syn.run_INLP_loop(10)
 print(f'the rank of P is :{np.linalg.matrix_rank(P)}')
 for P_i in P_is:

### 2022-06-08 paper/paper.tex

%
% File acl2020.tex
%
%% Based on the style files for ACL 2020, which were
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2020}
\usepackage{times}
\usepackage{latexsym}
\usepackage{multirow}
\renewcommand{\UrlFont}{\ttfamily\small}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B\textsc{ib}\TeX}

\title{Proposal for Analyzing Language Models: Separatablity of Syntax and Semantics}

\author{Qingxia Guo, Saiya Karamali, Lindsay Skinner, \and Gladys Wang
 \\ University of Washington \\ 
\texttt{\{qg07, karamali, skinnel, qinyanw\}@uw.edu}\\ 
}
\date{}

\begin{document}

\maketitle

\begin{abstract}
%Motivating question and or topics of interest.
Syntax and semantics are at the foundation every language, yet distinctions between the two are not readily agreed upon. We seek to explore how representations of these two information sets manifest in BERT embeddings. Specifically we investigate the degree of the linear separability of syntactic and semantic information in BERT embeddings, as well as quantify how important the linear component corresponding to one information set is to solving a classification task on the other information set. We use Iterative Nullspace Projection to decompose word-level BERT embeddings into syntactic, non-syntactic, semantic and non-semantic components to be used in syntactic and semantic classification tasks. Our results show that...
%How we address it and our results.
%Simplified conclusion. 
\end{abstract}

\section{Introduction}
\label{sec:introduction}

%Motivate your question
%Situate your question
%Summarize main findings and contributions

The boundary between semantics and syntax is a hotly debated topic in linguistics, but do large language models make such a distinction? If they do make such a distinction, do language model embeddings present this information in a way that is easily separated and recognized by humans? The objective of this project is to explore BERT\rq s \citep{bert} reliance on certain syntactic information when handling a semantic task, and vice versa. Specifically, we seek to quantify the importance of linearly-separable syntactic or semantic information when performing semantic or syntactic classification, respectively.

%may remove this
To achieve our goal, we construct a linear probing system for a task and then employ Iterative Nullspace Projection (INLP from here on) \citep{inlp} to generate a new embedding devoid of information learned from the probing task. We then measure the performance of this new embedding on downstream syntactic and semantic classification tasks. The design of our probing procedure follows \citep{amnesia}, which employs INLP to investigate whether BERT uses part-of-speech (POS) information when solving language modeling (LM) tasks. INLP has been used for a variety of tasks (\citep{inlp}, \citep{amnesia}, \citep{mbertGreek}, etc.) though this is the first case we know of in which INLP has been used to investigate the linear separability of syntactic and semantic information. 
%why did amnesia have citealp?

A novel method for removing information from an embedding, INLP iteratively trains linear models on a specific classification task, and projects the input on intersection of the nullspaces of those linear models. Our objective is that, by applying the INLP procedure to a syntactic task, we are able to separate the representation into a syntactic space and a non-syntactic space. We then compare the performance of a model that seeks to classify semantic labels using the original BERT embeddings with an otherwise identical model trained on embeddings projected onto the non-syntactic space, in order to see if BERT is using syntactic information when performing the semantic task. Conversely, we can also first probe a semantic task, thus defining a semantic and non-semantic space, and then investigate the performance of embeddings projected onto those spaces when performing a syntactic classification task. Once we derive the semantic space and syntactic space from the experiment, we further investigate on the separability of the two spaces by comparing the embedding projections. %See section ???
%Make the research question very clear
%Should a model be able to solve a semantic task without syntactic information?
%Vice versa? (Just be very explicit here)

To evaluate the separability of syntax and semantics, we use two task: one task for the probing system and INLP procedure, and one task for evaluating performance on embeddings before and after INLP. We choose Combinatory Categorical Grammar (CCG from here on) tagging \citep{ccg-bank} as the syntactic task and semantic category labeling \citep{propbank} as the semantic task.

The remainder of the paper proceeds as follows: Section \ref{sec:related} explores previous work related to our experiment. Section \ref{sec:method} provides a description of the probing and evaluation tasks and gives an overview of the experiment pipeline. Section \ref{sec:result} reviews our experiments and affiliated results. Section \ref{sec:discussion} discusses the implications of those results. Finally, section \ref{sec:conclusion} gives an overview of the entire process and outlines possible next steps.

\section{Related Work}
\label{sec:related}

%Quick
%Summarize papers similar to yours
%highlight how they motivate your approach
%how are you different

%Related work discussion goes here.

The separation and overlap between syntax and semantics has been of interest to linguists for many years. More recently, with the growing popularity of large language models, computational linguists have begun to explore how large language models deal with the boundaries of these information sets in word and sentence embeddings.

In \citep{disentangle} the authors use paraphrase pairs and new target syntax to train a semantic encoder, syntactic encoder and decoder to learn separate representations of the semantic and syntactic information contained in BART embeddings, in order to create semantically equivalent paraphrases with the new syntactic structure. Alongside the encoders they also train an adversarial syntax discriminator to try and predict the source syntax from the semantic embeddings, thus encouraging the disentanglement of the semantic and syntactic information by training the semantic embedder to remove as much syntactic information as possible. Their results show that one can achieve some removal of syntactic information from semantic embeddings, so disentanglement of some information is possible. Though they do not achieve perfect separation of the two information sets. Other non-linear approaches to syntactic-semantic information disentanglement have been carried out in \citep{multiDis}

%Similarly, the authors of \citep{multiDis} train a variational autoencoder with two latent variables for syntactic representations and semantic representations in order to test the separability of these two information sets in sentence embeddings across various models. To encourage the separation of these information sets the authors use multi-task training that incorporates paraphrase reconstruction loss, discriminative paraphrase loss and word position loss. They then use syntactic similarity tasks to measure their success in disentangling the syntactic and semantic information sets across various model embeddings.

%How we differ from these
Unlike the aforementioned studies, we seek to explore the linear separability of syntactic and semantic information in large language model embeddings at the word level. To accomplish this task we apply the Iterative Nullspace Projection method to syntactic (CCG) and semantic labeling tasks in order to define the syntactic and semantic components of BERT embeddings that will be used in our downstream classification tasks.

INLP, introduced in \citep{inlp}, is a method to define a linear guarding function that masks all the linear information in a word embedding that may be used for a downstream classification task. In the original paper the authors use this method to remove gender bias from BERT embeddings of biographical descriptions and then measure how easy it is to determine an individual's gender from the guarded embedding by using various downstream classification methods. Beyond using the INLP method to guard protected attributes, the authors hypothesize several additional use cases for this procedure, including information disentanglement.

The authors of \citep{amnesia} use INLP for exactly this task. They use INLP in order to separate and guard certain linguistic information sets from BERT embeddings in order to better understand what information is being used by large language models, and not just what is encoded. The main premise behind this paper is that if a particular property is used to solve a task, then the removal of that property should negatively influence the model's ability to solve that task. Conversely, if the removal of a property has little influence on the model's ability to perform a task then we know that property is not a significant contributing factor in the model's ability to perform that task. Specifically, \citep{amnesia} seeks to quantify the importance of the information sets used for part-of-speech tagging, syntactic dependency labeling, named entity recognition and syntactic constituency boundaries on BERT's ability to perform the language modeling task.

We take a similar approach to \citep{amnesia} by separating the information sets used for CCG tagging and semantic labeling from word-level BERT embeddings. However, as we are interested in the linear separability of these two information sets, we will test how the removal of these information sets impacts the embeddings' performance on semantic labeling and CCG tagging, respectively, rather than language modeling.

\section{Methods}
\label{sec:method}

%How are you addressing your research question
%models, data, etc.

We construct two separate probing tasks to isolate the syntactic and semantic information in word-level BERT embeddings. The embeddings are separated into syntactic and non-syntactic, and semantic and non-semantic components via INLP which is described in section \ref{sec:inlp-method}. These embedding components are then combined to form new embeddings, which are evaluated on the same tasks that were used for probing. %If time permits, we will also evaluate how these new embeddings perform on the language modeling task.

\subsection{The Iterative Null-Space Projection method}
\label{sec:inlp-method}

The INLP method first introduced in \citep{inlp}, is used to create a guarding function that masks all the linear information contained in a set of vectors, $X$, that can be used map each vector to $c \in C$, where $C$ is the set of all categories. This is accomplished by training a linear classifier, a matrix $W$, that is applied to each $x \in X$ in order to predict the correct category $c$ with the greatest possible accuracy. In other words, $Wx$ defines a distribution over the set of categories $C$ and we assign $x$ to the class $c \in C$ which is allotted the greatest probability by $Wx$. Note that the classifier's accuracy must be greater than that achieved by guessing the majority category, otherwise $x$ contains no linear information relevant for the categorization task and thus no guarding function is needed. Once $W$ is determined, for any $x \in X$ we can remove the information that $W$ uses to predict $c$ by projecting $x$ onto the null-space of $W$, $N(W) = \{x | Wx=0\}$. Call this projection function $P_1$ and let $\hat{x} = P_1(x)$. This removes all of the linear information in $x$ that $W$ used to predict the category $c$. 
%Need to cite INLP paper

However, this process does not necessarily remove all of the linear information in $x$ that could be used to predict $c$. For example, $x$ may contain redundant information and $W$ may have only used one set of this information for its prediction. In this case, the redundant information would still be present in $\hat{x}$. Thus, we must repeat the above process, defining a new linear classifier $\hat{W}$ that uses $\hat{x}$  to predict $c$. If $\hat{W}$ is still able to predict $c$ with a greater than majority class guess accuracy, then we know that $\hat{x}$ contained linear information about $c$. As above, we project $\hat{x}$ onto the null-space of $\hat{W}$ via the projection function $P_2$ and define a new $\hat{x} = P_2(P_1(x))$.

We iteratively apply this process until no linear information remains in $\hat{x}$, i.e. a linear classifier is unable to predict the correct category $c$ with any probability greater than that achieved by guessing the majority class. The final $\hat{x} = P_n(P_{n-1}(\dots P_1(x)))$ contains no linear information about the categories in $C$ and we call $P(x) = P_n(P_{n-1}(\dots P_1(x)))$ the guarding function.

We will pair the INLP method with the probing tasks described in sections \ref{sec:syntactic} and \ref{sec:semantics} in order to create two guarding functions that will enable us to isolate the linear components of BERT embeddings that contain syntax-specific and semantics-specific information.

\subsection{Data}
\label{sec:data}

%Description of the data set used and the information it contains relevant to the tasks of interest. Talk about pre-processing the syntax trees to get CCG tags. 
%We will use the English V4 subset of the CoNLL 2012 shared task data \citep{2012-conll} . We must perform two pre-processing steps for this data to be used for our probing and evaluation tasks. The first is to apply \citet{ccg-bank}'s CCG Derivation algorithm to the parse tree field in the dataset, in order to create the CCG tags for each word. If this proves to be too time-consuming or computationally expensive then we shall change the syntactic probing task to utilize the POS tags available in the dataset, in place of CCG tags. The second task is to use the SRL frames in the dataset to generate (verb, BIO-argument-tag) pairs that will act as the categories for the semantic probing task.

%Need to update this section to reflect the new dataset
We use the NAME data set to test the linear separability of the semantic and syntactic information in word-level BERT embeddings. %NEED TO FILL THIS OUT STILL

\subsection{Syntactic probing task}
\label{sec:syntactic}

The syntactic probing task involves training a linear classifier on the final layer BERT embeddings in order to predict the CCG tag associated with each word. We will use this classifier in the INLP algorithm in order to create a guarding function for the information that is necessary to complete the CCG labeling task. For a given embedding, $x$, the projection that results from applying this guarding function, $P_{syn}$, to the embedding will represent the non-syntactic information contained in the embedding and will from now on be referred to as the ``non-syntactic component'' of the embedding, $v_{no syn} = P_{syn} x$. We can then determine the ``syntactic component'' of the embedding by taking the difference of the embedding vector with the non-syntactic component, $v_{syn} = x - v_{no syn}$.

\subsection{Semantic probing task}
\label{sec:semantics}

Similar to the above, the semantic probing tasks involves training a linear classifier on the final layer BERT embeddings in order to predict the semantic tag (described in the data section) associated with each word. This classifier is used in the INLP algorithm in order to create a guarding function, $P_{sem}$, for the information necessary to complete the Semantic tag labeling task. As described in the Syntactic probing task section, we shall use the resulting guarding function to decompose the original embedding into a ``non-semantic component'', $v_{no sem} = P_{sem} x$, and a ``semantic component'', $v_{sem} = x - v_{no sem}$.

\subsection{Evaluation tasks}
\label{sec:eval}

Our goal is to determine which information sets captured in the BERT embeddings are relevant for our evaluation tasks. We thus use the components derived from the probing tasks to create new embeddings that isolate specific types of information. These embeddings are then evaluated on the syntactic and semantic tasks that were used for probing, and their performance is compared to that of the original embeddings. We also compare the performance of each model trained on one of these embeddings with another trained on new embeddings that are created by randomly removing the same number of dimensions from the original embeddings as are removed by the INLP guarding function. In doing so we can test the extent to which the loss of the particular information set of interest is responsible for the drop in performance, as opposed to a general loss of information.

The new embeddings to be tested include the syntactic component, the non-syntactic component, the semantic component and the non-semantic component derived from the probing tasks. Additionally, we can create an embedding that contains syntactic information and removes semantic information by linearly projecting the syntactic component onto the non-semantic component, $v_{syn - sem} = P_{sem} v_{syn}$. Using a similar process, we can create an embedding that contains semantic information and removes the syntactic information present, $v_{sem - syn} = P_{syn} v_{sem}$. Finally, we can create an embedding that contains the semantic information captured by the syntactic component, by linearly projecting the syntactic component onto the semantic component, $v_{syn * sem} = (v_{syn} \cdot v_{sem}) * v_{sem}$. And similarly, we can create an embedding that contains the syntactic information captured by the semantic component, $v_{sem * syn} = (v_{sem} \cdot v_{syn}) * v_{syn}$

We will assess each of these embedding types, the original BERT embeddings and the embeddings created by randomly removing directional information on the CCG and Semantic labeling tasks that were used in the probes.

%We have also hypothesized several additional assessment tasks that we would like to undertake, if time permits, or relegate to future work. The first of these tasks is to assess how each embedding type performs on the language modeling task. We would also like to perform the evaluation classification task using a feed-forward neural network with a single hidden layer that contains 10 nodes, in order to determine if there is any task-relevant non-linear information present in the embeddings. If time permits, we would also like to look for patterns in the performance of different embeddings, e.g. explore if a particular embedding type tends to perform better/worse on one of the evaluation tasks for words of a particular POS compared to others. Finally, if time permits we would like to repeat the above procedure to explore the embeddings output by different layers of the BERT model.

%FFNN description is entirely arbitrary, it was just something we talked about during one of our meetings. If anyone has a particular architecture in mind here, please feel free to update this section.

\section{Results}
\label{sec:result}

%Summarize Main findings

This is where our results will go.

\begin{table*}[h]
    \centering
    \begin{tabular}{llll}
    \hline
    \textbf{Embedding} & \textbf{Directions Removed} & \textbf{CCG Tagging} &\textbf{Semantic Labeling} \\
    \hline
    %\multirow{5}{*}{Coding the experiment} & data preprocessing & Qingxia & 05/04 \\
    %& building probing system \& INLP & Saiya \& Qingxia & 05/25 \\
    %& building evaluation system & Gladys & 05/25 \\
    %& merging subparts together & Saiya & 06/02\\ 
    %& running different task pairs & Saiya & 06/02\\
    Original & 0 & p$\%$ & p$\%$ \\
    \hline
    %\multirow{2}{*}{Presenting related topic} & \multirow{2}{*}{presenting INLP method} & \multirow{2}{*}{Lindsay \& Gladys} & \multirow{2}{*}{05/11}\\
    %& & & \\
    Rand($|v_{nosem}|$) & 0 & p$\%$ & p$\%$ \\
    Rand($|v_{nosyn}|$) & 0 & p$\%$ & p$\%$ \\
    Rand($|v_{syn-sem}|$) & 0 & p$\%$ & p$\%$ \\
    Rand($|v_{sem-syn}|$) & 0 & p$\%$ & p$\%$ \\
    Rand($|v_{syn*sem}|$) & 0 & p$\%$ & p$\%$ \\
    Rand($|v_{sem*syn}|$) & 0 & p$\%$ & p$\%$ \\
    \hline
    %\multirow{2}{*}{Writing paper} & \multirow{2}{*}{writing out experiment } & \multirow{2}{*}{Lindsay} & \multirow{2}{*}{06/09}  \\
    %& & & \\
    $v_{nosem}$ & 0 & p$\%$ & p$\%$ \\
    $v_{nosyn}$) & 0 & p$\%$ & p$\%$ \\
    $v_{syn-sem}$ & 0 & p$\%$ & p$\%$ \\
    $v_{sem-syn}$ & 0 & p$\%$ & p$\%$ \\
    $v_{syn*sem}$ & 0 & p$\%$ & p$\%$ \\
    $v_{sem*syn}$ & 0 & p$\%$ & p$\%$ \\
    \hline
    \hline
    \end{tabular}
    \caption{\label{role description} Description of how to read results
    }
    \end{table*}

%In this section, we consider what each evaluation task tells us about the embeddings that are being probed. When we isolate the syntactic component and run it on the syntactic and semantic tasks, we learn how successfully the component responsible for CCG tagging has been isolated, and we also learn how effective the syntactic component alone is on the semantic task. Similarly, running the semantic component on the evaluation tasks tells us how well we've isolated the semantic component and how effective it is on the syntactic task. If the removing of syntactic information leads to an insignificant decrease in the model\rq s performance in a semantic task (or removing semantic information leads to insignificant decrease in performance in a syntactic task), we can conclude that syntactic and semantic information in BERT\rq s representation is linearly separable. Conversely, a significant decrease in performance will indicate low separability between syntactic and semantic information in BERT \rq s representation. Finally, running the non-syntactic and non-semantic components on the evaluation tasks tells us whether any information not identified by INLP is at all useful for the evaluation tasks.

%Next, we consider the potential results of the various projected word embeddings on the evaluation tasks. Each of these tell us how much overlap there is between the syntactic and semantic components of the contextual word embeddings. Projecting the syntactic component onto the non-semantic component removes semantic information from the syntactic component, and projecting the semantic component onto the non-syntactic component removes syntactic information from the semantic component. Projecting the syntactic component onto the semantic component gives us the semantic information that is also part of the syntactic component, and projecting the semantic component onto the syntactic component gives us the syntactic information that is also part of the semantic component. Running all of these embeddings on the semantic and syntactic tasks tells us how separated the semantic and syntactic components are, and how important the overlapping portions are to each task.

\section{Discussion}
\label{sec:discussion}

%What do we learn
%What are some limitations? (only looking at linear info)(only looking at info relevant for these tasks)

This is where our discussion of results will go.

\section{Conclusion}
\label{sec:conclusion}

%Summarize main finding of the paper
%mention future work and extensions

This is where our conclusion will go. This is where our future work will go.

%Move all discussions of future work in the above to here

\bibliography{acl2020.bib}
\bibliographystyle{acl_natbib.bst}

\end{document}

### 2022-06-09 paper/paper.tex
Data section added - needs work
Added the data section with some info about the parallel meaning bank, but was unable to find all relevant info on the sites provided and unable to load the data on my end, so I need someone else to check this and fill in the missing info if possible.

@@
 
-%Description of the data set used and the information it contains relevant to the tasks of interest. Talk about pre-processing the syntax trees to get CCG tags. 
-%We will use the English V4 subset of the CoNLL 2012 shared task data \citep{2012-conll} .
-We must perform two pre-processing steps for this data to be used for our probing and evaluation tasks.
-The first is to apply \citet{ccg-bank}'s CCG Derivation algorithm to the parse tree field in the dataset, in order to create the CCG tags for each word.
-If this proves to be too time-consuming or computationally expensive then we shall change the syntactic probing task to utilize the POS tags available in the dataset, in place of CCG tags.
-The second task is to use the SRL frames in the dataset to generate (verb, BIO-argument-tag) pairs that will act as the categories for the semantic probing task.
-
 %Need to update this section to reflect the new dataset
-We use the NAME data set to test the linear separability of the semantic and syntactic information in word-level BERT embeddings. %NEED TO FILL THIS OUT STILL
+We use the English gold standard of Parallel Meaning Bank v4.0 \citep{pmbData} to test the linear separability of the semantic and syntactic information in word-level BERT embeddings.
+This data set contains word-level semantic tags and NAME FEATURE USED FOR CCG DERIVATIONS %Fill this in
+ that we parse using CITE CCG PARSER %Fill this in
+ to develop word-level CCG tags, both of which are used to train the linear classifiers used for probing and evaluation. The gold-standard data set is comprised of XX words. %Fill this in
+ %Do we drop anything or do any other pre-processing that I missed here?
 

### 2022-06-09 paper/paper.tex
fixed minor typos

@@
 
-\title{Proposal for Analyzing Language Models: Separatablity of Syntax and Semantics}
+\title{Exploring the linear separablity of syntactic and semantic information in BERT embeddings }
 
@@
 
-To evaluate the separability of syntax and semantics, we use two task: one task for the probing system and INLP procedure, and one task for evaluating performance on embeddings before and after INLP.
+To evaluate the separability of syntax and semantics, we use two tasks: one task for the probing system and INLP procedure, and one task for evaluating performance on embeddings before and after INLP.
 We choose Combinatory Categorical Grammar (CCG from here on) tagging \citep{ccg-bank} as the syntactic task and semantic category labeling \citep{propbank} as the semantic task.
@@
 
-Similar to the above, the semantic probing tasks involves training a linear classifier on the final layer BERT embeddings in order to predict the semantic tag (described in the data section) associated with each word.
+Similar to the above, the semantic probing task involves training a linear classifier on the final layer BERT embeddings in order to predict the semantic tag (described in the data section) associated with each word.
 This classifier is used in the INLP algorithm in order to create a guarding function, $P_{sem}$, for the information necessary to complete the Semantic tag labeling task.

### 2022-06-09 paper/paper.tex
adding result

@@
 
-In \citep{disentangle} the authors use paraphrase pairs and new target syntax to train a semantic encoder, syntactic encoder and decoder to learn separate representations of the semantic and syntactic information contained in BART embeddings, in order to create semantically equivalent paraphrases with the new syntactic structure.
+\citealp{disentangle} use paraphrase pairs and new target syntax to train a semantic encoder, syntactic encoder and decoder to learn separate representations of the semantic and syntactic information contained in BART embeddings, in order to create semantically equivalent paraphrases with the new syntactic structure.
 Alongside the encoders they also train an adversarial syntax discriminator to try and predict the source syntax from the semantic embeddings, thus encouraging the disentanglement of the semantic and syntactic information by training the semantic embedder to remove as much syntactic information as possible.
@@
 %Need to update this section to reflect the new dataset
-We use the English gold standard of Parallel Meaning Bank v4.0 \citep{pmbData} to test the linear separability of the semantic and syntactic information in word-level BERT embeddings.
-This data set contains word-level semantic tags and NAME FEATURE USED FOR CCG DERIVATIONS %Fill this in
- that we parse using CITE CCG PARSER %Fill this in
- to develop word-level CCG tags, both of which are used to train the linear classifiers used for probing and evaluation. The gold-standard data set is comprised of XX words. %Fill this in
+% might change later if choose to run experiment in silver
+We use the English Parallel Meaning Bank v4.0 \citep{pmbData} to test the linear separability of the semantic and syntactic information in word-level BERT embeddings.
+This dataset is consist of gold standard and silver standard of word-level semantic tags.
+The gold standard contains 5,438 sentences with annotations that are manually verfied while the silver standard contains 62,739 sentences with autogenerated annotations.
+All of our experiments will be conducted on gold standard data.
+
+The original dataset does not include CCG Tags, but \citealp{pmbData} utilized CCG parser to produce CCG tags.
+We follow a similar procedure to apply a CCG parser \citep{ccg_tagger} to develop word-level CCG tags.
+Once we obtain both CCG tags and semantics tags for the dataset, we can perform syntactic and semantics probing task as desired.%Fill this in
  %Do we drop anything or do any other pre-processing that I missed here?
@@
     \hline
-    %\multirow{5}{*}{Coding the experiment} & data preprocessing & Qingxia & 05/04 \\
-    %& building probing system \& INLP & Saiya \& Qingxia & 05/25 \\
-    %& building evaluation system & Gladys & 05/25 \\
-    %& merging subparts together & Saiya & 06/02\\ 
-    %& running different task pairs & Saiya & 06/02\\
-    Original & 0 & p$\%$ & p$\%$ \\
+    Original & 0 & $84.75\%$ & $88.56\%$ \\
     \hline
-    %\multirow{2}{*}{Presenting related topic} & \multirow{2}{*}{presenting INLP method} & \multirow{2}{*}{Lindsay \& Gladys} & \multirow{2}{*}{05/11}\\
-    %& & & \\
-    Rand($|v_{nosem}|$) & 0 & p$\%$ & p$\%$ \\
-    Rand($|v_{nosyn}|$) & 0 & p$\%$ & p$\%$ \\
-    Rand($|v_{syn-sem}|$) & 0 & p$\%$ & p$\%$ \\
-    Rand($|v_{sem-syn}|$) & 0 & p$\%$ & p$\%$ \\
+
+    Rand($|v_{nosem}|$) & 77 & $84.06\%$ & $87.71\%$ \\
+    Rand($|v_{nosyn}|$) & 159 & $82.26\%$ & $87.57\%$ \\
+    Rand($|v_{syn-sem}|$) & 81 & p$83.11\%$ & $87.61\%$ \\
+    Rand($|v_{sem-syn}|$) & 161 & $82.46\%$ & $86.04\%$ \\
     Rand($|v_{syn*sem}|$) & 0 & p$\%$ & p$\%$ \\
@@
     \hline
-    %\multirow{2}{*}{Writing paper} & \multirow{2}{*}{writing out experiment } & \multirow{2}{*}{Lindsay} & \multirow{2}{*}{06/09}  \\
-    %& & & \\
-    $v_{nosem}$ & 0 & p$\%$ & p$\%$ \\
-    $v_{nosyn}$) & 0 & p$\%$ & p$\%$ \\
-    $v_{syn-sem}$ & 0 & p$\%$ & p$\%$ \\
-    $v_{sem-syn}$ & 0 & p$\%$ & p$\%$ \\
+    $v_{nosem}$ & 77 & $27.93\%$ & $17.28\%$ \\
+    $v_{nosyn}$ & 159 & $23.76\%$ & $49.43\%$ \\
+    $v_{syn-sem}$ & 81 & $34.24\%$ & $33.91\%$ \\
+    $v_{sem-syn}$ & 161 & $26.96\%$ & $50.21\%$ \\
     $v_{syn*sem}$ & 0 & p$\%$ & p$\%$ \\

### 2022-06-09 paper/paper.tex
adding table of descritpion

@@
 
-%Summarize Main findings
+We first probe our tasks on the original embeddings, and linear classifers can successfully predict both CCG tags and semantics tags as resulted in table \ref{role description}.
+Then we apply INLP method to the linear classifers to derive guarding matrices $P_{syn}$ and $P_{sem}$ which will project the embeddings to the nullspace of syntactic space and semantic space respectively.
+With some linear tranformation with the original embeddings and the guarding matrices, we are able to extract the embeddings described in table \ref{description}.
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{p{4cm}p{3cm}}
+        expression & description\\ \hline 
+        $v_{orig}$ & Original BERT embeddings  \\
+        $v_{nosem} = P_{sem} v_{orig}$ & Gained after INLP with semantics task \\
+        $v_{nosyn}= P_{syn} v_{orig}$ & Gained after INLP with syntactic task \\
+        $v_{sem} = v_{orig}-v_{no_sem}$ & Semantics representation \\
+        $v_{syn} = v_{orig}-v_{no_syn}$ & Syntactic representation \\
+        $v_{syn-sem} = P_{sem} v_{syn}$ &Contains syntactic features only \\
+        $v_{sem-syn} = P_{syn} v_{sem}$ &Contains semantic features only \\
+        $v_{syn*sem} = \langle v_{syn}, v_{sem} \rangle \cdot \frac{v_{syn}}{\|v_{sem}\|}$ & Syntactic representations projected on semantic space\\
+        $v_{sem*syn} = \langle v_{syn}, v_{sem} \rangle \cdot \frac{v_{syn}}{\|v_{syn}\|}$ & Semantics representations projected on semantic space\\
 
-This is where our results will go.
+        \hline
+    \end{tabular}
+    \caption{\label{description} Description of Embeddings
+    }
+\end{table}
+
+To further compare the impact of INLP, we
 
@@
     \hline
-    \textbf{Embedding} & \textbf{Directions Removed} & \textbf{CCG Tagging} &\textbf{Semantic Labeling} \\
+    \textbf{Embedding} & \textbf{Directions Removed} & \textbf{CCG Tagging} &\textbf{Semantic Tagging} \\
     \hline
-    Original & 0 & $84.75\%$ & $88.56\%$ \\
+    $v_{orig}$ & 0 & $84.75\%$ & $88.56\%$ \\
     \hline
@@
     \end{tabular}
-    \caption{\label{role description} Description of how to read results
+    \caption{\label{role description} Experiment Result of different Embeddings
     }
@@
 \label{sec:discussion}
+
+The INLP methods
 

### 2022-06-09 paper/paper.tex
adding description

--- 2022-06-09 paper/paper.tex
@@
 Using a similar process, we can create an embedding that contains semantic information and removes the syntactic information present, $v_{sem - syn} = P_{syn} v_{sem}$.
-Finally, we can create an embedding that contains the semantic information captured by the syntactic component, by linearly projecting the syntactic component onto the semantic component, $v_{syn * sem} = (v_{syn} \cdot v_{sem}) * v_{sem}$.
-And similarly, we can create an embedding that contains the syntactic information captured by the semantic component, $v_{sem * syn} = (v_{sem} \cdot v_{syn}) * v_{syn}$
+Finally, we can create an embedding that contains the semantic information captured by the syntactic component, by linearly projecting the syntactic component onto the semantic component, $v_{syn * sem} = \langle v_{syn} ,v_{sem} \rangle u_{sem}$ where $u_{sem}$ is the unit vector for $v_{sem}$.
+And similarly, we can create an embedding that contains the syntactic information captured by the semantic component, $v_{sem * syn} = \langle v_{sem} , v_{syn} \rangle  u_{syn}$
 
@@
 
-%Summarize Main findings
+We first probe our tasks on the original embeddings, and linear classifers can successfully predict both CCG tags and semantics tags (around 85\% accuracy) as resulted in table \ref{role description}.
+Then we apply INLP method to the linear classifers to derive guarding matrices $P_{syn}$ and $P_{sem}$ which will project the embeddings to the nullspace of syntactic space and semantic space respectively.
+With some linear tranformation with the original embeddings and the guarding matrices, we are able to extract the embeddings described in table \ref{description}.
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{p{4cm}p{3cm}}
+        expression & description\\ \hline 
+        $v_{orig}$ & Original BERT embeddings  \\
+        $v_{nosem} = P_{sem} v_{orig}$ & Gained after INLP with semantics task \\
+        $v_{nosyn}= P_{syn} v_{orig}$ & Gained after INLP with syntactic task \\
+        $v_{sem} = v_{orig}-v_{nosem}$ & Semantics representation \\
+        $v_{syn} = v_{orig}-v_{nosyn}$ & Syntactic representation \\
+        $v_{syn-sem} = P_{sem} v_{syn}$ &Contains syntactic features only \\
+        $v_{sem-syn} = P_{syn} v_{sem}$ &Contains semantic features only \\
+        $v_{syn*sem} = \langle v_{syn}, v_{sem} \rangle \cdot \frac{v_{syn}}{\|v_{sem}\|}$ & Syntactic representations projected on semantic space\\
+        $v_{sem*syn} = \langle v_{syn}, v_{sem} \rangle \cdot \frac{v_{syn}}{\|v_{syn}\|}$ & Semantics representations projected on semantic space\\
 
-This is where our results will go.
+        \hline
+    \end{tabular}
+    \caption{\label{description} Description of Embeddings
+    }
+\end{table}
+
+To further compare the impact of INLP, we conduct experiments with random directions removed.
+For each embeddings, we record the number of removed directions using $Null(P)$ where $P$ is the projection matrix.
+For embeddings not derived from matrix multiplication, we obtain the number from $Null(M)$ where $M$ is the embedding matrix with size (768,instance number).\footnote{This is not necessarily equivalent to the number of direction removed.
+$Null(M) \geq Null(P)$ for the corresponding projection $P$ , but in practice the numbers are very close} Then we create embeddings with the same number of directions randomly removed and probe the tasks.
 

### 2022-06-09 src/utils.py
added function to calc kept directions for embeddings
dirKeptCount in utils outputs a tuple of the number of directions kept for each embedding type

@@ -131,3 +131,21 @@ def embeddingComponentBreakdown(x,P_sem,P_syn):
     #sem_syn_emb = sem_emb.dot(syn_emb) * syn_emb/torch.sqrt(torch.sum(syn_emb**2,dim=1))
 
     return no_sem_emb.T, no_syn_emb.T, syn_less_sem_emb.T, sem_less_syn_emb.T, syn_sem_emb.T, sem_syn_emb.T
+
+def dirKeptCount(P_sem,P_syn):
+    no_sem_ct = torch.matrix_rank(P_sem).item()
+    no_syn_ct = torch.matrix_rank(P_sem).item()
+    sem_ct = 768 - no_sem_ct
+    syn_ct = 768 - no_syn_ct
+    mat1 = P_sem - P_sem.matmul(P_syn)
+    syn_less_sem_ct = 768 - torch.matrix_rank(mat1).item()
+    mat2 = P_syn - P_syn.matmul(P_sem)
+    sem_less_syn_ct = 768 - torch.matrix_rank(mat2).item()
+    mat3a = torch.eye(768)-P_sem
+    mat3b = torch.eye(768)-P_syn
+    mat3 = mat3a.matmul(mat3b)
+    syn_sem_ct = torch.matrix_rank(mat3).item()
+    mat4 = mat3b.matmul(mat3a)
+    sem_syn_ct = torch.matrix_rank(mat4).item()
+
+    return syn_ct, sem_ct, syn_less_sem_ct, sem_less_syn_ct, syn_sem_ct, sem_syn_ct

### 2022-06-09 paper/paper.tex
adding disscussion

@@
 
-We first probe our tasks on the original embeddings, and linear classifers can successfully predict both CCG tags and semantics tags (around 85\% accuracy) as resulted in table \ref{role description}.
+We first probe our tasks on the original embeddings, and linear classifers can successfully predict both CCG tags and semantics tags (around 85\% testing accuracy) as resulted in table \ref{role description}.
 Then we apply INLP method to the linear classifers to derive guarding matrices $P_{syn}$ and $P_{sem}$ which will project the embeddings to the nullspace of syntactic space and semantic space respectively.
@@
 $Null(M) \geq Null(P)$ for the corresponding projection $P$ , but in practice the numbers are very close} Then we create embeddings with the same number of directions randomly removed and probe the tasks.
+All the testing accuracy of our final experiments are in table \ref{role description}.
 
@@
     \hline
-    \textbf{Embedding} & \textbf{Directions Removed} & \textbf{CCG Tagging} &\textbf{Semantic Tagging} \\
+    \textbf{Embedding} & \textbf{Directions Removed} & \textbf{CCG Tagging} &\textbf{Semantics Tagging} \\
     \hline
@@
     \end{tabular}
-    \caption{\label{role description} Experiment Result of different Embeddings
+    \caption{\label{role description} Experiment Result of Different Embeddings
     }
@@
 
-The INLP methods
+The INLP method successfully guard information retrieved from probe task.
+When performing CCG tagging, the model accuracy drops significantly (from 84.06 \% to 23.76\%), and similarly for semantics tagging.
+To measure the relevancy of semantics information to syntactic task, we compared the performance of CCG tagging with embeddings $v_{nosem}$ with that of embeddings $v_{nosyn}$.
+The both performances have considerable decrease as to the original embeddings.
+The model is using representation from both space when probe the syntactic task.
+Meanwhile, the performances of semantics tagging with the two embeddings differ significantly, with $v_{nosyn}$ performs almost three times better than $v_{nosem}$.
+The result suggests that the semantics task is less dependent on the syntactic information than the opposite direction, which contradicts the assumption that semantics space will be rel
+
+To investigate the compositionality of semantics space and syntactic space, we also calculated the average cosine simliarity scores between embeddings in table \ref{sim}.
 
@@
 
-This is where our discussion of results will go.
+\begin{table}[h]
+    \centering
+    \begin{tabular}{p{4cm}p{3cm}}
+        expression & similarity score\\ \hline 
+        $v_{orig}, v_{nosem}$ &  $0.4183$ \\ 
+        $v_{orig}, v_{nosyn}$ & $0.3609$\\ 
+        $v_{nosyn}, v_{nosem}$ & $0.2830$ \\
+        $v_{syn-sem}, v_{sem-syn}$ & $0.2579$ \\
+
+        \hline
+    \end{tabular}
+    \caption{\label{sim} Similarity of Semantics and Syntactic Representations
+    }
+\end{table}
 

### 2022-06-09 paper/paper.tex
finished dicussion section

@@
     \centering
-    \begin{tabular}{p{4cm}p{3cm}}
-        expression & description\\ \hline 
+    \begin{tabular}{p{4cm}p{3cm}}\hline
+        \textbf{Expression} & \textbf{Description}\\ \hline 
         $v_{orig}$ & Original BERT embeddings  \\
@@
     Rand($|v_{nosyn}|$) & 159 & $82.26\%$ & $87.57\%$ \\
-    Rand($|v_{syn-sem}|$) & 81 & p$83.11\%$ & $87.61\%$ \\
+    Rand($|v_{syn-sem}|$) & 81 & $83.11\%$ & $87.61\%$ \\
     Rand($|v_{sem-syn}|$) & 161 & $82.46\%$ & $86.04\%$ \\
@@
 \label{sec:discussion}
-
 The INLP method successfully guard information retrieved from probe task.
 When performing CCG tagging, the model accuracy drops significantly (from 84.06 \% to 23.76\%), and similarly for semantics tagging.
+In contrast, when the directions are randomly removed, the performance remains relatively the same for both tasks.
+
 To measure the relevancy of semantics information to syntactic task, we compared the performance of CCG tagging with embeddings $v_{nosem}$ with that of embeddings $v_{nosyn}$.
@@
 Meanwhile, the performances of semantics tagging with the two embeddings differ significantly, with $v_{nosyn}$ performs almost three times better than $v_{nosem}$.
-The result suggests that the semantics task is less dependent on the syntactic information than the opposite direction, which contradicts the assumption that semantics space will be rel
+The result suggests that the semantics task is less dependent on the syntactic information than the opposite direction, which contradicts the assumption that syntactic information will be more pertinent to semantics task than vice versa.
 
-To investigate the compositionality of semantics space and syntactic space, we also calculated the average cosine simliarity scores between embeddings in table \ref{sim}.
-
-%What do we learn
-%What are some limitations? (only looking at linear info)(only looking at info relevant for these tasks)
+We also investigate the compositionality of semantics space and syntactic space, by averaging cosine simliarity scores between embeddings in table \ref{sim}.
+The simliarity score between $v_orig$ and $v_{nosem}$ is higher than between $v_{orig}$ and $v_{nosyn}$, which might suggest that BERT embeddings contain more syntactic information than semantics.
+More experiments need to be conducted to verify this assumption.
 
@@
     \begin{tabular}{p{4cm}p{3cm}}
-        expression & similarity score\\ \hline 
+        \hline
+        \textbf{Expressions} & \textbf{Similarity Score}\\ \hline 
         $v_{orig}, v_{nosem}$ &  $0.4183$ \\ 
@@
 
+%What do we learn
+%What are some limitations? (only looking at linear info)(only looking at info relevant for these tasks)
+
 \section{Conclusion}
 \label{sec:conclusion}
+
+Our probing experiment 
+Though INLP successfully produces the desired guard function, it is worth also worth noting that our dataset is relatively small compared
 

### 2022-06-09 paper/paper.tex
adding conclusion

@@
 This dataset is consist of gold standard and silver standard of word-level semantic tags.
-The gold standard contains 5,438 sentences with annotations that are manually verfied while the silver standard contains 62,739 sentences with autogenerated annotations.
+The gold standard contains 5,438 sentences with annotations that are manually verified while the silver standard contains 62,739 sentences with autogenerated annotations.
 All of our experiments will be conducted on gold standard data.
@@
 \label{sec:conclusion}
+It has been established that linear classifer is successful in various linguistics probing tasks \citep{language-transfer}.
+Our experiment has confirmed linear classifer is robust for probing CCG tagging and semantics tagging on Parallel Meaning Bank \citep{pmbData}.
 
-Our probing experiment 
-Though INLP successfully produces the desired guard function, it is worth also worth noting that our dataset is relatively small compared
+We employed INLP methods and successfully damaged the capablity of linear classifer to probe the designated tasks.
+After gained guard matrices that remove task specific informations from the embeddings, we examined the separablity of semantics and syntactic representation by evaluation tasks on various embeddings.
+Our analysis concluded:first, removing either syntactic or semantic representation will hurt probing tasks; second, syntactic task will be more dependent on semantic information than the opposite direction; last, BERT embeddings is more similar to syntactic representation than semantic representation.
+
+Though INLP successfully produces the desired result, it is worth also worth noting that our dataset is relatively small compared to the number of parameters in linear classifier.
+In the future, we will expect this epxeriment to be reproduced at a larger scale so that the INLP method can be further validated.
 
@@
 
-This is where our conclusion will go. This is where our future work will go.
+% This is where our conclusion will go. This is where our future work will go.
 

### 2022-06-09 paper/paper.tex
commenting out projection part

@@
 Using a similar process, we can create an embedding that contains semantic information and removes the syntactic information present, $v_{sem - syn} = P_{syn} v_{sem}$.
-Finally, we can create an embedding that contains the semantic information captured by the syntactic component, by linearly projecting the syntactic component onto the semantic component, $v_{syn * sem} = \langle v_{syn} ,v_{sem} \rangle u_{sem}$ where $u_{sem}$ is the unit vector for $v_{sem}$.
+% Finally, we can create an embedding that contains the semantic information captured by the syntactic component, by linearly projecting the syntactic component onto the semantic component, $v_{syn * sem} = \langle v_{syn} ,v_{sem} \rangle u_{sem}$ where $u_{sem}$ is the unit vector for $v_{sem}$.
 And similarly, we can create an embedding that contains the syntactic information captured by the semantic component, $v_{sem * syn} = \langle v_{sem} , v_{syn} \rangle  u_{syn}$
@@
         $v_{sem-syn} = P_{syn} v_{sem}$ &Contains semantic features only \\
-        $v_{syn*sem} = \langle v_{syn}, v_{sem} \rangle \cdot \frac{v_{syn}}{\|v_{sem}\|}$ & Syntactic representations projected on semantic space\\
-        $v_{sem*syn} = \langle v_{syn}, v_{sem} \rangle \cdot \frac{v_{syn}}{\|v_{syn}\|}$ & Semantics representations projected on semantic space\\
+        % $v_{syn*sem} = \langle v_{syn}, v_{sem} \rangle \cdot \frac{v_{syn}}{\|v_{sem}\|}$ & Syntactic representations projected on semantic space\\
+        % $v_{sem*syn} = \langle v_{syn}, v_{sem} \rangle \cdot \frac{v_{syn}}{\|v_{syn}\|}$ & Semantics representations projected on semantic space\\
 
@@
     Rand($|v_{sem-syn}|$) & 161 & $82.46\%$ & $86.04\%$ \\
-    Rand($|v_{syn*sem}|$) & 0 & p$\%$ & p$\%$ \\
-    Rand($|v_{sem*syn}|$) & 0 & p$\%$ & p$\%$ \\
+    % Rand($|v_{syn*sem}|$) & 0 & p$\%$ & p$\%$ \\
+    % Rand($|v_{sem*syn}|$) & 0 & p$\%$ & p$\%$ \\
     \hline
@@
     $v_{sem-syn}$ & 161 & $26.96\%$ & $50.21\%$ \\
-    $v_{syn*sem}$ & 0 & p$\%$ & p$\%$ \\
-    $v_{sem*syn}$ & 0 & p$\%$ & p$\%$ \\
+    % $v_{syn*sem}$ & 0 & p$\%$ & p$\%$ \\
+    % $v_{sem*syn}$ & 0 & p$\%$ & p$\%$ \\
     \hline
@@
 
-Though INLP successfully produces the desired result, it is worth also worth noting that our dataset is relatively small compared to the number of parameters in linear classifier.
-In the future, we will expect this epxeriment to be reproduced at a larger scale so that the INLP method can be further validated.
+Though INLP successfully produces the desired result, it is worth noting that our dataset is relatively small compared to the number of parameters in linear classifier.
+In the future, we will expect this epxeriment to be reproduced at a larger scale so that the experiment findings can be further validated.
 

### 2022-06-09 paper/paper.tex
fixing typos

@@
 
-The original dataset does not include CCG Tags, but \citealp{pmbData} utilized CCG parser to produce CCG tags.
-We follow a similar procedure to apply a CCG parser \citep{ccg_tagger} to develop word-level CCG tags.
+The original dataset does not include CCG tags.
+\citealp{pmbData} utilized CCG parser to produce CCG tags.
+We followed a similar procedure and applied a CCG parser \citep{ccg_tagger} to develop word-level CCG tags.
 Once we obtain both CCG tags and semantics tags for the dataset, we can perform syntactic and semantics probing task as desired.%Fill this in

### 2022-06-10 paper/paper.tex
some edits to paper
Still need to edit more

@@
 
-%Need to update this section to reflect the new dataset
-% might change later if choose to run experiment in silver
 We use the English Parallel Meaning Bank v4.0 \citep{pmbData} to test the linear separability of the semantic and syntactic information in word-level BERT embeddings.
-This dataset is consist of gold standard and silver standard of word-level semantic tags.
+This dataset consists of gold standard and silver standard word-level semantic tags.
 The gold standard contains 5,438 sentences with annotations that are manually verified while the silver standard contains 62,739 sentences with autogenerated annotations.
-All of our experiments will be conducted on gold standard data.
+All of our experiments are conducted on gold standard data.
 
-The original dataset does not include CCG tags.
-\citealp{pmbData} utilized CCG parser to produce CCG tags.
-We followed a similar procedure and applied a CCG parser \citep{ccg_tagger} to develop word-level CCG tags.
-Once we obtain both CCG tags and semantics tags for the dataset, we can perform syntactic and semantics probing task as desired.%Fill this in
- %Do we drop anything or do any other pre-processing that I missed here?
+The original dataset does not include CCG tags, however \citealp{pmbData} utilized a CCG parser to produce CCG tags.
+We follow a similar procedure and apply a CCG parser \citep{ccg_tagger} to develop word-level CCG tags.
+Once we obtain both CCG tags and semantics tags for the dataset, we can perform the syntactic and semantics probing task as desired.
 
@@
 We will use this classifier in the INLP algorithm in order to create a guarding function for the information that is necessary to complete the CCG labeling task.
-For a given embedding, $x$, the projection that results from applying this guarding function, $P_{syn}$, to the embedding will represent the non-syntactic information contained in the embedding and will from now on be referred to as the ``non-syntactic component'' of the embedding, $v_{no syn} = P_{syn} x$.
-We can then determine the ``syntactic component'' of the embedding by taking the difference of the embedding vector with the non-syntactic component, $v_{syn} = x - v_{no syn}$.
+For a given embedding, $v_{orig}$, the projection that results from applying this guarding function, $P_{syn}$, to the embedding will represent the non-syntactic information contained in the embedding and will from now on be referred to as the ``non-syntactic component'' of the embedding, $v_{no syn} = P_{syn} v_{orig}$.
+We can then determine the ``syntactic component'' of the embedding by taking the difference of the embedding vector with the non-syntactic component, $v_{syn} = v_{orig} - v_{no syn}$.
 
@@
 This classifier is used in the INLP algorithm in order to create a guarding function, $P_{sem}$, for the information necessary to complete the Semantic tag labeling task.
-As described in the Syntactic probing task section, we shall use the resulting guarding function to decompose the original embedding into a ``non-semantic component'', $v_{no sem} = P_{sem} x$, and a ``semantic component'', $v_{sem} = x - v_{no sem}$.
+As described in the Syntactic probing task section, we shall use the resulting guarding function to decompose the original embedding into a ``non-semantic component'', $v_{no sem} = P_{sem} v_{orig}$, and a ``semantic component'', $v_{sem} = v_{orig} - v_{no sem}$.
 
@@
 
-We first probe our tasks on the original embeddings, and linear classifers can successfully predict both CCG tags and semantics tags (around 85\% testing accuracy) as resulted in table \ref{role description}.
-Then we apply INLP method to the linear classifers to derive guarding matrices $P_{syn}$ and $P_{sem}$ which will project the embeddings to the nullspace of syntactic space and semantic space respectively.
-With some linear tranformation with the original embeddings and the guarding matrices, we are able to extract the embeddings described in table \ref{description}.
+We first evaluate our two tasks on the original embeddings, and determine that linear classifiers can successfully predict both CCG tags and semantics tags (around 85\% testing accuracy), as shown in table \ref{role description}.
+We then apply INLP method to derive the guarding matrices $P_{syn}$ and $P_{sem}$, which are used to project the original embeddings onto the complements of the syntactic information sub-space and the semantic information sub-space.
+By applying linear transformations to the original embeddings and their projections, we are able to extract the embeddings described in table \ref{description}.
 \begin{table}[ht]
@@
         $v_{nosyn}= P_{syn} v_{orig}$ & Gained after INLP with syntactic task \\
-        $v_{sem} = v_{orig}-v_{nosem}$ & Semantics representation \\
+        $v_{sem} = v_{orig}-v_{nosem}$ & Semantic representation \\
         $v_{syn} = v_{orig}-v_{nosyn}$ & Syntactic representation \\

### 2022-06-10 paper/paper.tex
fix citation parentheses

--- 2022-06-09 paper/paper.tex
@@
 We then measure the performance of this new embedding on downstream syntactic and semantic classification tasks.
-The design of our probing procedure follows \citep{amnesia}, which employs INLP to investigate whether BERT uses part-of-speech (POS) information when solving language modeling (LM) tasks.
-INLP has been used for a variety of tasks (\citep{inlp}, \citep{amnesia}, \citep{mbertGreek}, etc.) though this is the first case we know of in which INLP has been used to investigate the linear separability of syntactic and semantic information.
+The design of our probing procedure follows \citealp{amnesia}, which employs INLP to investigate whether BERT uses part-of-speech (POS) information when solving language modeling (LM) tasks.
+INLP has been used for a variety of tasks (\citealp{inlp}; \citealp{amnesia}; \citealp{mbertGreek}, etc.) though this is the first case we know of in which INLP has been used to investigate the linear separability of syntactic and semantic information.
 %why did amnesia have citealp?
@@
 Though they do not achieve perfect separation of the two information sets.
-Other non-linear approaches to syntactic-semantic information disentanglement have been carried out in \citep{multiDis}
+Other non-linear approaches to syntactic-semantic information disentanglement have been carried out in \citealp{multiDis}
 
@@
 
-INLP, introduced in \citep{inlp}, is a method to define a linear guarding function that masks all the linear information in a word embedding that may be used for a downstream classification task.
+INLP, introduced in \citealp{inlp}, is a method to define a linear guarding function that masks all the linear information in a word embedding that may be used for a downstream classification task.
 In the original paper the authors use this method to remove gender bias from BERT embeddings of biographical descriptions and then measure how easy it is to determine an individual's gender from the guarded embedding by using various downstream classification methods.
@@
 
-The authors of \citep{amnesia} use INLP for exactly this task.
+The authors of \citealp{amnesia} use INLP for exactly this task.
 They use INLP in order to separate and guard certain linguistic information sets from BERT embeddings in order to better understand what information is being used by large language models, and not just what is encoded.
@@
 Conversely, if the removal of a property has little influence on the model's ability to perform a task then we know that property is not a significant contributing factor in the model's ability to perform that task.
-Specifically, \citep{amnesia} seeks to quantify the importance of the information sets used for part-of-speech tagging, syntactic dependency labeling, named entity recognition and syntactic constituency boundaries on BERT's ability to perform the language modeling task.
+Specifically, \citealp{amnesia} seeks to quantify the importance of the information sets used for part-of-speech tagging, syntactic dependency labeling, named entity recognition and syntactic constituency boundaries on BERT's ability to perform the language modeling task.
 
-We take a similar approach to \citep{amnesia} by separating the information sets used for CCG tagging and semantic labeling from word-level BERT embeddings.
+We take a similar approach to \citealp{amnesia} by separating the information sets used for CCG tagging and semantic labeling from word-level BERT embeddings.
 However, as we are interested in the linear separability of these two information sets, we will test how the removal of these information sets impacts the embeddings' performance on semantic labeling and CCG tagging, respectively, rather than language modeling.
@@
 
-The INLP method first introduced in \citep{inlp}, is used to create a guarding function that masks all the linear information contained in a set of vectors, $X$, that can be used map each vector to $c \in C$, where $C$ is the set of all categories.
+The INLP method first introduced in \citealp{inlp}, is used to create a guarding function that masks all the linear information contained in a set of vectors, $X$, that can be used map each vector to $c \in C$, where $C$ is the set of all categories.
 This is accomplished by training a linear classifier, a matrix $W$, that is applied to each $x \in X$ in order to predict the correct category $c$ with the greatest possible accuracy.

### 2022-06-10 paper/paper.tex
Edited paper
Removed pieces that were incomplete and made some grammar and spelling fixes

@@
 We seek to explore how representations of these two information sets manifest in BERT embeddings.
-Specifically we investigate the degree of the linear separability of syntactic and semantic information in BERT embeddings, as well as quantify how important the linear component corresponding to one information set is to solving a classification task on the other information set.
+Specifically we investigate the degree of the linear separability of syntactic and semantic information in BERT embeddings, as well as quantify how important the linear component corresponding to one information set is to solving a classification task that targets the other information set.
 We use Iterative Nullspace Projection to decompose word-level BERT embeddings into syntactic, non-syntactic, semantic and non-semantic components to be used in syntactic and semantic classification tasks.
-Our results show that...
+Our results show that there is significant overlap between the syntactic and semantic components in BERT embeddings and the loss of either information set will negatively impact the performance of a linear classifier trained for either of the classification objectives.
+Somewhat surprisingly, we find that the loss of semantic information has a much greater impact on the performance of the syntax classifier than the loss of syntactic information has on the semantic classifier.
 %How we address it and our results.
@@
 The boundary between semantics and syntax is a hotly debated topic in linguistics, but do large language models make such a distinction?
-If they do make such a distinction, do language model embeddings present this information in a way that is easily separated and recognized by humans?
+If so, do language model embeddings present this information in a way that is easily separated and recognized by humans?
 The objective of this project is to explore BERT\rq s \citep{bert} reliance on certain syntactic information when handling a semantic task, and vice versa.
@@
 
-The new embeddings to be tested include the syntactic component, the non-syntactic component, the semantic component and the non-semantic component derived from the probing tasks.
+%The new embeddings to be tested include the non-syntactic component and the non-semantic component derived from the probing tasks.
 Additionally, we can create an embedding that contains syntactic information and removes semantic information by linearly projecting the syntactic component onto the non-semantic component, $v_{syn - sem} = P_{sem} v_{syn}$.
@@
 
-We will assess each of these embedding types, the original BERT embeddings and the embeddings created by randomly removing directional information on the CCG and Semantic labeling tasks that were used in the probes.
+We will assess each of the non-syntactic and non-semantic embedding types, the original BERT embeddings and the embeddings created by randomly removing directional information on the CCG and Semantic labeling tasks that were used in the probes.
 
@@
 
-\section{Results}
-\label{sec:result}
-
-We first evaluate our two tasks on the original embeddings, and determine that linear classifiers can successfully predict both CCG tags and semantics tags (around 85\% testing accuracy), as shown in table \ref{role description}.
-We then apply INLP method to derive the guarding matrices $P_{syn}$ and $P_{sem}$, which are used to project the original embeddings onto the complements of the syntactic information sub-space and the semantic information sub-space.
-By applying linear transformations to the original embeddings and their projections, we are able to extract the embeddings described in table \ref{description}.
 \begin{table}[ht]
@@
         $v_{syn} = v_{orig}-v_{nosyn}$ & Syntactic representation \\
-        $v_{syn-sem} = P_{sem} v_{syn}$ &Contains syntactic features only \\
-        $v_{sem-syn} = P_{syn} v_{sem}$ &Contains semantic features only \\
+        %$v_{syn-sem} = P_{sem} v_{syn}$ &Contains syntactic features only \\
+        %$v_{sem-syn} = P_{syn} v_{sem}$ &Contains semantic features only \\
         % $v_{syn*sem} = \langle v_{syn}, v_{sem} \rangle \cdot \frac{v_{syn}}{\|v_{sem}\|}$ & Syntactic representations projected on semantic space\\
@@
 
-To further compare the impact of INLP, we conduct experiments with random directions removed.
-For each embeddings, we record the number of removed directions using $Null(P)$ where $P$ is the projection matrix.
+\section{Results}
+\label{sec:result}
+
+We first evaluate our two tasks on the original embeddings, and determine that linear classifiers can successfully predict both CCG tags and semantics tags (around 85\% testing accuracy), as shown in table \ref{role description}.
+We then apply the INLP method to derive the guarding matrices $P_{syn}$ and $P_{sem}$, which are used to project the original embeddings onto the complements of the syntactic information sub-space and the semantic information sub-space.
+By applying linear transformations to the original embeddings and their projections, we are able to extract the embeddings described in table \ref{description}.
+
+To ensure a fair assessment of the impact of the information loss, we conduct experiments for which we start with the original BERT embeddings and randomly remove the same number of directions that our derived embeddings lost.
+For the embeddings derived from projection alone, we record the number of removed directions using $Null(P)$, where $P$ is the projection matrix.
 For embeddings not derived from matrix multiplication, we obtain the number from $Null(M)$ where $M$ is the embedding matrix with size (768,instance number).\footnote{This is not necessarily equivalent to the number of direction removed.
-$Null(M) \geq Null(P)$ for the corresponding projection $P$ , but in practice the numbers are very close} Then we create embeddings with the same number of directions randomly removed and probe the tasks.
-All the testing accuracy of our final experiments are in table \ref{role description}.
+$Null(M) \geq Null(P)$ for the embedding matrix $M$ that corresponds to projection $P$, but in practice the numbers are very close.
+With this approach we err on the side of removing more random directions from our random embeddings, resulting in a more conservative comparison when assessing our derived embeddings.} Then we create embeddings with the same number of directions randomly removed and train the linear classifiers on these embeddings.
+The testing accuracies from our experiments can be found in table \ref{role description}.
 
@@
     Rand($|v_{nosyn}|$) & 159 & $82.26\%$ & $87.57\%$ \\
-    Rand($|v_{syn-sem}|$) & 81 & $83.11\%$ & $87.61\%$ \\
-    Rand($|v_{sem-syn}|$) & 161 & $82.46\%$ & $86.04\%$ \\
+    %Rand($|v_{syn-sem}|$) & 81 & $83.11\%$ & $87.61\%$ \\
+    %Rand($|v_{sem-syn}|$) & 161 & $82.46\%$ & $86.04\%$ \\
     % Rand($|v_{syn*sem}|$) & 0 & p$\%$ & p$\%$ \\
@@
     $v_{nosyn}$ & 159 & $23.76\%$ & $49.43\%$ \\
-    $v_{syn-sem}$ & 81 & $34.24\%$ & $33.91\%$ \\
-    $v_{sem-syn}$ & 161 & $26.96\%$ & $50.21\%$ \\
+    %$v_{syn-sem}$ & 81 & $34.24\%$ & $33.91\%$ \\
+    %$v_{sem-syn}$ & 161 & $26.96\%$ & $50.21\%$ \\
     % $v_{syn*sem}$ & 0 & p$\%$ & p$\%$ \\
@@
 \label{sec:discussion}
-The INLP method successfully guard information retrieved from probe task.
-When performing CCG tagging, the model accuracy drops significantly (from 84.06 \% to 23.76\%), and similarly for semantics tagging.
-In contrast, when the directions are randomly removed, the performance remains relatively the same for both tasks.
+The INLP method successfully guards the information used in the probing task.
+When performing CCG tagging, the model accuracy drops significantly when comparing the $v_{nosyn}$ embeddings with the Rand($|v_{nosyn}|$) embeddings (from 82.26\% to 23.76\%), and similarly for semantics tagging (from 87.71\% to 17.28\%).
+In contrast, when the directions are randomly removed, the performance remains relatively the similar to the classifiers' performances on the original BERT embeddings for both tasks (84.75\% and 88.56\%, respectively).
 
-To measure the relevancy of semantics information to syntactic task, we compared the performance of CCG tagging with embeddings $v_{nosem}$ with that of embeddings $v_{nosyn}$.
-The both performances have considerable decrease as to the original embeddings.
+%To measure the importance of semantic information for the syntactic task, we compare the performance of CCG tagging with embeddings $v_{nosem}$ to that of embeddings $v_{nosyn}$.
+Both performances have considerable decrease as to the original embeddings.
 The model is using representation from both space when probe the syntactic task.
@@
 
-We also investigate the compositionality of semantics space and syntactic space, by averaging cosine simliarity scores between embeddings in table \ref{sim}.
-The simliarity score between $v_orig$ and $v_{nosem}$ is higher than between $v_{orig}$ and $v_{nosyn}$, which might suggest that BERT embeddings contain more syntactic information than semantics.
+To measure the importance of semantic information for the syntactic task, we compare the performance of the CCG tagging classifier on $v_{nosem}$ with that of Rand($|v_{nosem}|$) and see a performance drop of 56.13\%.
+Similarly, we can measure the importance of syntactic information for the semantic task by comparing the performance of the semantic tagging classifier on $v_{nosyn}$ with that of Rand($|v_{nosyn}|$) and see a performance drop of 38.14\%.
+We can see that the removal of each information set has a significant impact on the performance of a linear classifier trained on either classification task.
+This suggests that the syntactic and semantic information in BERT embeddings is not easily disentangled.
+
+What is surprising is that the loss of semantic information has a more significant impact on the syntactic classification task than the loss of syntactic information does on the semantic tagging task.
+These results suggests that the semantics task is less dependent on syntactic information than the opposite direction, which challenges the assumption that syntactic information is more pertinent to semantic comprehension than vice versa.
+
+%WOULD LIKE TO INCLUDE THIS IF WE GET SYN/SEM PERFORMANCE
+%We are also interested in the separability of the semantic and syntactic subspaces that arise from BERT embeddings.
+To this end we investigated the performance of embeddings projected onto the the syntactic and semantic sub-spaces with and without the semantic and syntactic information present, respectively.
+We see a significant drop in the performance of the CCG classifier when comparing the performance of $v_{syn}$ embeddings with the performance of $v_{syn-sem}$ embeddings.
+VALUE\%.
+We also see a significant drop in the performance of the semantic classifier when comparing the $v_{sem}$ embeddings with the $v_{sem-syn}$ embeddings, VALUE\%.
+Both of these performance drops indicate that the two information sets are not easily disentangled.
+
+%We also explore the compositionality of the semantic sub-space and syntactic sub-space, by investigating the average cosine-similarity scores between embeddings, found in table \ref{sim}.
+The similarity score between $v_{orig}$ and $v_{nosem}$ is higher than between $v_{orig}$ and $v_{nosyn}$, which might suggest that BERT embeddings contain more syntactic information than semantics.
 More experiments need to be conducted to verify this assumption.
+\footnote{We ran into some concerns when evaluating the $(I-P)$ projection matrices for $v_{syn}$ and $v_{sem}$ that warrant caution when evaluating these results.
+We suspect the nullspace was not properly computed, likely due to computational rounding.
+Thus treat these results with caution.}
 
-\begin{table}[h]
-    \centering
-    \begin{tabular}{p{4cm}p{3cm}}
-        \hline
-        \textbf{Expressions} & \textbf{Similarity Score}\\ \hline 
-        $v_{orig}, v_{nosem}$ &  $0.4183$ \\ 
-        $v_{orig}, v_{nosyn}$ & $0.3609$\\ 
-        $v_{nosyn}, v_{nosem}$ & $0.2830$ \\
-        $v_{syn-sem}, v_{sem-syn}$ & $0.2579$ \\
+%\begin{table}[h]
+    %\centering
+   %\begin{tabular}{p{4cm}p{3cm}}
+        %\hline
+        %\textbf{Expressions} & \textbf{Similarity Score}\\ \hline 
+        %$v_{orig}, v_{nosem}$ &  $0.4183$ \\ 
+        %$v_{orig}, v_{nosyn}$ & $0.3609$\\ 
+        %$v_{nosyn}, v_{nosem}$ & $0.2830$ \\
+        %$v_{syn-sem}, v_{sem-syn}$ & $0.2579$ \\
 
-        \hline
-    \end{tabular}
-    \caption{\label{sim} Similarity of Semantics and Syntactic Representations
-    }
-\end{table}
+        %\hline
+    %\end{tabular}
+    %\caption{\label{sim} Similarity of Semantics and Syntactic Representations
+    %}
+%\end{table}
 
@@
 \label{sec:conclusion}
-It has been established that linear classifer is successful in various linguistics probing tasks \citep{language-transfer}.
-Our experiment has confirmed linear classifer is robust for probing CCG tagging and semantics tagging on Parallel Meaning Bank \citep{pmbData}.
+It has been established that linear classifiers are successful in various linguistics probing tasks \citep{language-transfer}.
+Our experiment has confirmed that linear classifiers can perform CCG tagging and semantic tagging on the Parallel Meaning Bank data set \citep{pmbData} with a fairly high rate of success.
+We then employed INLP and successfully guarded the information contained in BERT embeddings that linear classifiers use to perform the aforementioned classification tasks.
 
-We employed INLP methods and successfully damaged the capablity of linear classifer to probe the designated tasks.
-After gained guard matrices that remove task specific informations from the embeddings, we examined the separablity of semantics and syntactic representation by evaluation tasks on various embeddings.
+Using these guarding functions we were able to explore the importance and separability of the syntactic and semantic information contained in BERT embeddings.
+We evaluated the classification tasks on various derived embeddings and concluded that not only is the linear syntactic and semantic information essential for their respective classification tasks, these information sets are also very important for the opposing classification tasks as well.
+Thus the two information sets are not easily separated.
+Somewhat surprising, we determined that the semantic information was more influential on the success of the syntactic classifier than the other way around.
+%something about linear separability with last two embeddings.
+
+%After gained guard matrices that remove task specific informations from the embeddings, we examined the separablity of semantics and syntactic representation by evaluation tasks on various embeddings.
 Our analysis concluded:first, removing either syntactic or semantic representation will hurt probing tasks; second, syntactic task will be more dependent on semantic information than the opposite direction; last, BERT embeddings is more similar to syntactic representation than semantic representation.
@@
 Though INLP successfully produces the desired result, it is worth noting that our dataset is relatively small compared to the number of parameters in linear classifier.
-In the future, we will expect this epxeriment to be reproduced at a larger scale so that the experiment findings can be further validated.
+In the future, we will expect this experiment to be reproduced at a larger scale so that the experiment findings can be further validated.
+Additionally, in future work we hope to perform a similar analysis on embeddings derived from different layers of the BERT architecture, in order to determine the importance and separability of these information sets at each layer.
+In doing so we hope to better understand how BERT processes different types of linguistic information throughout the encoding process.
 

### 2022-06-10 paper/paper.tex
minor tweaks of the writing; 'semantics tags' --> 'semantic tags'

@@
 %Motivating question and or topics of interest.
-Syntax and semantics are at the foundation every language, yet distinctions between the two are not readily agreed upon.
+Syntax and semantics are at the foundation of every language, yet distinctions between the two are not readily agreed upon.
 We seek to explore how representations of these two information sets manifest in BERT embeddings.
@@
 We follow a similar procedure and apply a CCG parser \citep{ccg_tagger} to develop word-level CCG tags.
-Once we obtain both CCG tags and semantics tags for the dataset, we can perform the syntactic and semantics probing task as desired.
+Once we obtain both CCG tags and semantic tags for the dataset, we can perform the syntactic and semantic probing task as desired.
 
@@
 Similar to the above, the semantic probing task involves training a linear classifier on the final layer BERT embeddings in order to predict the semantic tag (described in the data section) associated with each word.
-This classifier is used in the INLP algorithm in order to create a guarding function, $P_{sem}$, for the information necessary to complete the Semantic tag labeling task.
-As described in the Syntactic probing task section, we shall use the resulting guarding function to decompose the original embedding into a ``non-semantic component'', $v_{no sem} = P_{sem} v_{orig}$, and a ``semantic component'', $v_{sem} = v_{orig} - v_{no sem}$.
+This classifier is used in the INLP algorithm in order to create a guarding function, $P_{sem}$, for the information necessary to complete the semantic tag labeling task.
+As described in the syntactic probing task section, we use the resulting guarding function to decompose the original embedding into a ``non-semantic component'', $v_{no sem} = P_{sem} v_{orig}$, and a ``semantic component'', $v_{sem} = v_{orig} - v_{no sem}$.
 
@@
 
-We will assess each of the non-syntactic and non-semantic embedding types, the original BERT embeddings and the embeddings created by randomly removing directional information on the CCG and Semantic labeling tasks that were used in the probes.
+We will assess each of the non-syntactic and non-semantic embedding types, the original BERT embeddings and the embeddings created by randomly removing directional information on the CCG and semantic labeling tasks that were used in the probes.
 
@@
         $v_{orig}$ & Original BERT embeddings  \\
-        $v_{nosem} = P_{sem} v_{orig}$ & Gained after INLP with semantics task \\
+        $v_{nosem} = P_{sem} v_{orig}$ & Gained after INLP with semantic task \\
         $v_{nosyn}= P_{syn} v_{orig}$ & Gained after INLP with syntactic task \\
@@
 
-We first evaluate our two tasks on the original embeddings, and determine that linear classifiers can successfully predict both CCG tags and semantics tags (around 85\% testing accuracy), as shown in table \ref{role description}.
+We first evaluate our two tasks on the original embeddings, and determine that linear classifiers can successfully predict both CCG tags and semantic tags (around 85\% testing accuracy), as shown in table \ref{role description}.
 We then apply the INLP method to derive the guarding matrices $P_{syn}$ and $P_{sem}$, which are used to project the original embeddings onto the complements of the syntactic information sub-space and the semantic information sub-space.
@@
     \hline
-    \textbf{Embedding} & \textbf{Directions Removed} & \textbf{CCG Tagging} &\textbf{Semantics Tagging} \\
+    \textbf{Embedding} & \textbf{Directions Removed} & \textbf{CCG Tagging} &\textbf{Semantic Tagging} \\
     \hline
@@
 The INLP method successfully guards the information used in the probing task.
-When performing CCG tagging, the model accuracy drops significantly when comparing the $v_{nosyn}$ embeddings with the Rand($|v_{nosyn}|$) embeddings (from 82.26\% to 23.76\%), and similarly for semantics tagging (from 87.71\% to 17.28\%).
+When performing CCG tagging, the model accuracy drops significantly when comparing the $v_{nosyn}$ embeddings with the Rand($|v_{nosyn}|$) embeddings (from 82.26\% to 23.76\%), and similarly for semantic tagging (from 87.71\% to 17.28\%).
 In contrast, when the directions are randomly removed, the performance remains relatively the similar to the classifiers' performances on the original BERT embeddings for both tasks (84.75\% and 88.56\%, respectively).
@@
 What is surprising is that the loss of semantic information has a more significant impact on the syntactic classification task than the loss of syntactic information does on the semantic tagging task.
-These results suggests that the semantics task is less dependent on syntactic information than the opposite direction, which challenges the assumption that syntactic information is more pertinent to semantic comprehension than vice versa.
+These results suggest that the semantics task is less dependent on syntactic information than the opposite direction, which challenges the assumption that syntactic information is more pertinent to semantic comprehension than vice versa.
 

### 2022-07-30 src/LinearClassifier.py
fixing some bugs in the system

@@ -5,6 +5,9 @@ import numpy as np
 import scipy
 from typing import List
 import tqdm
+import random
+
+#random.seed(42)
 
 ## defining GPU here
 device = torch.device("cuda") if torch.cuda.is_available() else torch.device('cpu')
@@ -20,10 +23,11 @@ class LinearClassifier(torch.nn.Module):
 		dev_y: dev label for stopping criterion
 		'''
 		super().__init__()
+		random.seed(42)
 		## everything defined in GPU
-		self.embeddings = input_embeddings
+		self.embeddings = input_embeddings.double()
 		self.output = output
-		self.linear = torch.nn.Linear(input_embeddings.shape[1], tag_size,device=device)
+		self.linear = torch.nn.Linear(input_embeddings.shape[1], tag_size,device=device,dtype=torch.double)
 		# class weight performs really worse
 		# cls_weight = compute_class_weight('balanced',classes=np.array(range(tag_size)),y=output.numpy())
 		# cls_weight = torch.tensor(cls_weight,dtype=torch.float)
@@ -33,11 +37,24 @@ class LinearClassifier(torch.nn.Module):
 		# embedding size = [batch_size, embed_dim]
 		# output size = [batch_size,tag_size]
 		emb = embeddings.to(device)
+		emb = emb.double()
 		fc = self.linear(emb)
 		return fc
 
+	def eval(self,dev_x,dev_y):
+		with torch.no_grad():
+			# TODO: batching if necessary
+			dev_x = dev_x.to(device)
+			dev_y = dev_y.to(device)
+			dev_pred = self.forward(dev_x)
+			loss = self.loss_func(dev_pred,dev_y)
+
+		final_dev =  torch.argmax(dev_pred,dim=1).cpu().numpy()
+		print(f'dev accuracy score:{accuracy_score(dev_y.cpu().numpy(),final_dev):.4f}')
+		return dev_pred, loss.item()
+
 	
-	def optimize(self,lr=0.01,num_epochs=100):
+	def optimize(self,lr=0.01,num_epochs=10):
 		optimizer = torch.optim.AdamW(self.linear.parameters(), lr = lr)
 		best_predictions = None
 		best_loss = float('inf')
@@ -69,7 +86,7 @@ class LinearClassifier(torch.nn.Module):
 		#final_out = output.numpy()
 		#dev_out = self.dev_y.numpy()
 		train_acc = accuracy_score(self.output.numpy(),final_pred)
[…]

### 2022-07-30 src/eval_after_inlp.py

@@ -1,0 +1,54 @@
+import torch
+import numpy as np
+
+from load_bert import DataProcessing
+from LinearClassifier import LinearClassifier, INLPTraining
+from eval_classifier import EvalClassifier
+
+from utils import embeddingComponentBreakdown,remove_random_directions
+
+P_sem = torch.load('sem_space_removed.pt')
+P_syn = torch.load('syn_space_removed.pt')
+
+gold_train = DataProcessing('gold','train')
+gold_train.bert_tokenize()
+train_emb = gold_train.get_bert_embeddings('load')
+
+emb_tr, y_tr_syn,num_tags_syn = gold_train.from_sents_to_words('syn',train_emb)
+
+
+gold_dev = DataProcessing('gold','test')
+gold_dev.bert_tokenize()
+dev_emb = gold_dev.get_bert_embeddings('load')
+
+emb_dev, y_dev_syn, dum = gold_dev.from_sents_to_words('syn',dev_emb)
+
+# print(P_sem.shape)
+# print(P_syn.shape)
+new_emb_trs = embeddingComponentBreakdown(emb_tr.T.double(),P_sem,P_syn)
+new_emb_devs = embeddingComponentBreakdown(emb_dev.T.double(),P_sem,P_syn)
+sim =torch.nn.functional.cosine_similarity(new_emb_trs[1],new_emb_trs[0])
+print(sum(sim)/len(sim))
+
+sys.exit()
+# print(new_emb_trs[4].shape)
+# print(768-torch.matrix_rank(new_emb_trs[4]))
+
+# print(new_emb_trs[2].shape)
+
+# new_emb_tr = torch.matmul(P,emb_tr.T).T
+# new_emb_dev = torch.matmul(P,emb_dev.T).T
+
+# new_emb_tr = remove_random_directions(emb_tr.numpy(),77)
+# new_emb_dev = remove_random_directions(emb_dev.numpy(),77)
+
+# new_emb_tr = torch.tensor(new_emb_tr)
+# new_emb_dev = torch.tensor(new_emb_dev)
+
+t_after = LinearClassifier(new_emb_trs[4],y_tr_syn,num_tags_syn)
+t_after.optimize()
+t_after.eval(new_emb_devs[4],y_dev_syn)
+
+
+# print(768-torch.matrix_rank(P_sem))
+# print(768-torch.matrix_rank(P_syn))

### 2022-07-30 src/eval_classifier.py

@@ -6,23 +6,27 @@ from sklearn.metrics import accuracy_score
 device = torch.device("cuda") if torch.cuda.is_available() else torch.device('cpu')
 
 
-class EvalClassifier(LinearClassifier):
-    def __init__(self,input_embeddings,output,tag_size,dev_x,dev_y):
+class EvalClassifier(torch.nn.Module):
+    def __init__(self,input_embeddings,output,tag_size,hidden_dim,dev_x,dev_y):
         '''
         A Eval Classifier for eval task, inherited from Linear Classifer
         But With batching and different logistics regression
         '''
-        super().__init__(input_embeddings,output,tag_size)
+        super().__init__()
+        self.embeddings = input_embeddings
+        self.output = output
         self.dev_x = dev_x.to(device)
         self.dev_y = dev_y.to(device)
-        # TODO: tranfer output to one hot vector
-        #self.output_vecs = 
-        #self.loss_func = torch.nn.NLLLoss()
+        self.linear = torch.nn.Linear(input_embeddings.shape[1], tag_size,device=device)
+        #self.linear2 = torch.nn.Linear(hidden_dim,tag_size,device=device)
+        self.loss_func = torch.nn.CrossEntropyLoss()
     
     def forward(self,embeddings):
-        fc = self.linear(embeddings)
-        #output_probs = torch.nn.functional.softmax(fc,dim=1)
-        return fc
+        x = self.linear(embeddings)
+        # x = torch.nn.functional.sigmoid(x)
+        # out = self.linear2(x)
+        
+        return x
     
     def batching_data(self,dset='train',batch_size=32):
         if dset == 'train':
@@ -40,8 +44,8 @@ class EvalClassifier(LinearClassifier):
             loss = self.loss_func(dev_pred,self.dev_y)
         return dev_pred, loss.item()
     
-    def optimize(self,batch_size=32, lr=0.001,num_epochs=12):
-        optimizer = torch.optim.Adagrad(self.linear.parameters(), lr = lr)
+    def optimize(self,batch_size=24954, lr=0.01,num_epochs=100):
+        optimizer = torch.optim.AdamW(self.linear.parameters(), lr = lr)
         best_predictions_tr = None
         best_predictions_dv = None
         best_loss=float('inf')

### 2022-07-30 src/generate_labels.py

@@ -1,0 +1,32 @@
+import pandas as pd
+import numpy as np
+import pickle
+from sklearn import preprocessing
+import json
+import sklearn
+from sklearn.model_selection import StratifiedShuffleSplit
+
+df1 = pd.read_pickle('/home2/qg07/homework_test/575project/data/pmb_gold/gold_train.pkl')
+df2 = pd.read_pickle('/home2/qg07/homework_test/575project/data/pmb_gold/gold_dev.pkl')
+df3 = pd.read_pickle('/home2/qg07/homework_test/575project/data/pmb_gold/gold_test.pkl')
+
+a = pd.concat([df1,df2,df3])
+a = a.drop('index',axis=1)
+
+a.to_pickle('../data/pmb_gold/gold_all.pkl')
+
+
+#ss = StratifiedShuffleSplit(n_splits=1,test_size=0.2,random_state=0)
+
+#print(a['semantics_tags'])
+#sys.exit()
+# ccg_tag_list = np.concatenate(a['ccg_tags'].tolist())
+# le = preprocessing.LabelEncoder()
+# p = le.fit_transform(ccg_tag_list)
+
+# print(len(le.classes_))
+# sys
+# d = dict(zip(le.classes_,le.transform(le.classes_).tolist()))
+# with open('../data/pmb_silver/syn_mapping.json','w') as f:
+#     json.dump(d,f)
+# #print(a)

### 2022-07-30 src/load_bert.py

@@ -185,9 +185,9 @@ def encoding_srl(srls:List[Dict], srl_ref=None):
     return srl_ref, res
 
 # loading data from saved pickle files
-train_df = pd.read_pickle('../data/pmb_silver/silver_train.pkl')
-labels = np.concatenate(train_df['ccg_tags'].tolist())
-print(len(np.unique(labels)))
+# train_df = pd.read_pickle('../data/pmb_silver/silver_train.pkl')
+# labels = np.concatenate(train_df['ccg_tags'].tolist())
+# print(len(np.unique(labels)))
 # train_input = train_df['text'].tolist()
 
 # seq_lens = [len(item) for item in train_input]

### 2022-07-30 src/probe.py

@@ -11,44 +11,76 @@ from typing import Dict, List
 from utils import bert_tokenization, get_projection_to_intersection_of_nullspaces, get_rowspace_projection, train_linear_classifier
 import numpy as np
 from tqdm import tqdm
+import random
 
 from load_bert import DataProcessing
 from LinearClassifier import LinearClassifier, INLPTraining
 from eval_classifier import EvalClassifier
+
+from argparse import ArgumentParser
+parser = ArgumentParser(__doc__)
+
+parser.add_argument('--dataset',dest='dataset',type=str, default='gold',help='choosing either gold or silver standard data')
+args = parser.parse_args()
 #logging.basicConfig(level-logging.INFO) #turn on detailed logging
 
 ## defining GPU here
+random.seed(42)
+
 device = torch.device("cuda") if torch.cuda.is_available() else torch.device('cpu')
 
-gold_train = DataProcessing('silver','train')
+gold_train = DataProcessing(args.dataset,'train')
 gold_train.bert_tokenize()
 train_emb = gold_train.get_bert_embeddings('load')
 
 ## all in cpu right now
-emb_tr, y_tr, num_tags = gold_train.from_sents_to_words('sem',train_emb)
+emb_tr, y_tr_sem, num_tags_sem = gold_train.from_sents_to_words('syn',train_emb)
+dum, y_tr_syn,num_tags_syn = gold_train.from_sents_to_words('syn',train_emb)
 
-gold_dev = DataProcessing('silver','dev')
+gold_dev = DataProcessing(args.dataset,'test')
 gold_dev.bert_tokenize()
 dev_emb = gold_dev.get_bert_embeddings('load')
 
 ## all in cpu right now
-emb_dev, y_dev, num_tags = gold_dev.from_sents_to_words('sem',dev_emb)
+emb_dev, y_dev_sem, num_tags_sem = gold_dev.from_sents_to_words('syn',dev_emb)
+dum,y_dev_syn,num_tags_syn =gold_dev.from_sents_to_words('syn',dev_emb)
+
 
 if torch.cuda.is_available():
 	torch.cuda.empty_cache()
-print(num_tags)
 
[…]

### 2022-07-30 src/run.sh

@@ -4,5 +4,5 @@ source ~/anaconda3/etc/profile.d/conda.sh
 
 conda activate news_emo
 
-python load_bert.py
+python probe.py
  

### 2022-07-30 src/utils.py

@@ -113,21 +113,39 @@ def train_linear_classifier(X, y, num_epochs, num_tags, input_dim):
 def embeddingComponentBreakdown(x,P_sem,P_syn):
     """
     Takes in and outputs numpy arrays
-    x is the original BERT embeddings (768,1)
+    x is the original BERT embeddings (768,num_inst)
     P_sem is the projection matrix defined in the INLP loop for the semantic tagging task (768,768)
     P_syn is the projection matrix defined in the INLP loop for the syntactic tagging task (768,768)
     """
+    P_sem = P_sem/torch.norm(P_sem)
+    P_syn = P_syn/torch.norm(P_syn)
+    print(torch.norm(P_sem))
     no_sem_emb = P_sem.matmul(x)
     no_syn_emb = P_syn.matmul(x)
     sem_emb = x - no_sem_emb
     syn_emb = x - no_syn_emb
+
+    #<x-nosem,x-nosyn> = <nosem,nosyn> + <x,x> - <x,P_sem x> - <x,P_syn x>
+    print(torch.matrix_rank(sem_emb))
     syn_less_sem_emb = P_sem.matmul(syn_emb)
     sem_less_syn_emb = P_syn.matmul(sem_emb)
 
-    syn_sem_emb = syn_emb - P_sem.matmul(syn_emb)
-    sem_syn_emb = sem_emb - P_syn.matmul(sem_emb)
-
+    # syn_sem_emb = syn_emb - P_sem.matmul(syn_emb)
+    # sem_syn_emb = sem_emb - P_syn.matmul(sem_emb)
+    
+    dot_prod = torch.mul(sem_emb,syn_emb)
+    #print(dot_prod.shape)
+    dot_prod = torch.sum(dot_prod,dim=0)
+    #print(dot_prod.shape)
+    unit_sem = torch.nn.functional.normalize(sem_emb,dim=0)
+    # print(torch.matrix_rank(no_sem_emb))
+    # print(torch.norm(P_sem))
+    # print(torch.norm(P_syn))
+    unit_syn = torch.nn.functional.normalize(syn_emb,dim=0)
+    syn_sem_emb = dot_prod*unit_sem
+    #print(syn_sem_emb.shape)
+    sem_syn_emb = dot_prod*unit_syn
[…]

### 2022-07-30 paper/paper.tex

@@
 
-A novel method for removing information from an embedding, INLP iteratively trains linear models on a specific classification task, and projects the input on intersection of the nullspaces of those linear models.
+A novel method for removing information from an embedding, INLP iteratively trains linear models on a specific classification task, and projects the input on the intersection of the nullspaces of those linear models.
 Our objective is that, by applying the INLP procedure to a syntactic task, we are able to separate the representation into a syntactic space and a non-syntactic space.
@@
 Conversely, we can also first probe a semantic task, thus defining a semantic and non-semantic space, and then investigate the performance of embeddings projected onto those spaces when performing a syntactic classification task.
-Once we derive the semantic space and syntactic space from the experiment, we further investigate on the separability of the two spaces by comparing the embedding projections.
+The performance of these embeddings on their opposing classification tasks will also give us an indication of how linearly separable the two information sets are.%Once we derive the semantic space and syntactic space from the experiment, we further investigate on the separability of the two spaces by comparing the embedding projections.
 %See section ???
@@
 We follow a similar procedure and apply a CCG parser \citep{ccg_tagger} to develop word-level CCG tags.
-Once we obtain both CCG tags and semantic tags for the dataset, we can perform the syntactic and semantic probing task as desired.
+Once we obtain both CCG tags and semantic tags for the dataset, we can perform the syntactic and semantic probing tasks as desired.
 
@@
 
-Similar to the above, the semantic probing task involves training a linear classifier on the final layer BERT embeddings in order to predict the semantic tag (described in the data section) associated with each word.
+Similar to the above, the semantic probing task involves training a linear classifier on the final layer BERT embeddings in order to predict the semantic tag associated with each word.
 This classifier is used in the INLP algorithm in order to create a guarding function, $P_{sem}$, for the information necessary to complete the semantic tag labeling task.
@@
         $v_{syn} = v_{orig}-v_{nosyn}$ & Syntactic representation \\
+        Rand($n$) & BERT embeddings with n random directions removed \\
         %$v_{syn-sem} = P_{sem} v_{syn}$ &Contains syntactic features only \\
@@
 To ensure a fair assessment of the impact of the information loss, we conduct experiments for which we start with the original BERT embeddings and randomly remove the same number of directions that our derived embeddings lost.
-For the embeddings derived from projection alone, we record the number of removed directions using $Null(P)$, where $P$ is the projection matrix.
-For embeddings not derived from matrix multiplication, we obtain the number from $Null(M)$ where $M$ is the embedding matrix with size (768,instance number).\footnote{This is not necessarily equivalent to the number of direction removed.
+For the embeddings derived from projection matrices, we record the number of removed directions using $Null(P)$, where $P$ is the projection matrix.
+.\footnote{For embeddings not derived from matrix multiplication, we obtain the number from $Null(M)$ where $M$ is the embedding matrix with size (768,instance number).
+This is not necessarily equivalent to the number of direction removed.
 $Null(M) \geq Null(P)$ for the embedding matrix $M$ that corresponds to projection $P$, but in practice the numbers are very close.
-With this approach we err on the side of removing more random directions from our random embeddings, resulting in a more conservative comparison when assessing our derived embeddings.} Then we create embeddings with the same number of directions randomly removed and train the linear classifiers on these embeddings.
+With this approach we err on the side of removing more random directions from our random embeddings, resulting in a more conservative comparison when assessing our derived embeddings.
+Also note that we decided not to include these types of embeddings in our final assessment due to concerning behavior we witnessed after performing the linear transformation to obtain the embeddings.
+Specifically, the rank of the resulting embedding matrix did not match our expectations.
+This may be due to a rounding or an issue with the orthonormalization process, but we were unable to confirm.
+So we decided to remove all embeddings derived in this manner from our evaluation process until we are able to determine the source of this behavior.
+This means that $v_{syn}$ and $v_{sem}$ will not be included in our final evaluation.} Then we create embeddings with the same number of directions randomly removed and train the linear classifiers on these embeddings.
 The testing accuracies from our experiments can be found in table \ref{role description}.
@@
 In contrast, when the directions are randomly removed, the performance remains relatively the similar to the classifiers' performances on the original BERT embeddings for both tasks (84.75\% and 88.56\%, respectively).
+This suggests that the syntactic and semantic information contained in the original BERT embeddings is not highly concentrated and that removing a small amount of one information set will not have a significant impact on the classification task.
 
@@
 What is surprising is that the loss of semantic information has a more significant impact on the syntactic classification task than the loss of syntactic information does on the semantic tagging task.
-These results suggest that the semantics task is less dependent on syntactic information than the opposite direction, which challenges the assumption that syntactic information is more pertinent to semantic comprehension than vice versa.
+These results suggest that the semantic task is less dependent on syntactic information than the opposite direction, which challenges the assumption that syntactic information is more pertinent to semantic comprehension than vice versa.
 
@@
 
-Using these guarding functions we were able to explore the importance and separability of the syntactic and semantic information contained in BERT embeddings.
+Using the INLP-derived guarding functions we were able to explore the importance and separability of the syntactic and semantic information contained in BERT embeddings.
 We evaluated the classification tasks on various derived embeddings and concluded that not only is the linear syntactic and semantic information essential for their respective classification tasks, these information sets are also very important for the opposing classification tasks as well.
@@
 
-Though INLP successfully produces the desired result, it is worth noting that our dataset is relatively small compared to the number of parameters in linear classifier.
+Though INLP successfully produces the desired result, it is worth noting that our dataset is relatively small compared to the number of parameters in the linear classifier.
 In the future, we will expect this experiment to be reproduced at a larger scale so that the experiment findings can be further validated.
-Additionally, in future work we hope to perform a similar analysis on embeddings derived from different layers of the BERT architecture, in order to determine the importance and separability of these information sets at each layer.
+Additionally, we plan to increase the variety of training and evaluation tasks in order to develop a more general understanding of how syntactic and semantic information is encoded in BERT embeddings.
+Finally, we intend to perform a similar analysis on embeddings derived from different layers of the BERT architecture, in order to determine the importance and separability of these information sets at each layer.
 In doing so we hope to better understand how BERT processes different types of linguistic information throughout the encoding process.

### 2022-12-12 paper/paper.tex
1st draft of added layerwise section

@@
 Somewhat surprisingly, we find that the loss of semantic information has a much greater impact on the performance of the syntax classifier than the loss of syntactic information has on the semantic classifier.
+Our results of layer-wise analysis also provide incentive to set an additional stopping condition of INLP, regarding the rank of the projection matrix.
 %How we address it and our results.
@@
 
+\subsection{Layer-wise evaluation}
+In addition to the final layer BERT embeddings, we perform a similar analysis on the embeddings derived from different layers of the BERT architecture, in order to determine the separability of these information sets at each layer.
+For embedding $v_{orig_i}$ from layer $i,$ a linear classifier is trained for each probing task to acquire guarding functions $P_{syn_i}$ and $P_{sem_i},$ respectively.
+Applying these projection functions, we are able to acquire $v_{nosyn_i}$ and $v_{nosem_i}.$ Subtracting them from the original embedding, we get the semantic representation $v_{sem_i}$ and the syntatic representation $v_{syn_i}.$ We also randomly remove the same number of dimensions in the original embedding for comparison.
+
+By comparing the experiment results across different information sets and different layers, we hope to better understand how BERT processes different types of linguistic information throughout the encoding process.
+
 \begin{table}[ht]
@@
         $v_{syn} = v_{orig}-v_{nosyn}$ & Syntactic representation \\
-        Rand($n$) & BERT embeddings with n random directions removed \\
+        Rand($v$, $n$) & embeddings $v$ with n random directions removed \\
+        Subscript $i$ & Objects relating to the $i$-th layer of the BERT architecture rather than the final layer\\
         %$v_{syn-sem} = P_{sem} v_{syn}$ &Contains syntactic features only \\
@@
 \label{sec:result}
-
+\subsection{Final layer evaluation}
 We first evaluate our two tasks on the original embeddings, and determine that linear classifiers can successfully predict both CCG tags and semantic tags (around 85\% testing accuracy), as shown in table \ref{role description}.
@@
 For the embeddings derived from projection matrices, we record the number of removed directions using $Null(P)$, where $P$ is the projection matrix.
-.\footnote{For embeddings not derived from matrix multiplication, we obtain the number from $Null(M)$ where $M$ is the embedding matrix with size (768,instance number).
+\footnote{For embeddings not derived from matrix multiplication, we obtain the number from $Null(M)$ where $M$ is the embedding matrix with size (768,instance number).
 This is not necessarily equivalent to the number of direction removed.
@@
 
-    Rand($|v_{nosem}|$) & 77 & $84.06\%$ & $87.71\%$ \\
-    Rand($|v_{nosyn}|$) & 159 & $82.26\%$ & $87.57\%$ \\
+    Rand($v_orig, |v_{nosem}|$) & 77 & $84.06\%$ & $87.71\%$ \\
+    Rand($|v_orig, v_{nosyn}|$) & 159 & $82.26\%$ & $87.57\%$ \\
     %Rand($|v_{syn-sem}|$) & 81 & $83.11\%$ & $87.61\%$ \\
@@
     \end{table*}
+
+\subsection{Intermediate Layer Evaluation}
+
+Linear classifiers are generally able to achieve a test accuracy greater than 90\% for both CCG tagging and semantics tagging.
+However, after applying the INLP algorithm and projecting the layer-wise embeddings to the nullspaces of the information sets, we observe that we get the same results across all layers for each evaluation task, with the accuracy being the majority class of the evaluation task.
+Evaluations of Rand($v_i, |v_{nosyn_i}|$) and Rand($v_i, |v_{nosem_i}|$) result in the same majority class accuracy.
+
+Upon a close inspection of the INLP process and the projections of the original intermediate layer embeddings, $v_{nosem_i}$ and $v_{nosyn_i}$, we realize that, the INLP process continues to run even if it already removes more ranks than BERT’s hidden size, which is 768 in our case, because the desired dev accuracy is still not met.
+Once the rank of the projection matrix reaches the limit, the INLP process simply reduces the magnitude of each elements in the embeddings.
+In some cases, the process eventually zeroes out the embeddings, which explains the identical yet trivial result we get from the evaluation task.
 
@@
 
-To measure the importance of semantic information for the syntactic task, we compare the performance of the CCG tagging classifier on $v_{nosem}$ with that of Rand($|v_{nosem}|$) and see a performance drop of 56.13\%.
-Similarly, we can measure the importance of syntactic information for the semantic task by comparing the performance of the semantic tagging classifier on $v_{nosyn}$ with that of Rand($|v_{nosyn}|$) and see a performance drop of 38.14\%.
+To measure the importance of semantic information for the syntactic task, we compare the performance of the CCG tagging classifier on $v_{nosem}$ with that of Rand($v_orig, |v_{nosem}|$) and see a performance drop of 56.13\%.
+Similarly, we can measure the importance of syntactic information for the semantic task by comparing the performance of the semantic tagging classifier on $v_{nosyn}$ with that of Rand($v_{orig}, |v_{nosyn}|$) and see a performance drop of 38.14\%.
 We can see that the removal of each information set has a significant impact on the performance of a linear classifier trained on either classification task.
@@
 These results suggest that the semantic task is less dependent on syntactic information than the opposite direction, which challenges the assumption that syntactic information is more pertinent to semantic comprehension than vice versa.
+
+As an unexpected finding of our experiment, we are unable to fully remove the syntactic/semantic information from the intermediate embedding by training the linear classifier to make prediction that is no better than the majority, without removing more ranks than BERT's hidden size, on all intermediate layers of BERT.
+However, removing more ranks than BERT’s hidden size, whether through the INLP algorithm or randomly, results in a degenerate embedding where every element is reduced to an extremely small magnitude that the linear probe on the evaluation task will only reach the majority class accuracy.
+This seems to reveal that, on the intermediate layers of BERT, the target information is not linearly separable from the original embedding.
+Besides using the majority class accuracy as the stopping condition, researchers hoping to use INLP to guard information from BERT embeddings should also make sure the loop stops before removing too many ranks for non-trivial results.
 
@@
 
+By performing a similar analysis on embeddings derived from different layers of BERT architecture, we discovered that the syntactic information set and semantic information set are not linearly separable from the original intermediate layer embeddings.
+Attempting to completely remove these information sets will remove too much that the embeddings become degenerate.
+
 %After gained guard matrices that remove task specific informations from the embeddings, we examined the separablity of semantics and syntactic representation by evaluation tasks on various embeddings.
@@
 
-Though INLP successfully produces the desired result, it is worth noting that our dataset is relatively small compared to the number of parameters in the linear classifier.
-In the future, we will expect this experiment to be reproduced at a larger scale so that the experiment findings can be further validated.
-Additionally, we plan to increase the variety of training and evaluation tasks in order to develop a more general understanding of how syntactic and semantic information is encoded in BERT embeddings.
-Finally, we intend to perform a similar analysis on embeddings derived from different layers of the BERT architecture, in order to determine the importance and separability of these information sets at each layer.
-In doing so we hope to better understand how BERT processes different types of linguistic information throughout the encoding process.
+Though INLP successfully produces interesting results on various tasks, it is worth noting that our dataset is relatively small compared to the number of parameters in the linear classifier.
+Reproduing this experiment at a larger scale will be helpful in further validating the experiment results.
+Additionally, the variety of training and evaluation tasks can be increased for a broader understanding of how syntactic and semantic information is encoded in BERT embeddings.
 

### 2022-12-15 paper/paper.tex
shorten abstract, section 1 and 2

@@
 %Motivating question and or topics of interest.
-Syntax and semantics are at the foundation of every language, yet distinctions between the two are not readily agreed upon.
+Distinctions between Syntax and Semantics are not readily agreed upon.
 We seek to explore how representations of these two information sets manifest in BERT embeddings.
@@
 We use Iterative Nullspace Projection to decompose word-level BERT embeddings into syntactic, non-syntactic, semantic and non-semantic components to be used in syntactic and semantic classification tasks.
-Our results show that there is significant overlap between the syntactic and semantic components in BERT embeddings and the loss of either information set will negatively impact the performance of a linear classifier trained for either of the classification objectives.
-Somewhat surprisingly, we find that the loss of semantic information has a much greater impact on the performance of the syntax classifier than the loss of syntactic information has on the semantic classifier.
-Our results of layer-wise analysis also provide incentive to set an additional stopping condition of INLP, regarding the rank of the projection matrix.
+Our results show that there is significant overlap between the syntactic and semantic components in BERT embeddings to the degree that they are not linearly separable.
+Our results also indicate a factor of consideration when applying INLP, regarding the rank of the projection matrix.
 %How we address it and our results.
@@
 
-The boundary between semantics and syntax is a hotly debated topic in linguistics, but do large language models make such a distinction?
-If so, do language model embeddings present this information in a way that is easily separated and recognized by humans?
+The boundary between semantics and syntax has been hotly debated, but do language model embeddings present this information in a way that is easily separated and recognized by humans?
 The objective of this project is to explore BERT\rq s \citep{bert} reliance on certain syntactic information when handling a semantic task, and vice versa.
@@
 The design of our probing procedure follows \citealp{amnesia}, which employs INLP to investigate whether BERT uses part-of-speech (POS) information when solving language modeling (LM) tasks.
-INLP has been used for a variety of tasks (\citealp{inlp}; \citealp{amnesia}; \citealp{mbertGreek}, etc.) though this is the first case we know of in which INLP has been used to investigate the linear separability of syntactic and semantic information.
-%why did amnesia have citealp?
 
@@
 Our objective is that, by applying the INLP procedure to a syntactic task, we are able to separate the representation into a syntactic space and a non-syntactic space.
-We then compare the performance of a model that seeks to classify semantic labels using the original BERT embeddings with an otherwise identical model trained on embeddings projected onto the non-syntactic space, in order to see if BERT is using syntactic information when performing the semantic task.
-Conversely, we can also first probe a semantic task, thus defining a semantic and non-semantic space, and then investigate the performance of embeddings projected onto those spaces when performing a syntactic classification task.
-The performance of these embeddings on their opposing classification tasks will also give us an indication of how linearly separable the two information sets are.%Once we derive the semantic space and syntactic space from the experiment, we further investigate on the separability of the two spaces by comparing the embedding projections.
+We then compare the performance of a linear classifier for semantic labels using the original BERT embeddings with an otherwise identical model trained on embeddings projected onto the non-syntactic space.
+Conversely, we can define a semantic and non-semantic space by probing a semantic task, and then investigate the performance of embeddings projected onto those spaces when performing a syntactic classification task.
+The performance of these embeddings on their opposing classification tasks will give us an indication of how linearly separable the two information sets are.%Once we derive the semantic space and syntactic space from the experiment, we further investigate on the separability of the two spaces by comparing the embedding projections.
 %See section ???
@@
 
-The separation and overlap between syntax and semantics has been of interest to linguists for many years.
-More recently, with the growing popularity of large language models, computational linguists have begun to explore how large language models deal with the boundaries of these information sets in word and sentence embeddings.
+The separation and overlap between syntax and semantics has been of interest to linguists for years.
+More recently, with the growing popularity of large language models, computational linguists have begun to explore how large language models deal with the boundaries of these information sets.
 
@@
 Alongside the encoders they also train an adversarial syntax discriminator to try and predict the source syntax from the semantic embeddings, thus encouraging the disentanglement of the semantic and syntactic information by training the semantic embedder to remove as much syntactic information as possible.
-Their results show that one can achieve some removal of syntactic information from semantic embeddings, so disentanglement of some information is possible.
+Their results show that disentanglement of some information is possible.
 Though they do not achieve perfect separation of the two information sets.
@@
 In the original paper the authors use this method to remove gender bias from BERT embeddings of biographical descriptions and then measure how easy it is to determine an individual's gender from the guarded embedding by using various downstream classification methods.
-Beyond using the INLP method to guard protected attributes, the authors hypothesize several additional use cases for this procedure, including information disentanglement.
+Beyond this example, the authors hypothesize several additional use cases for this procedure, including information disentanglement.
 
 The authors of \citealp{amnesia} use INLP for exactly this task.
-They use INLP in order to separate and guard certain linguistic information sets from BERT embeddings in order to better understand what information is being used by large language models, and not just what is encoded.
+They use INLP to separate and guard certain linguistic information sets from BERT embeddings in order to better understand what information is being used by large language models, and not just what is encoded.
 The main premise behind this paper is that if a particular property is used to solve a task, then the removal of that property should negatively influence the model's ability to solve that task.
-Conversely, if the removal of a property has little influence on the model's ability to perform a task then we know that property is not a significant contributing factor in the model's ability to perform that task.
 Specifically, \citealp{amnesia} seeks to quantify the importance of the information sets used for part-of-speech tagging, syntactic dependency labeling, named entity recognition and syntactic constituency boundaries on BERT's ability to perform the language modeling task.
 
-We take a similar approach to \citealp{amnesia} by separating the information sets used for CCG tagging and semantic labeling from word-level BERT embeddings.
-However, as we are interested in the linear separability of these two information sets, we will test how the removal of these information sets impacts the embeddings' performance on semantic labeling and CCG tagging, respectively, rather than language modeling.
+We take a similar approach to \citealp{amnesia} by separating the information sets used for CCG tagging and semantic labeling from word-level BERT embeddings, and test how the removal of these information sets impacts the embeddings' performance on these tasks.
 
@@
 
-    Rand($v_orig, |v_{nosem}|$) & 77 & $84.06\%$ & $87.71\%$ \\
-    Rand($|v_orig, v_{nosyn}|$) & 159 & $82.26\%$ & $87.57\%$ \\
+    Rand($v_{orig}, |v_{nosem}|$) & 77 & $84.06\%$ & $87.71\%$ \\
+    Rand($v_{orig}, |v_{nosyn}|$) & 159 & $82.26\%$ & $87.57\%$ \\
     %Rand($|v_{syn-sem}|$) & 81 & $83.11\%$ & $87.61\%$ \\

### 2022-12-15 src/LinearClassifier.py
increase # of epochs and fixing some small issues

@@ -52,7 +52,7 @@ class LinearClassifier(torch.nn.Module):
 
 		final_dev =  torch.argmax(dev_pred,dim=1).cpu().numpy()
 		acc = accuracy_score(dev_y.cpu().numpy(),final_dev)
-		print(f'dev accuracy score:{acc:.4f}')
+		#print(f'dev accuracy score:{acc:.4f}')
 		return dev_pred, loss.item(),acc
 
 	
@@ -61,10 +61,11 @@ class LinearClassifier(torch.nn.Module):
 		dataloader = DataLoader(data_set,batch_size=batch_size)
 		return dataloader
 	
-	def optimize(self,lr=0.01,num_epochs=10):
+	def optimize(self,lr=0.01,num_epochs=100,dev_x=None,dev_y=None):
 		optimizer = torch.optim.AdamW(self.linear.parameters(), lr = lr)
 		best_predictions = None
 		best_loss = float('inf')
+		best_dev_acc=0
 		stop_count = 0
 		output = self.output.to(device)
 		dataloader = self.batched_input(self.embeddings,output)
@@ -81,15 +82,20 @@ class LinearClassifier(torch.nn.Module):
 				preds.append(pred)
 				total_loss += loss.item()
 			
-			total_loss = total_loss/len(dataloader)
 			preds = torch.cat(preds)
 			#print(f'epoch: {epoch+1}, loss = {loss.item():.4f}')
 			#implement stopping criterion
+			if dev_x is not None:
+				d1,total_loss,dev_acc = self.eval(dev_x,dev_y)
+			else:
+				total_loss = total_loss/len(dataloader)
 			if total_loss<best_loss:
 				best_loss=total_loss
 				best_model= self.linear
 				best_predictions = preds
 				stop_count=0
+				if dev_x is not None:
+					best_dev_acc = dev_acc
 			else:
 				if stop_count ==3:
 					break
@@ -101,9 +107,10 @@ class LinearClassifier(torch.nn.Module):
 		#dev_out = self.dev_y.numpy()
 		train_acc = accuracy_score(self.output.numpy(),final_pred)
 		print(f'train accuracy score:{train_acc:.4f}')
+		print(f'dev accuracy score{best_dev_acc:.4f}')
 		#print(f'dev accuracy score:{accuracy_score(self.dev_y.numpy(),final_dev):.4f}')
 			
-		return best_model,train_acc
[…]

### 2022-12-15 src/eval_after_inlp.py

@@ -13,6 +13,7 @@ parser = ArgumentParser(__doc__)
 
 parser.add_argument('--random',dest='random',action='store_true',default=False,help='nulling out random directions')
 parser.add_argument('--task',dest='task',type=str,default='syn',help='choosing a task =[syn|sem] for probing task after INLP,default is syn')
+parser.add_argument('--layer',dest='layer',type=int, default=-1)
 
 args = parser.parse_args()
 
@@ -23,7 +24,10 @@ if not args.random:
 else:
     print('null task: random')
 
-P_null = torch.load('../data/pmb_gold/'+null_task+'_space_removed.pt')
+if args.layer !=-1:
+    P_null = torch.load('../data/pmb_gold/'+null_task+'_space_removed_'+str(args.layer)+'.pt')
+else:
+    P_null = torch.load('../data/pmb_gold/'+null_task+'_space_removed.pt')
 print('evaluation task: '+ args.task)
 
 
@@ -33,16 +37,23 @@ print('evaluation task: '+ args.task)
 
 gold_train = DataProcessing('gold','train')
 gold_train.bert_tokenize()
-train_emb = gold_train.get_bert_embeddings('load')
+train_emb = gold_train.get_bert_embeddings('load',layerwise=args.layer)
 
 emb_tr, y_tr,num_tags = gold_train.from_sents_to_words(args.task,train_emb)
 
 
-gold_dev = DataProcessing('gold','test')
+gold_dev = DataProcessing('gold','dev')
 gold_dev.bert_tokenize()
-dev_emb = gold_dev.get_bert_embeddings('load')
+dev_emb = gold_dev.get_bert_embeddings('load',layerwise=args.layer)
 
 emb_dev, y_dev, dum = gold_dev.from_sents_to_words(args.task,dev_emb)
+
+
+gold_test = DataProcessing('gold','test')
+gold_test.bert_tokenize()
+test_emb = gold_test.get_bert_embeddings('load',layerwise=args.layer)
+
+emb_test, y_test, dum = gold_test.from_sents_to_words(args.task,test_emb)
 
 # print(P_sem.shape)
 # print(P_syn.shape)
@@ -75,13 +86,16 @@ emb_dev, y_dev, dum = gold_dev.from_sents_to_words(args.task,dev_emb)
 if not args.random:
     train_nulled = P_null.matmul(emb_tr.T.double()).T
[…]

### 2022-12-15 src/probe.py

@@ -22,219 +22,85 @@ parser = ArgumentParser(__doc__)
 
 parser.add_argument('--dataset',dest='dataset',type=str, default='gold',help='choosing either gold or silver standard data')
 parser.add_argument('--task',dest='task',type=str,default='syn',help='choosing a task =[syn|sem] for INLP,default is syn')
+parser.add_argument('--layer',dest='layer',type=int,default=-1,help='choosing if running on layer embeddings')
 parser.add_argument('--load',dest='load',action='store_true', default=False,help='load the preexisting embeddings if it is already generated')
 args = parser.parse_args()
-#logging.basicConfig(level-logging.INFO) #turn on detailed logging
 
-## defining GPU here
+
+
 random.seed(42)
 
 load_option = 'load' if args.load else 'save'
-# layerwise = 4
-
-investigate_layer = 3
-projection_layer = 5
+layerwise= args.layer
 
 
+# defining GPU here
 device = torch.device("cuda") if torch.cuda.is_available() else torch.device('cpu')
 
 
-# print('Loading Dataset: '+args.dataset)
+print('Loading Dataset: '+args.dataset)
 
-# gold_train = DataProcessing(args.dataset,'train')
-# gold_train.bert_tokenize()
-# train_emb = gold_train.get_bert_embeddings(load_option, layerwise=layerwise)
+train_data = DataProcessing(args.dataset,'train')
+train_data.bert_tokenize()
+train_emb_by_sents = train_data.get_bert_embeddings(load_option, layerwise=layerwise)
 
-# ## all in cpu right now
-# emb_tr, y_tr_sem, num_tags_sem = gold_train.from_sents_to_words(args.task,train_emb)
-# dum, y_tr_syn,num_tags_syn = gold_train.from_sents_to_words(args.task,train_emb)
-
-# gold_dev = DataProcessing(args.dataset,'test')
-# gold_dev.bert_tokenize()
-# dev_emb = gold_dev.get_bert_embeddings(load_option, layerwise=layerwise)
-
-# ## all in cpu right now
-# emb_dev, y_dev_sem, num_tags_sem = gold_dev.from_sents_to_words(args.task,dev_emb)
[…]

### 2022-12-15 src/run.sh

@@ -2,6 +2,8 @@
 
 source ~/anaconda3/etc/profile.d/conda.sh
 
-conda activate 575env
+conda activate news_emo
 
-python debug.py --dataset gold --task sem
+python probe.py --dataset gold --task syn --layer 8
+
+#python eval_after_inlp.py --random --task syn

### 2022-12-16 paper/paper.tex
updating the result

@@
 
-\aclfinalcopy % Uncomment this line for the final submission
+%\aclfinalcopy % Uncomment this line for the final submission
 %\def\aclpaperid{***} %  Enter the acl Paper ID here
@@
 %Motivating question and or topics of interest.
-Distinctions between Syntax and Semantics are not readily agreed upon.
-We seek to explore how representations of these two information sets manifest in BERT embeddings.
-Specifically we investigate the degree of the linear separability of syntactic and semantic information in BERT embeddings, as well as quantify how important the linear component corresponding to one information set is to solving a classification task that targets the other information set.
-We use Iterative Nullspace Projection to decompose word-level BERT embeddings into syntactic, non-syntactic, semantic and non-semantic components to be used in syntactic and semantic classification tasks.
-Our results show that there is significant overlap between the syntactic and semantic components in BERT embeddings to the degree that they are not linearly separable.
-Our results also indicate a factor of consideration when applying INLP, regarding the rank of the projection matrix.
+Relations between syntax and semantics are not readily agreed upon.
+We seek to explore how representations of syntax and semantic information sets manifest in BERT embeddings, particularly the degree of the linear separability of each other in BERT embeddings by applying Iterative Nullspace Projection (INLP) to decompose BERT embeddings into syntactic and semantic subspaces.
+We also investigate how important the linear component corresponding to one information set is to solving a classification task that targets the other information set.
+Our results show that both syntactic and semantics informations are not linearly represented in BERT embeddings.
+Therefore INLP fails separate syntactic and semantic space from BERT embeddings and does not provide interpretable results.
+The results also indicate a factor of consideration when applying INLP, regarding the rank of the projection matrix.
 %How we address it and our results.
@@
 %may remove this
-To achieve our goal, we construct a linear probing system for a task and then employ Iterative Nullspace Projection (INLP from here on) \citep{inlp} to generate a new embedding devoid of information learned from the probing task.
-We then measure the performance of this new embedding on downstream syntactic and semantic classification tasks.
-The design of our probing procedure follows \citealp{amnesia}, which employs INLP to investigate whether BERT uses part-of-speech (POS) information when solving language modeling (LM) tasks.
+To achieve our goal, we apply a novel method Iterative Nullspace Projection (INLP from here) \citep{inlp}for removing information from an embedding.
+INLP iteratively trains linear models on a specific classification task, and projects the input on the intersection of the nullspaces of those linear models.
 
-A novel method for removing information from an embedding, INLP iteratively trains linear models on a specific classification task, and projects the input on the intersection of the nullspaces of those linear models.
+Our experiment scheme follows \citealp{amnesia}, which employs INLP to investigate whether BERT uses part-of-speech (POS) information when solving language modeling (LM) tasks.
+Similarly, we construct a linear probing system for a task and then employ INLP to generate a new embedding devoid of information learned from the probing task.
+We then evaluate the performance of this new embedding on another downstream task.
+Then we will perform the same procedure but switch the probing task and downstream evaluating task.
+To evaluate the separability of syntactic and semantic representation, we need two tasks that could extract those information on word level.
+Hence, we choose Combinatory Categorical Grammar (CCG from here on) tagging \citep{ccg-bank} as the syntactic task and semantic tagging \citep{semantics_tagging} as the semantic task.
+
+% insert a figure if possible
 Our objective is that, by applying the INLP procedure to a syntactic task, we are able to separate the representation into a syntactic space and a non-syntactic space.
@@
 %Vice versa? (Just be very explicit here)
-
-To evaluate the separability of syntax and semantics, we use two tasks: one task for the probing system and INLP procedure, and one task for evaluating performance on embeddings before and after INLP.
-We choose Combinatory Categorical Grammar (CCG from here on) tagging \citep{ccg-bank} as the syntactic task and semantic category labeling \citep{propbank} as the semantic task.
 
@@
 Unlike the aforementioned studies, we seek to explore the linear separability of syntactic and semantic information in large language model embeddings at the word level.
-To accomplish this task we apply the Iterative Nullspace Projection method to syntactic (CCG) and semantic labeling tasks in order to define the syntactic and semantic components of BERT embeddings that will be used in our downstream classification tasks.
+To accomplish this task we apply the INLP method to syntactic (CCG) and semantic tasks in order to define the syntactic and semantic components of BERT embeddings that will be used in our downstream classification tasks.
 
@@
 
-We take a similar approach to \citealp{amnesia} by separating the information sets used for CCG tagging and semantic labeling from word-level BERT embeddings, and test how the removal of these information sets impacts the embeddings' performance on these tasks.
+We take a similar approach to \citealp{amnesia} by separating the information sets used for CCG tagging and semantic tagging from word-level BERT embeddings, and test how the removal of these information sets impacts the embeddings' performance on these tasks.
 
-\section{Methods}
+\section{Experiment}
 \label{sec:method}
@@
 
-We construct two separate probing tasks to isolate the syntactic and semantic information in word-level BERT embeddings.
-The embeddings are separated into syntactic and non-syntactic, and semantic and non-semantic components via INLP which is described in section \ref{sec:inlp-method}.
-These embedding components are then combined to form new embeddings, which are evaluated on the same tasks that were used for probing.
+To isolate the syntactic and semantic information from word-level BERT embeddings efficiently, we implement INLP using method described in section \ref{sec:inlp-method}.
+CCG tagging and semantics tagging are probing tasks for INLP to extract relevant information from embeddings, which are described in section \ref{sec:data},\ref{sec:probing} .
+%These embedding components are then combined to form new embeddings, which are evaluated on the same tasks that were used for probing.%
+ We also conduct experiments using BERT embeddings from different layers to see which layer might contain more syntactic or semantics information, as described in \ref{layer}.
 %If time permits, we will also evaluate how these new embeddings perform on the language modeling task.
@@
 This is accomplished by training a linear classifier, a matrix $W$, that is applied to each $x \in X$ in order to predict the correct category $c$ with the greatest possible accuracy.
-In other words, $Wx$ defines a distribution over the set of categories $C$ and we assign $x$ to the class $c \in C$ which is allotted the greatest probability by $Wx$.
+%In other words, $Wx$ defines a distribution over the set of categories $C$ and we assign $x$ to the class $c \in C$ which is allotted the greatest probability by $Wx$.
 Note that the classifier's accuracy must be greater than that achieved by guessing the majority category, otherwise $x$ contains no linear information relevant for the categorization task and thus no guarding function is needed.
+%
 Once $W$ is determined, for any $x \in X$ we can remove the information that $W$ uses to predict $c$ by projecting $x$ onto the null-space of $W$, $N(W) = \{x | Wx=0\}$.
@@
 However, this process does not necessarily remove all of the linear information in $x$ that could be used to predict $c$. 
-For example, $x$ may contain redundant information and $W$ may have only used one set of this information for its prediction.
+%For example, $x$ may contain redundant information and $W$ may have only used one set of this information for its prediction.
 In this case, the redundant information would still be present in $\hat{x}$.
+%
 Thus, we must repeat the above process, defining a new linear classifier $\hat{W}$ that uses $\hat{x}$  to predict $c$.
@@
 We iteratively apply this process until no linear information remains in $\hat{x}$, i.e.
-a linear classifier is unable to predict the correct category $c$ with any probability greater than that achieved by guessing the majority class.
+a linear classifier is unable to predict the correct category $c$ with any probability greater than that achieved by guessing the majority class\footnote{The stopping criterion follows \citealp{amnesia}, iterations will stop if the linear classifier achieve within one point above majority accuracy on development set.}.
 The final $\hat{x} = P_n(P_{n-1}(\dots P_1(x)))$ contains no linear information about the categories in $C$ and we call $P(x) = P_n(P_{n-1}(\dots P_1(x)))$ the guarding function.
 
-We will pair the INLP method with the probing tasks described in sections \ref{sec:syntactic} and \ref{sec:semantics} in order to create two guarding functions that will enable us to isolate the linear components of BERT embeddings that contain syntax-specific and semantics-specific information.
+% adding the formula to efficiently find the intersection
+
+The projection matrix $P$ derived by matrix multipcliaitons $P_n\cdot P_{n-1} \dots P_1$ can be susceptible to numerical errors, therefore \citealp{inlp} utilized the following formula using \textit{rowspace projection}\footnote{$P_R(W_i)$ in the formula means row space projection of weight matrix $W$.}  proposed by \citealp{ben-israel_2015} to compute the intersection of null spaces of weight matrix.
+Then projection matrix $P$ is derived from null space projection of the intersection, $P=P_{\cap_{i=1}^n N(W_i)}$, instead.
+Our experiment follow the same computation.
+
+\begin{quote}
+    $\cap_{i=1}^n N(W_i) = N(\sum_{i=1}^n(P_R(W_i)) )$ 
+\end{quote}
+
+% We will pair the INLP method with the probing tasks described in sections \ref{sec:syntactic} and \ref{sec:semantics} in order to create two guarding functions that will enable us to isolate the linear components of BERT embeddings that contain syntax-specific and semantics-specific information.
 
@@
 The gold standard contains 5,438 sentences with annotations that are manually verified while the silver standard contains 62,739 sentences with autogenerated annotations.
-All of our experiments are conducted on gold standard data.
+All of our experiments are conducted on gold standard data with standard train/dev/test split\footnote{Gold standard dataset contains total of 34706 words, with 80\% of training and dev data, and 20\% of testing data}.
 
@@
 We follow a similar procedure and apply a CCG parser \citep{ccg_tagger} to develop word-level CCG tags.
-Once we obtain both CCG tags and semantic tags for the dataset, we can perform the syntactic and semantic probing tasks as desired.
+Once we obtain both CCG tags and semantic tags for the dataset, we can perform word-level syntactic and semantic probing tasks as desired.
+The total number of labels in CCG tags and Semantics tags are 159 and 72 respectively.
 
-\subsection{Syntactic probing task}
-\label{sec:syntactic}
+% adding the train dev test split and word count
+% train words = 24986
+% dev words 2781
+% test words = 6939
 
-The syntactic probing task involves training a linear classifier on the final layer BERT embeddings in order to predict the CCG tag associated with each word.
-We will use this classifier in the INLP algorithm in order to create a guarding function for the information that is necessary to complete the CCG labeling task.
-For a given embedding, $v_{orig}$, the projection that results from applying this guarding function, $P_{syn}$, to the embedding will represent the non-syntactic information contained in the embedding and will from now on be referred to as the ``non-syntactic component'' of the embedding, $v_{no syn} = P_{syn} v_{orig}$.
-We can then determine the ``syntactic component'' of the embedding by taking the difference of the embedding vector with the non-syntactic component, $v_{syn} = v_{orig} - v_{no syn}$.
+\subsection{Probing tasks}
+\label{sec:probing}
+% num tags = 159
+The  probing task involves training a linear classifier on the final layer BERT embeddings in order to predict the CCG tag or semantic tag associated with each word.
+We will use this classifier in the INLP algorithm in order to create a guarding function for the information that is necessary to complete the task.
+Take CCG tag as an example: for a given embedding, $v_{orig}$, the projection that results from applying this guarding function, $P_{syn}$ or $P_{sem}$, to the embedding will represent the non-syntactic information contained in the embedding and will from now on be referred to as the ``non-syntactic component'' of the embedding, $v_{no syn} = P_{syn} v_{orig}$.% We can then determine the ``syntactic component'' of the embedding by taking the difference of the embedding vector with the non-syntactic component, $v_{syn} = v_{orig} - v_{no syn}$.
 
-\subsection{Semantic probing task}
-\label{sec:semantics}
-
+% num tags = 72
 Similar to the above, the semantic probing task involves training a linear classifier on the final layer BERT embeddings in order to predict the semantic tag associated with each word.
 This classifier is used in the INLP algorithm in order to create a guarding function, $P_{sem}$, for the information necessary to complete the semantic tag labeling task.
-As described in the syntactic probing task section, we use the resulting guarding function to decompose the original embedding into a ``non-semantic component'', $v_{no sem} = P_{sem} v_{orig}$, and a ``semantic component'', $v_{sem} = v_{orig} - v_{no sem}$.
+As described in the syntactic probing task, we use the resulting guarding function to compute a ``non-semantic embedding'', $v_{no sem} = P_{sem} v_{orig}$.
+%, and a ``semantic component'', $v_{sem} = v_{orig} - v_{no sem}$.
 
@@
 We will assess each of the non-syntactic and non-semantic embedding types, the original BERT embeddings and the embeddings created by randomly removing directional information on the CCG and semantic labeling tasks that were used in the probes.
+All the embeddings are listed in table \ref{description}.
 
@@
 
-\subsection{Layer-wise evaluation}
-In addition to the final layer BERT embeddings, we perform a similar analysis on the embeddings derived from different layers of the BERT architecture, in order to determine the separability of these information sets at each layer.
-For embedding $v_{orig_i}$ from layer $i,$ a linear classifier is trained for each probing task to acquire guarding functions $P_{syn_i}$ and $P_{sem_i},$ respectively.
-Applying these projection functions, we are able to acquire $v_{nosyn_i}$ and $v_{nosem_i}.$ Subtracting them from the original embedding, we get the semantic representation $v_{sem_i}$ and the syntatic representation $v_{syn_i}.$ We also randomly remove the same number of dimensions in the original embedding for comparison.
-
-By comparing the experiment results across different information sets and different layers, we hope to better understand how BERT processes different types of linguistic information throughout the encoding process.
-
 \begin{table}[ht]
@@
         $v_{nosyn}= P_{syn} v_{orig}$ & Gained after INLP with syntactic task \\
-        $v_{sem} = v_{orig}-v_{nosem}$ & Semantic representation \\
-        $v_{syn} = v_{orig}-v_{nosyn}$ & Syntactic representation \\
-        Rand($v$, $n$) & embeddings $v$ with n random directions removed \\
-        Subscript $i$ & Objects relating to the $i$-th layer of the BERT architecture rather than the final layer\\
+        % $v_{sem} = v_{orig}-v_{nosem}$ & Semantic representation \\
+        % $v_{syn} = v_{orig}-v_{nosyn}$ & Syntactic representation \\
+        Rand($v$, $n$) & Embeddings $v$ with n random directions removed \\
+        % Subscript $i$ & Objects relating to the $i$-th layer of the BERT architecture rather than the final layer\\
         %$v_{syn-sem} = P_{sem} v_{syn}$ &Contains syntactic features only \\
@@
 
+\subsection{Layer-wise evaluation}
+\label{layer}
+In addition to the final layer BERT embeddings, we perform a similar analysis on the embeddings derived from different layers of the BERT architecture, in order to determine the separability of these information sets at each layer.
+For embedding $v_{orig_i}$ from layer $i,$ a linear classifier is trained for each probing task to acquire guarding functions $P_{syn_i}$ and $P_{sem_i},$ respectively.
+Applying these projection functions, we are able to acquire $v_{nosyn_i}$ and $v_{nosem_i}.$ Subtracting them from the original embedding, we get the semantic representation $v_{sem_i}$ and the syntatic representation $v_{syn_i}.$ We also randomly remove the same number of dimensions in the original embedding for comparison.
+
+By comparing the experiment results across different information sets and different layers, we hope to better understand how BERT processes different types of linguistic information throughout the encoding process.
+
 \section{Results}
@@
 For the embeddings derived from projection matrices, we record the number of removed directions using $Null(P)$, where $P$ is the projection matrix.
-\footnote{For embeddings not derived from matrix multiplication, we obtain the number from $Null(M)$ where $M$ is the embedding matrix with size (768,instance number).
+%\footnote{For embeddings not derived from matrix multiplication, we obtain the number from $Null(M)$ where $M$ is the embedding matrix with size (768,instance number).
 This is not necessarily equivalent to the number of direction removed.
@@
 So we decided to remove all embeddings derived in this manner from our evaluation process until we are able to determine the source of this behavior.
-This means that $v_{syn}$ and $v_{sem}$ will not be included in our final evaluation.} Then we create embeddings with the same number of directions randomly removed and train the linear classifiers on these embeddings.
+This means that $v_{syn}$ and $v_{sem}$ will not be included in our final evaluation.}
+Then we create embeddings with the same number of directions randomly removed and train the linear classifiers on these embeddings.
 The testing accuracies from our experiments can be found in table \ref{role description}.
@@
     \hline
-
-    Rand($v_{orig}, |v_{nosem}|$) & 77 & $84.06\%$ & $87.71\%$ \\
-    Rand($v_{orig}, |v_{nosyn}|$) & 159 & $82.26\%$ & $87.57\%$ \\
+    Majority Guess & N/A&$16.57\%$  &$22.93\%$ \\ \hline
+    Rand($v_{orig}, |v_{nosem}|$) & 792 & $16.57\%$ & $22.93\%$ \\
+    Rand($v_{orig}, |v_{nosyn}|$) & 795 & $16.57\%$ & $22.93\%$ \\
     %Rand($|v_{syn-sem}|$) & 81 & $83.11\%$ & $87.61\%$ \\
@@
     \hline
-    $v_{nosem}$ & 77 & $27.93\%$ & $17.28\%$ \\
-    $v_{nosyn}$ & 159 & $23.76\%$ & $49.43\%$ \\
+    $v_{nosem}$ & 792 & $16.57\%$ & $22.93\%$ \\
+    $v_{nosyn}$ & 795 & $16.57\%$ & $22.93\%$ \\
     %$v_{syn-sem}$ & 81 & $34.24\%$ & $33.91\%$ \\

### 2022-12-16 paper/paper.tex
edit result, discussion, and conclusion

@@
 \label{sec:result}
-\subsection{Final layer evaluation}
-We first evaluate our two tasks on the original embeddings, and determine that linear classifiers can successfully predict both CCG tags and semantic tags (around 85\% testing accuracy), as shown in table \ref{role description}.
+We first evaluate our two tasks on the original embeddings, and determine that linear classifiers can successfully predict both CCG tags and semantic tags (around 85\% and 89\% testing accuracy, respectively), as shown in table \ref{role description}.
 We then apply the INLP method to derive the guarding matrices $P_{syn}$ and $P_{sem}$, which are used to project the original embeddings onto the complements of the syntactic information sub-space and the semantic information sub-space.
@@
 
-To ensure a fair assessment of the impact of the information loss, we conduct experiments for which we start with the original BERT embeddings and randomly remove the same number of directions that our derived embeddings lost.
-For the embeddings derived from projection matrices, we record the number of removed directions using $Null(P)$, where $P$ is the projection matrix.
+To ensure a fair assessment of the impact of the information loss, we conduct experiments for which we start with the original BERT embeddings and randomly remove the same number of directions that our derived embeddings lost, and train the linear classifiers on these embeddings.
+The testing accuracies from our experiments can be found in table \ref{role description}.
+Curiously, our linear classifiers for evaluation tasks cannot do bettet than majority class.
+%For the embeddings derived from projection matrices, we record the number of removed directions using $Null(P)$, where $P$ is the projection matrix.
 %\footnote{For embeddings not derived from matrix multiplication, we obtain the number from $Null(M)$ where $M$ is the embedding matrix with size (768,instance number).
@@
 This means that $v_{syn}$ and $v_{sem}$ will not be included in our final evaluation.}
-Then we create embeddings with the same number of directions randomly removed and train the linear classifiers on these embeddings.
-The testing accuracies from our experiments can be found in table \ref{role description}.
+
+On the intermediate layers, linear classifiers are generally able to achieve a test accuracy greater than 85\% for both CCG tagging and semantics tagging.
+However, we observe the same majority case accuracy across all layers for each evaluation task.
+Evaluations of Rand($v_i, |v_{nosyn_i}|$) and Rand($v_i, |v_{nosem_i}|$) result in the same majority class accuracy.
 
@@
 
-\subsection{Intermediate Layer Evaluation}
-
-Linear classifiers are generally able to achieve a test accuracy greater than 90\% for both CCG tagging and semantics tagging.
-However, after applying the INLP algorithm and projecting the layer-wise embeddings to the nullspaces of the information sets, we observe that we get the same results across all layers for each evaluation task, with the accuracy being the majority class of the evaluation task.
-Evaluations of Rand($v_i, |v_{nosyn_i}|$) and Rand($v_i, |v_{nosem_i}|$) result in the same majority class accuracy.
-
-Upon a close inspection of the INLP process and the projections of the original intermediate layer embeddings, $v_{nosem_i}$ and $v_{nosyn_i}$, we realize that, the INLP process continues to run even if it already removes more ranks than BERT’s hidden size, which is 768 in our case, because the desired dev accuracy is still not met.
-Once the rank of the projection matrix reaches the limit, the INLP process simply reduces the magnitude of each elements in the embeddings.
-In some cases, the process eventually zeroes out the embeddings, which explains the identical yet trivial result we get from the evaluation task.
-
 %In this section, we consider what each evaluation task tells us about the embeddings that are being probed.
@@
 \label{sec:discussion}
-The INLP method successfully guards the information used in the probing task.
+%The INLP method successfully guards the information used in the probing task.
 When performing CCG tagging, the model accuracy drops significantly when comparing the $v_{nosyn}$ embeddings with the Rand($|v_{nosyn}|$) embeddings (from 82.26\% to 23.76\%), and similarly for semantic tagging (from 87.71\% to 17.28\%).
@@
 This suggests that the syntactic and semantic information contained in the original BERT embeddings is not highly concentrated and that removing a small amount of one information set will not have a significant impact on the classification task.
-
 %To measure the importance of semantic information for the syntactic task, we compare the performance of CCG tagging with embeddings $v_{nosem}$ to that of embeddings $v_{nosyn}$.
@@
 The result suggests that the semantics task is less dependent on the syntactic information than the opposite direction, which contradicts the assumption that syntactic information will be more pertinent to semantics task than vice versa.
-
-To measure the importance of semantic information for the syntactic task, we compare the performance of the CCG tagging classifier on $v_{nosem}$ with that of Rand($v_orig, |v_{nosem}|$) and see a performance drop of 56.13\%.
+%To measure the importance of semantic information for the syntactic task, we compare the performance of the CCG tagging classifier on $v_{nosem}$ with that of Rand($v_orig, |v_{nosem}|$) and see a performance drop of 56.13\%.
 Similarly, we can measure the importance of syntactic information for the semantic task by comparing the performance of the semantic tagging classifier on $v_{nosyn}$ with that of Rand($v_{orig}, |v_{nosyn}|$) and see a performance drop of 38.14\%.
@@
 This suggests that the syntactic and semantic information in BERT embeddings is not easily disentangled.
+%What is surprising is that the loss of semantic information has a more significant impact on the syntactic classification task than the loss of syntactic information does on the semantic tagging task.
+These results suggest that the semantic task is less dependent on syntactic information than the opposite direction, which challenges the assumption that syntactic information is more pertinent to semantic comprehension than vice versa.
+%
+We are surprised to find out that we are unable to fully remove the syntactic/semantic information from the embeddings by training the linear classifier to make prediction that is no better than the majority, without removing more ranks than BERT's hidden size.
+However, removing more ranks than BERT’s hidden size, whether through the INLP algorithm or randomly, results in a degenerate embedding where every element is reduced to an extremely small magnitude that the linear probe on the evaluation task will only reach the majority class accuracy.
+This is true on all layers of BERT.
+This seems to reveal that, the target information is not linearly separable from the original embeddings.
 
-What is surprising is that the loss of semantic information has a more significant impact on the syntactic classification task than the loss of syntactic information does on the semantic tagging task.
-These results suggest that the semantic task is less dependent on syntactic information than the opposite direction, which challenges the assumption that syntactic information is more pertinent to semantic comprehension than vice versa.
-
-As an unexpected finding of our experiment, we are unable to fully remove the syntactic/semantic information from the intermediate embedding by training the linear classifier to make prediction that is no better than the majority, without removing more ranks than BERT's hidden size, on all intermediate layers of BERT.
-However, removing more ranks than BERT’s hidden size, whether through the INLP algorithm or randomly, results in a degenerate embedding where every element is reduced to an extremely small magnitude that the linear probe on the evaluation task will only reach the majority class accuracy.
-This seems to reveal that, on the intermediate layers of BERT, the target information is not linearly separable from the original embedding.
-Besides using the majority class accuracy as the stopping condition, researchers hoping to use INLP to guard information from BERT embeddings should also make sure the loop stops before removing too many ranks for non-trivial results.
+Upon a close inspection of the INLP process and the projections of the original embeddings, $v_{nosem}$ and $v_{nosyn}$, we realize that, the INLP process continues to run even if it already removes more ranks than BERT’s hidden size, which is 768 in our case, because the desired dev accuracy is still not met.
+Once the rank of the projection matrix reaches the limit, the INLP process simply reduces the magnitude of each elements in the embeddings.
+In most cases, the process eventually zeroes out the embeddings, which explains the identical yet trivial result we get from the evaluation tasks across all layers.
 
@@
 Our experiment has confirmed that linear classifiers can perform CCG tagging and semantic tagging on the Parallel Meaning Bank data set \citep{pmbData} with a fairly high rate of success.
-We then employed INLP and successfully guarded the information contained in BERT embeddings that linear classifiers use to perform the aforementioned classification tasks.
+We then employed INLP to guard the information contained in BERT embeddings that linear classifiers use to perform the aforementioned classification tasks.
 
 Using the INLP-derived guarding functions we were able to explore the importance and separability of the syntactic and semantic information contained in BERT embeddings.
-We evaluated the classification tasks on various derived embeddings and concluded that not only is the linear syntactic and semantic information essential for their respective classification tasks, these information sets are also very important for the opposing classification tasks as well.
-Thus the two information sets are not easily separated.
-Somewhat surprising, we determined that the semantic information was more influential on the success of the syntactic classifier than the other way around.
+We evaluated the classification tasks on various derived embeddings and concluded that not only is the syntactic and semantic information essential for their respective classification tasks, these information sets are also very crucial for the opposing classification tasks as well.
+Thus the two information sets are not linearly separable from the original embeddings.
+Attempts to remove the information sets by INLP will result in projection matrices whose ranks are higher than the rank of embeddings.
+Applying the projection matrices will result in degenerate embeddings where all information is removed.
 %something about linear separability with last two embeddings.
 
-By performing a similar analysis on embeddings derived from different layers of BERT architecture, we discovered that the syntactic information set and semantic information set are not linearly separable from the original intermediate layer embeddings.
-Attempting to completely remove these information sets will remove too much that the embeddings become degenerate.
-
+Our results indicate that besides using the majority class accuracy as the stopping condition, researchers hoping to use INLP to guard information from BERT embeddings should also make sure the loop stops before too many ranks are removed.
+If the rank of the projection matrix $P$ is higher than the rank of the embedding matrix, only trivial results will be achieved.
 %After gained guard matrices that remove task specific informations from the embeddings, we examined the separablity of semantics and syntactic representation by evaluation tasks on various embeddings.

### 2022-12-16 paper/paper.tex
adding final changes to the paper

@@
 
-However, this process does not necessarily remove all of the linear information in $x$ that could be used to predict $c$. 
+%However, this process does not necessarily remove all of the linear information in $x$ that could be used to predict $c$. 
 %For example, $x$ may contain redundant information and $W$ may have only used one set of this information for its prediction.
@@
 %
-Thus, we must repeat the above process, defining a new linear classifier $\hat{W}$ that uses $\hat{x}$  to predict $c$.
+%Thus, we must repeat the above process, defining a new linear classifier $\hat{W}$ that uses $\hat{x}$  to predict $c$.
 If $\hat{W}$ is still able to predict $c$ with a greater than majority class guess accuracy, then we know that $\hat{x}$ contained linear information about $c$.
@@
 % num tags = 159
-The  probing task involves training a linear classifier on the final layer BERT embeddings in order to predict the CCG tag or semantic tag associated with each word.
+The  probing task involves training a linear classifier\footnote{The linear classifier will use Adam as optimizer \cite{adam} implemented in torch \cite{torch}.
+Therefore the total number of parameters will be dimensions of BERT embeddings $\cdot$ number of labels, which will be 122112 for syntactic task and 55296 for semantics.
+} on the final layer BERT embeddings in order to predict the CCG tag or semantic tag associated with each word.
 We will use this classifier in the INLP algorithm in order to create a guarding function for the information that is necessary to complete the task.

### 2022-12-16 paper/paper.tex
adding ethincal consideration

@@
 
+\section*{Ethical Consideration}
+\subsection*{Scientific Artifacts}
+Below is a list of scientific artifact used along with the names of their licenses. None of these artifacts restrict use for research purposes. We are not using this work for commercial purposes.
+
+\begin{itemize}
+    \item Parallel Meaning Bank\cite{pmbData} -  ODC-BY 1.0
+    \item BERT\cite{bert}  – Apache License Version 2.0
+    \item Deccg\cite{ccg_tagger} - MIT License
+    \item PyTorch\cite{torch} - PSF License Agreement
+\end{itemize}
+
+Of the artifacts used, the only data source is the Parallel Meaning Bank, which does not contain any sensitive information.
+We used only English data from the PMB, which is intended for syntax and semantics research and makes no attempt, as far as we are aware, to balance the demographic groups represented.
+This is not a problem for our work because we are not using the PMB to generate anything.
+
+\section*{Computational Experiments}
+The parameters used are the word embeddings and syntactic/semantic tags; this yields 122,112 parameters for the syntactic model and 55,296 parameters for the semantic model.
+The models altogether took two GPU models to run, and the results reflect a one-time run on a computing cluster.
+The run time is around 5 minutes for a single probing task.
 \bibliography{acl2020.bib}
